---
title: Cell density analysis
author: "AS Huffmyer"
date: '2025'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for symbiont cell counts and larval size. 

# Set up 

```{r}
if (!require("dplyr")) install.packages("dplyr")
if (!require("readr")) install.packages("readr")
if (!require("stringr")) install.packages("stringr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("grid")) install.packages("grid")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lattice")) install.packages("lattice")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lsmeans")) install.packages("lsmeans")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("car")) install.packages("car")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("corrr")) install.packages("corrr")
if (!require("GGally")) install.packages("GGally")
if (!require("cowplot")) install.packages("cowplot")

library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(car)
library(kableExtra)
library(ggcorrplot)
library(corrr)
library(GGally)
library(cowplot)
library(tidyverse)
```

# Cell Density 

## Load data 

Import data  
```{r}
sym_counts <- read_csv("data/size_cells/cell-density.csv")%>%select(!Notes)
numbers <- read_csv("data/size_cells/larval_counts.csv")

data <- left_join(sym_counts, numbers)
head(data)
```

## Calculations 

Calculate total cells.  
```{r}
# Calculate mean counts for each sample
df_cells <- data %>%
  dplyr::select(Tube.ID, Squares_counted, Resuspension.Vol.uL, Number.Larvae, Parent, Date,  matches("Count[1-6]")) %>%
  gather("rep", "count", -Tube.ID, -Squares_counted, -Resuspension.Vol.uL, -Number.Larvae, -Date, -Parent) %>%
  group_by(Tube.ID, Squares_counted, Number.Larvae, Parent, Date, Resuspension.Vol.uL) %>%
  dplyr::summarise(mean_count = mean(count, na.rm = TRUE))

# Normalize counts by homogenate volume (ul)
df_cells <- df_cells %>%
  mutate(cells.mL = (mean_count * 10000) / Squares_counted,
         cells = cells.mL * (Resuspension.Vol.uL/1000))
```

This generated the total number of cells in each tube. 

## Normalize to number of larvae 

Divide the total number of cells by the resuspension volume to obtain cells per uL 

```{r}
df_cells$cells.larvae=df_cells$cells/df_cells$Number.Larvae
```

## Visualize data   

Plot cells per unit protein data with mean and standard error.   

```{r}
plot1<-df_cells %>%
  ggplot(aes(x = as.character(Date), y = cells.larvae, color = Parent)) +
  labs(x = "Date",y = "Symbiont Cells per Larva") +
  scale_color_manual(values=c("orange", "red4", "darkgray"))+
  geom_jitter(width = 0.05) + # Plot all points
  geom_smooth(aes(group=1), method="lm", colour="black", se=TRUE)+
  theme_classic();plot1

ggsave(plot1, filename="figures/cells_size/cell_density_larva.png", width=6, height=5)
```

Values are within range of Huffmyer et al. 2025 (Current Biology) for cells per individual in Montipora capitata at similar ages.  

Present means and standard error of each group in a table.  

```{r}
df_cells%>%
  group_by(Parent, Date)%>%
  dplyr::summarise(n=length(cells.larvae),
            Mean=format(round(mean(cells.larvae), 0), 0), 
            SE=format(round(sd(cells.larvae)/sqrt(length(cells.larvae)),0),0))%>%
  kbl(caption="Descriptive statistics of Symbiodiniaceae cell densities per larva")%>%
  kable_classic(full_width=FALSE, html_font="Arial")%>%
  row_spec(0, bold = TRUE) 
```

Plot with mean and standard errors.  

```{r}
df_cells %>%
  group_by(Parent, Date)%>%
  dplyr::summarise(mean=mean(cells.larvae), se=(sd(cells.larvae)/sqrt(length(cells.larvae))))%>%
  
  ggplot(aes(x = Date, y = mean, color = Parent)) +
  labs(x = "",y = "Cell Density per Larva") +
  scale_color_manual(values=c("orange", "red4", "darkgray"))+
  geom_point()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  theme_classic()
```

Variable over time but with a general increase in densities. 

## Statistical analysis  

Run anova on cells per larvae by parent and date    
```{r}
model<-aov(cells.larvae ~ Parent * Date, data=df_cells)
summary(model)

qqPlot(residuals(model))
```

No significant difference in cell density by parent over time.  

# Larval Size 

```{r}
size <- read_csv("data/size_cells/larval_size.csv")%>%select(!Notes)%>%select(!Larvae.ID)
```

Calculate size as prolate spheroid (Huffmyer et al. 2025). 

V = 4 𝜋𝑎^2𝑏, where a is 1/2 width and b is 1/2 length

```{r}
size$volume <- 4 * pi * (0.5*size$Width.mm)^2 * (0.5 * size$Length.mm)
```

```{r}
hist(size$volume)
```

Values are in the same range as Huffmyer et al. 2025 for larval stages.    

Summarise as a mean for each tube. 

```{r}
size<-size%>%
  group_by(Tube.ID)%>%
  dplyr::summarise(volume = mean(volume))
```

Merge into data frame. 

```{r}
merged<-left_join(df_cells, size)
```

Plot size. 

```{r}
plot2<-merged %>%
  ggplot(aes(x = as.character(Date), y = volume, color = Parent)) +
  labs(x = "Date",y = "Larval Volume") +
  scale_color_manual(values=c("orange", "red4", "darkgray"))+
  geom_jitter(width = 0.05) + # Plot all points
  geom_smooth(aes(group=1), method="lm", colour="black", se=TRUE)+
  theme_classic();plot2

ggsave(plot2, filename="figures/cells_size/volume_larvae.png", width=6, height=5)
```

# Size normalize cell densities 

Size normalize cell density to larval size. 

```{r}
merged$cells.mm3<-merged$cells.larvae/merged$volume
```

Plot size normalized densities. 

```{r}
plot3<-merged %>%
  ggplot(aes(x = as.character(Date), y = cells.mm3, color = Parent)) +
  labs(x = "Date",y = "Size-Normalized Cell Density (cells mm3)") +
  scale_color_manual(values=c("orange", "red4", "darkgray"))+
  geom_jitter(width = 0.05) + # Plot all points
  geom_smooth(aes(group=1), method="lm", colour="black", se=TRUE)+
  theme_classic();plot3

ggsave(plot3, filename="figures/cells_size/cells_volume.png", width=6, height=5)
```