---
title: "Larval Respirometry Plotting and Analysis - Hawaii 2023 - F1F2 project" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
```


# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pr_data<-read.csv("output/F1F2_sdr/F1F2_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pr_data<-pr_data[!is.na(pr_data$Type),]

#format columns
pr_data$Group<-as.factor(pr_data$Group)
pr_data$SDR<-as.factor(pr_data$SDR)
pr_data$Plate<-as.factor(pr_data$Plate)
pr_data$Temperature<-as.factor(pr_data$Temperature)
```

Change level of par factor. 
```{r}
pr_data$PAR<-factor(pr_data$PAR, levels=c("0.1", "50", "100", "500", "0.2"))
```

Look for outliers in the data.  

```{r}
boxplot(pr_data$P.nmol.org.min)

boxplot(pr_data$P.nmol.org.min~pr_data$PAR)
```

Set values at 0 for light runs if <0 indicating no photosynthesis occurred. Set values at 0 for dark runs if >0 indicating no respiration occurred. 
```{r}
pr_data<-pr_data%>%
  mutate(P.nmol.org.min=if_else(P.nmol.org.min<0 & PAR==50, 0, 
                      if_else(P.nmol.org.min<0 & PAR==100, 0, 
                              if_else(P.nmol.org.min<0 & PAR==500, 0, 
                                      if_else(P.nmol.org.min>0 & PAR==0.1, 0, 
                                              if_else(P.nmol.org.min>0 & PAR==0.2, 0, P.nmol.org.min))))))

boxplot(pr_data$P.nmol.org.min~pr_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/F1F2_sdr/runs_temp.csv")
pr.temps = subset(pr.temps, select = -c(X) ) #remove empty column

pr_data$Plate<-as.integer(pr_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pr_data<-left_join(pr_data,pr.temps)

#round to 0.1°C 
pr_data<-pr_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

# Plot photosynthesis over PAR and temperature treatments  

Plot data with means   
```{r}
pr_plot1<-pr_data %>%
    group_by(PAR, Group, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Group))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Group), colour=Group), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Group)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("log P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1

ggsave("figures/F1F2_sdr/pr_means.png", pr_plot1, dpi=300, w=7, h=5, units="in")
```

Display by individual plate (duplicate plates at each temperature).  
```{r}
pr_plot1a<-pr_data %>%
    group_by(PAR, Group, Plate)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Group))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Group), colour=Group), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Group)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("log P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1a
```

View by plate with a loess line. 
```{r}
pr_plot2b<-pr_data %>%
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=Group)) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Group, group=Group), method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("log P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot2b
```
Plates look similar and consistent. 

Plot data with individual data points.    
```{r}
pr_plot3<-pr_data %>%
    
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=interaction(PAR, Group))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("log P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot3

ggsave("figures/F1F2_sdr/pr_dots.png", pr_plot3, dpi=300, w=7, h=5, units="in")
```

## Conduct anova 

Run anova on effect of PAR, group, and temperature. 
```{r}
summary(aov(P.nmol.org.min~PAR*Group*Temperature, data=pr_data))
```
There are strong effects of temperature and light and an interaction between light and group. 

Run model for 27 degrees. 
```{r}
model<-pr_data%>%
  filter(Temperature==27)%>%
  aov(P.nmol.org.min~PAR*Group, data=.)

summary(model)
```
Strong effect of light at 27C. 

Run model for 33
```{r}
model<-pr_data%>%
  filter(Temperature==33)%>%
  aov(P.nmol.org.min~PAR*Group, data=.)

summary(model)
```
Strong effect of light, group, and light x group at 33C. 

Run model for 36
```{r}
model<-pr_data%>%
  filter(Temperature==36)%>%
  aov(P.nmol.org.min~PAR*Group, data=.)

summary(model)
```
Strong effect of light at 36C. 

# Calculate R, P, and P:R 

Calculate gross and net photosynthesis, light enhanced respiration (LER), dark respiration (Rd), and P:R ratio at low (50 PAR), medium (100 PAR), and high (500 PAR) light. 

Here I am calculating P:R ratios by dividing gross photosynthesis by dark respiration. I tried dividing by LER, but due to a high number of 0's in the data, we loose most of the calculations.
```{r}
pr_data_calc<-pr_data%>%
  select(Date, Chamber.ID, Plate, Run, Temperature, Group, PAR, P.nmol.org.min)%>%
  pivot_wider(names_from=PAR, values_from=P.nmol.org.min)%>%
  mutate(ler=`0.1`)%>%
  mutate(rd=`0.2`)%>%
  mutate(ler_inverse=ler*-1)%>%
  mutate(rd_inverse=rd*-1)%>%
  mutate(p_net_low=`50`)%>%
  mutate(p_net_medium=`100`)%>%
  mutate(p_net_high=`500`)%>%
  mutate(p_gross_low=p_net_low-ler)%>%
  mutate(p_gross_medium=p_net_medium-ler)%>%
  mutate(p_gross_high=p_net_high-ler)%>%
  mutate(ratio_low=p_gross_low/(rd_inverse))%>%
  mutate(ratio_medium=p_gross_medium/(rd_inverse))%>%
  mutate(ratio_high=p_gross_high/(rd_inverse))%>%
  pivot_longer(names_to="Metric", values_to="Value", cols=c(ler:ratio_high))%>%
  select(!c(`0.1`, `100`, `500`, `0.2`))
```

Filter P:R ratio values. 
```{r}
boxplot(pr_data_calc$Value)

pr_data_calc<-pr_data_calc%>%
  filter(!c(Value<0 & Metric=="ratio_low"))%>% 
  filter(!c(Value<0 & Metric=="ratio_medium"))%>%
  filter(!c(Value<0 & Metric=="ratio_high"))

boxplot(pr_data_calc$Value)
```

Set level of factors. 
```{r}
levels(as.factor(pr_data_calc$Metric))

pr_data_calc$Metric<-factor(pr_data_calc$Metric, levels=c("p_gross_low", "p_gross_medium", "p_gross_high", "p_net_low", "p_net_medium", "p_net_high", "ratio_low", "ratio_medium", "ratio_high", "rd", "rd_inverse", "ler", "ler_inverse"))
```

View metrics. 
```{r}
metric_plot1<-pr_data_calc%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_medium")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
    filter(!Metric=="rd")%>%
    filter(!Metric=="rd_inverse")%>%
    filter(!Metric=="ler_inverse")%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1

ggsave("figures/F1F2_sdr/pr_metrics.png", metric_plot1, dpi=300, w=7, h=5, units="in")
```

View metrics across temperatures. 
```{r}
metric_plot2<-pr_data_calc%>%
    filter(Metric %in% c("ratio_high", "ratio_low", "ratio_medium"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2

ggsave("figures/F1F2_sdr/pr_ratio.png", metric_plot2, dpi=300, w=7, h=5, units="in")
```

View metrics. 
```{r}
metric_plot3<-pr_data_calc%>%
    filter(Metric %in% c("ler_inverse", "rd_inverse"))%>%
    #filter(!c(Metric=="rd_inverse" & Value<0))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot3

ggsave("figures/F1F2_sdr/ler_rd.png", metric_plot3, dpi=300, w=7, h=5, units="in")
```

## Run anova on differences. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ratio_high")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")

```
P:R Ratio at high light is affected by temperature and decreases with temperature. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ratio_medium")%>%
  aov(Value~Group*Temperature, data=.)

TukeyHSD(model, "Temperature")

summary(model)
```
P:R ratios decrease with temperature at medium light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ratio_low")%>%
  aov(Value~Group*Temperature, data=.)

TukeyHSD(model, "Temperature")

summary(model)
```
P:R ratios decrease with temperature at high light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ler_inverse")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Light enhanced respiration is reduced with high temperature.

```{r}
model<-pr_data_calc%>%
  filter(Metric=="rd_inverse")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")

TukeyHSD(model, "Group")
```
Dark respiration is not affected by temperature, but it is affected by group. F2 NB has lower respiration than F1 NB. F2 NB also has lower Rd than F2 WT.   

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_gross_high")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
P gross at high light decreases with temperature and is lowest at 36 degrees. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_gross_medium")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
At medium light, P gross decreases with temperature and is lowest at 36 degrees. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_gross_low")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Gross photosynthesis at low light is affected by temperature and is lowest at 36 degrees.

These measurements all make sense that P:R ratio is reduced with temperature. Respiration stays the same, but photosynthesis is decreasing. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_net_high")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Net photosynthesis at high light is affected by temperature, decreases at 36C. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_net_medium")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
At medium light P net decreases at 36C. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_net_low")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Group")
```
P net is lower in F1 NB at low light, but only slightly. 

Test effect of light on P:R ratio. 
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("ratio_low", "ratio_medium", "ratio_high"))%>%
  aov(Value~Metric*Temperature*Group, data=.)

summary(model)

TukeyHSD(model, "Metric")
```
P:R ratios are significantly different by light and temperature. There are higher P:R ratios at high light but only when at 27C. 

Test effect of light on P gross.  
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("p_gross_low", "p_gross_medium", "p_gross_high"))%>%
  aov(Value~Metric*Temperature*Group, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")
TukeyHSD(model, "Group")

emm<-emmeans(model, ~ Group | Temperature)
pairs(emm)
```
P gross increases with light and is different between high and low light (increases with light). P gross decreases with temperature strongly. There is also an interaction between temperature and group. There are differences between F1 WT and F2 NB. 

Test effect of light on P net.   
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("p_net_low", "p_net_medium", "p_net_high"))%>%
  aov(Value~Metric*Temperature*Group, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")
```
P net is affected by light and increases with each light level. P net is also affected by temperature and decreases at  36C. 
