---
title: "Larval Respirometry Plotting and Analysis - Hawaii 2023 - F1F2 project" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Need to QC individual export plots 
Need to run each metric (ratio, Pnet, Pgross, Rd, LER) against PAR - currently separated by PAR. 

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
```


# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pr_data<-read.csv("output/F1F2_sdr/F1F2_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pr_data<-pr_data[!is.na(pr_data$Type),]

#format columns
pr_data$Group<-as.factor(pr_data$Group)
pr_data$SDR<-as.factor(pr_data$SDR)
pr_data$Plate<-as.factor(pr_data$Plate)
pr_data$Temperature<-as.factor(pr_data$Temperature)
```

Change level of par factor. 
```{r}
pr_data$PAR<-factor(pr_data$PAR, levels=c("0.1", "50", "100", "500", "0.2"))
```

Look for outliers in the data.  

```{r}
boxplot(pr_data$P.nmol.org.min)
pr_data<-pr_data%>%filter(P.nmol.org.min < 0.10)
boxplot(pr_data$P.nmol.org.min)

boxplot(pr_data$P.nmol.org.min~pr_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/F1F2_sdr/runs_temp.csv")
pr.temps = subset(pr.temps, select = -c(X) ) #remove empty column

pr_data$Plate<-as.integer(pr_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pr_data<-left_join(pr_data,pr.temps)

#round to 0.1°C 
pr_data<-pr_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

# Plot photosynthesis over PAR and temperature treatments  

Plot data with means   
```{r}
pr_plot1<-pr_data %>%
  filter(P.nmol.org.min<0.1)%>%
    group_by(PAR, Group, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Group))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Group), colour=Group), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Group)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1

ggsave("figures/F1F2_sdr/pr_means.png", pr_plot1, dpi=300, w=7, h=5, units="in")
```

Display by individual plate (duplicate plates at each temperature).  
```{r}
pr_plot1a<-pr_data %>%
  filter(P.nmol.org.min<0.1)%>%
    group_by(PAR, Group, Plate)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Group))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Group), colour=Group), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Group)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1a
```

View by plate with a loess line. 
```{r}
pr_plot2b<-pr_data %>%
  filter(P.nmol.org.min<0.1)%>%
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=Group)) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Group, group=Group), method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot2b
```
Plates look similar and consistent. 

Plot data with individual data points.    
```{r}
pr_plot3<-pr_data %>%
  filter(P.nmol.org.min<0.1)%>%
    
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=interaction(PAR, Group))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot3

ggsave("figures/F1F2_sdr/pr_dots.png", pr_plot3, dpi=300, w=7, h=5, units="in")
```

## Conduct anova 

Run anova on effect of PAR, group, and temperature. 
```{r}
summary(aov(P.nmol.org.min~PAR*Group*Temperature, data=pr_data))
```
There are strong effects of temperature and light. 

Run model for 27 degrees. 
```{r}
model<-pr_data%>%
  filter(Temperature==27)%>%
  aov(P.nmol.org.min~PAR*Group, data=.)

summary(model)
```
Strong effect of light at 27C. 

Run model for 33
```{r}
model<-pr_data%>%
  filter(Temperature==33)%>%
  aov(P.nmol.org.min~PAR*Group, data=.)

summary(model)
```
Strong effect of light at 33C. 

# Calculate R, P, and P:R 

Calculate gross and net photosynthesis, light enhanced respiration (LER), dark respiration (Rd), and P:R ratio at low (50 PAR), medium (100 PAR), and high (500 PAR) light. 
```{r}
pr_data_calc<-pr_data%>%
  select(Date, Chamber.ID, Plate, Run, Temperature, Group, PAR, P.nmol.org.min)%>%
  pivot_wider(names_from=PAR, values_from=P.nmol.org.min)%>%
  mutate(ler=`0.1`)%>%
  mutate(rd=`0.2`)%>%
  mutate(ler_inverse=ler*-1)%>%
  mutate(rd_inverse=rd*-1)%>%
  mutate(p_net_low=`50`)%>%
  mutate(p_net_medium=`100`)%>%
  mutate(p_net_high=`500`)%>%
  mutate(p_gross_low=p_net_low-ler)%>%
  mutate(p_gross_medium=p_net_medium-ler)%>%
  mutate(p_gross_high=p_net_high-ler)%>%
  mutate(ratio_low=p_gross_low/(ler_inverse))%>%
  mutate(ratio_medium=p_gross_medium/(ler_inverse))%>%
  mutate(ratio_high=p_gross_high/(ler_inverse))%>%
  pivot_longer(names_to="Metric", values_to="Value", cols=c(ler:ratio_high))%>%
  select(!c(`0.1`, `100`, `500`, `0.2`))

boxplot(pr_data_calc$Value)

pr_data_calc<-pr_data_calc%>%
  filter(Value>-5)%>% #%>%
  filter(Value<10)%>%
  filter(!c(Metric=="p_gross_low" & Value >0.1))%>%
filter(!c(Metric=="p_net_low" & Value >0.1))

boxplot(pr_data_calc$Value)
```

Set level of factors. 
```{r}
levels(as.factor(pr_data_calc$Metric))

pr_data_calc$Metric<-factor(pr_data_calc$Metric, levels=c("p_gross_low", "p_gross_medium", "p_gross_high", "p_net_low", "p_net_medium", "p_net_high", "ratio_low", "ratio_medium", "ratio_high", "rd", "rd_inverse", "ler", "ler_inverse"))
```

View metrics. 
```{r}
metric_plot1<-pr_data_calc%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_medium")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
    filter(!Metric=="rd")%>%
    filter(!Metric=="rd_inverse")%>%
    filter(!Metric=="ler_inverse")%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="lm", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1

ggsave("figures/F1F2_sdr/pr_metrics.png", metric_plot1, dpi=300, w=7, h=5, units="in")
```

View metrics across temperatures. 
```{r}
metric_plot2<-pr_data_calc%>%
    filter(Metric %in% c("ratio_high", "ratio_low", "ratio_medium"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="lm", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2

ggsave("figures/F1F2_sdr/pr_ratio.png", metric_plot2, dpi=300, w=7, h=5, units="in")
```

View metrics. 
```{r}
metric_plot3<-pr_data_calc%>%
    filter(Metric %in% c("ler_inverse", "rd_inverse"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Group), color=Group), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Group, group=Group), method="lm", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Group", values=c("brown4", "darkblue", "orange", "lightblue"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot3

ggsave("figures/F1F2_sdr/ler_rd.png", metric_plot3, dpi=300, w=7, h=5, units="in")
```

## Run anova on differences. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ratio_high")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)
```
No difference in P:R ratio at high (500 PAR) light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ratio_medium")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)
```
No difference in P:R ratio in medium (100 PAR) light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ratio_low")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Group")
```
Significant effect of group (p=0.045) on P:R at low (50 PAR) light - F2 NB has lower P:R ratios than WT F2.  

```{r}
model<-pr_data_calc%>%
  filter(Metric=="ler")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Light enhanced respiration is lower at 33 compared to 27. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="rd")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")

TukeyHSD(model, "Group")
```
Dark respiration is different by group and by temperature. Rd is higher at 33 degrees. Rd is also lower in the F2 NB than the F1 NB. The F2 NB also has lower Rd than the F2 WT. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_gross_high")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)
```
No effect on gross photosynthesis at high light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_gross_medium")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Gross photosynthesis at medium light (100) PAR is significantly affected by temperature. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_gross_low")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Gross photosynthesis at low light is affected by temperature. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_net_high")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
Net photosynthesis at high light is affected by temperature. P net increases at higher temperatures at high light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_net_medium")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)
```
No effect on P net at medium light. 

```{r}
model<-pr_data_calc%>%
  filter(Metric=="p_net_low")%>%
  aov(Value~Group*Temperature, data=.)

summary(model)
```
No effect on P net at low light. 

Test effect of light on P:R ratio. 
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("ratio_low", "ratio_medium", "ratio_high"))%>%
  aov(Value~Metric*Temperature*Group, data=.)

summary(model)

TukeyHSD(model, "Metric")
```
P:R ratios are significantly different by light. Ratios at high light are different than low and medium light. 

Test effect of light on P gross.  
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("p_gross_low", "p_gross_medium", "p_gross_high"))%>%
  aov(Value~Metric*Temperature*Group, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Group | Temperature)
pairs(emm)
```
P gross increases with light and is different between all light levels. P gross is also lower at high temperature. There is also an interaction betwen temperature and group. At 27 degrees, P gross is higher in F1 WT than F2 WT.  

Test effect of light on P net.   
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("p_net_low", "p_net_medium", "p_net_high"))%>%
  aov(Value~Metric*Temperature*Group, data=.)

summary(model)

TukeyHSD(model, "Metric")
```
P net is affected by light and increases with each light level. 
