---
title: "Fungia Respirometry plotting and analysis - Hawaii 2023"
author: "jillashey"
date: "2023-07-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script plots and analyzes the initial respirataion Fungia larval run. These larvae are aposymbiotic and acquire their symbionts from the environment, which is why photosynthesis was not run. I assessed how many individuals would need to be loaded into each well to get a good signal. I loaded approximately 8, 12, 16 or 20 larvae per well. 

## Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("nlme" %in% rownames(installed.packages()) == 'FALSE') install.packages('nlme')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('nlme')
```

## Data visualization + manipulation

Load data from LoLinR.

```{r}
r_data <- read.csv("output/fungia/calculated_normalized_rates.csv")
sample_info <- read.csv("data/fungia/PR_Fungia_Sample_Info.csv")
```

Format dataset.

```{r}
#remove all rows of wells that did not have samples or blanks
r_data<-r_data[!is.na(r_data$Type),]

#format columns
r_data$Temperature<-as.factor(r_data$Temperature)
r_data$Tank<-as.factor(r_data$Tank)
r_data$SDR<-as.factor(r_data$SDR)
r_data$Plate<-as.factor(r_data$Plate)
r_data$Org.Number <- as.factor(r_data$Org.Number)
```

If a respiration value is above 0 (indicating oxygen production), turn it into 0. 
```{r}
r_data<-r_data%>%
  mutate(R.nmol.org.min=if_else(R.nmol.org.min>0, 0, R.nmol.org.min))

boxplot(r_data$R.nmol.org.min)
```

Calculate the inverse. 
```{r}
r_data<-r_data%>%
  mutate(R.nmol.org.min=R.nmol.org.min*-1)

boxplot(r_data$R.nmol.org.min)
```

Look for outliers in the data.

```{r}
r_data<-r_data%>%filter(R.nmol.org.min < 0.006)
boxplot(r_data$R.nmol.org.min)
```
Removed one outlier 

## Plot respiration over organism number 

```{r}
r_plot1 <- r_data %>%
  group_by(Org.Number) %>%
  dplyr::summarise(mean=mean(R.nmol.org.min, na.rm=TRUE), sd=sd(R.nmol.org.min, na.rm=TRUE), N=length(R.nmol.org.min), se=sd/sqrt(N)) %>%
  
  ggplot(., aes(x = Org.Number, y = mean)) +
    #facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("No. organisms") + 
    #scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot1
```

Respiration peaks with 12 larave and then decreases with more than 12. This could be due to oxygen depletion in the well with >12 larave. 

## Look for outliers 

View and remove outliers.

```{r}
#identify outliers by Temperature and Treatment groups
outlier.plot <- ggbetweenstats(r_data, Org.Number, R.nmol.org.min, outlier.tagging = TRUE)
outlier.plot

ggsave("figures/fungia/outliersbygroup_pr.png", outlier.plot, dpi=300, w=16, h=8, units="in")

#set quantile values
q <- c(0.1, 0.9)

# calculate quantile values by Temperature and Treatment groups
Quants <- r_data %>%
  group_by(Org.Number) %>%
  summarize(quant25 = quantile(R.nmol.org.min, probs = q[1], na.rm=TRUE),
            quant75 = quantile(R.nmol.org.min, probs = q[2], na.rm=TRUE),
            IQRbyGroup=IQR(R.nmol.org.min, na.rm=TRUE))

# Add group name 
## only do this if I do TPCs

#Calculate Quantile upper and lower ranges 
Quants$upper <-  Quants$quant75+1.5*Quants$IQRbyGroup # Upper Range  
Quants$lower <- Quants$quant25-1.5*Quants$IQRbyGroup # Lower Range

#join outlier cutoffs with rate data
r_data <- left_join(r_data, Quants)

#remove outliers from rates
r_data_outlier <- r_data %>%
  filter(R.nmol.org.min < upper) %>%
  filter(R.nmol.org.min > lower) #%>%
  #filter(rate < 0.125)

# Visualize data following outlier removal
outlier.plot2 <- ggbetweenstats(r_data_outlier, Org.Number, R.nmol.org.min, outlier.tagging = TRUE)
outlier.plot2

ggsave("figures/fungia/removed_outliersbygroup_pr.png", outlier.plot2, dpi=300, w=16, h=8, units="in")
```

I think the outlier removal worked! It removed one outlier from the 8 org.number group. 

## Plot data after outlier removal 

```{r}
r_plot2 <- r_data_outlier %>%
  group_by(Org.Number) %>%
  dplyr::summarise(mean=mean(R.nmol.org.min, na.rm=TRUE), sd=sd(R.nmol.org.min, na.rm=TRUE), N=length(R.nmol.org.min), se=sd/sqrt(N)) %>%

  ggplot(., aes(x = Org.Number, y = mean)) +
    #facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("No. organisms") + 
    #scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot2

ggsave("figures/fungia/pr_means_orgnumber.png", r_plot2, dpi=300, w=7, h=5, units="in")
```

## Running anova on respiration rates by organism number

I think that 8 v. 20 organisms will differ significantly by one another, but no other differences.

```{r}
aov <- r_data_outlier %>%
  aov(R.nmol.org.min~Org.Number, data=.)
summary(aov)

# No effect of org number! 

TukeyHSD(aov)
```

Since there is no effect of organism number on the respiration rate, I think that we should load 8-12 fungia per well. This makes it relatively easier to count out larvae and load the wells for respiration. 

AH recommends using 10 larvae per well. This will make counting easier and will prevent oxygen depletion that likely drives reduced respiration rates at >12 larvae. It may not be significant here, but there is definitely a visible trend we should be mindful of. 


