---
title: "Larval PI Curve Plotting and Analysis - Hawaii 2023" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
```

# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pi_data<-read.csv("output/larval_pi_curves/calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pi_data<-pi_data[!is.na(pi_data$Type),]

#format columns
pi_data$PAR<-as.factor(pi_data$PAR)
pi_data$Tank<-as.factor(pi_data$Tank)
pi_data$Symbiont<-as.factor(pi_data$Symbiont)
pi_data$SDR<-as.factor(pi_data$SDR)
pi_data$Run<-as.factor(pi_data$Run)
```

Look for outliers in the data.  

```{r}
boxplot(pi_data$P.nmol.org.min)
#pi_data<-pi_data%>%filter(P.nmol.org.min > -0.10)
#boxplot(pi_data$P.nmol.org.min)
```

Calculate mean temperature values for each run.    

```{r}
pi.temps<-read.csv("output/larval_pi_curves/runs_temp.csv")
pi.temps = subset(pi.temps, select = -c(X) ) #remove empty column

#format "run" column
pi.temps<-pi.temps %>% #Format as "#" rather than "Run#"
  mutate(Run = str_sub(Run, 4, -1))
pi.temps$Run<-as.integer(pi.temps$Run) #format as integer
pi.temps$Run<-as.factor(pi.temps$Run) #format as factor
```

Add temperature data to master data frame.  
```{r}
pi_data<-left_join(pi_data,pi.temps)

#round to 0.1°C 
pi_data<-pi_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

# Plot photosynthesis over PAR treatments  

Plot data with means   
```{r}
pi_plot1<-pi_data %>%
    group_by(PAR, Symbiont)%>%
    dplyr::summarise(mean=mean(abs(P.nmol.org.min), na.rm=TRUE), sd=sd(abs(P.nmol.org.min), na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = abs(mean), group=interaction(PAR, Symbiont))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot1

ggsave("figures/larval_pi_curves/pi_means.png", pi_plot1, dpi=300, w=7, h=5, units="in")
```

Plot data with a loess line   
```{r}
pi_plot2<-pi_data %>%
    
    ggplot(., aes(x = as.factor(PAR), y = abs(P.nmol.org.min), group=Symbiont)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=TRUE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot2

ggsave("figures/larval_pi_curves/pi_loess.png", pi_plot2, dpi=300, w=7, h=5, units="in")
```

# Apply PI curve calculations and models 

## Look for outliers 

View and remove outliers.   
```{r, warning=FALSE, message=FALSE}
#identify outliers by Temperature and Treatment groups
outlier.plot <- ggbetweenstats(pi_data, PAR, P.nmol.org.min, outlier.tagging = TRUE)
outlier.plot

ggsave("figures/larval_pi_curves/outliersbygroup.png", outlier.plot, dpi=300, w=16, h=8, units="in")

#set quantile values
q <- c(0.25, 0.75)

# calculate quantile values by PAR and Treatment groups
Quants <- pi_data %>%
  group_by(PAR, Symbiont)%>%
  summarise(quant25 = quantile(P.nmol.org.min, probs = q[1], na.rm=TRUE),
            quant75 = quantile(P.nmol.org.min, probs = q[2], na.rm=TRUE),
            IQRbyGroup=IQR(P.nmol.org.min, na.rm=TRUE))

# NOT WORKING 






# add a group name by pasting PAR and Treatment
Quants$group <-paste0(Quants$PAR,"_", Quants$Symbiont)

#Calculate Quantile upper and lower ranges 
Quants$upper <-  Quants$quant75+1.5*Quants$IQRbyGroup # Upper Range  
Quants$lower <- Quants$quant25-1.5*Quants$IQRbyGroup # Lower Range

#join outlier cutoffs with rate data
pi_data <- left_join(pi_data, Quants)







#remove outliers from rates
tpc_data <- tpc_data %>%
  filter(R.nmol.org.min < upper) %>%
  filter(R.nmol.org.min > lower) #%>%
  #filter(rate < 0.125)

# Visualize data following outlier removal
outlier.plot2 <- ggbetweenstats(tpc_data, Temp.C, R.nmol.org.min, outlier.tagging = TRUE)
outlier.plot2

ggsave("figures/tpc_sdr/removed_outliersbygroup.png", outlier.plot2, dpi=300, w=16, h=8, units="in")
```

## Fitting TPC    

TPC fitting 
Padifeld et al **rTPC and nls.multstart: A new pipeline to fit thermal performance curves in r**  
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13585  

Sharpe Schoolfield 1981 model
Schoolfield, R. M., Sharpe, P. J. H., & Magnuson, C. E. (1981). Non-linear regression of biological temperature-dependent rate models based on absolute reaction-rate theory. Journal of theoretical biology, 88(4), 719-731. https://doi.org/10.1016/0022-5193(81)90246-0

```{r}
# choose model
get_model_names()
#sharpeschoolhigh_1981

# get start vals
start_vals <- get_start_vals(tpc_data$Temp.C,tpc_data$R.nmol.org.min, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc_data$Temp.C,tpc_data$R.nmol.org.min, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc_data$Temp.C,tpc_data$R.nmol.org.min, model_name = 'sharpeschoolhigh_1981')

#view values
start_vals
low_lims
upper_lims
```

### Cladocopium TPC 

Run a model fit for Cladocopium. 
```{r}
# CURVE FIT 
C_data<-tpc_data%>%
  filter(Symbiont=="Cladocopium")

#fit 
C.fit <- nls_multstart(R.nmol.org.min~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 28),
                                                     data = C_data,
                                                     iter = 500,
                                                     start_lower = start_vals - 1,
                                                     start_upper = start_vals + 1,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

C.fit

#generate the predicted data
C.new_data <- data.frame(temp = seq(from=min(C_data$Temp.C), to=max(C_data$Temp.C), by=((max(C_data$Temp.C)-min(C_data$Temp.C))/(nrow(C_data)-1)))) #set min and max temperatures and add prediction for number of rows in data frame 

#C.new_data <- data.frame(temp = seq(from=min(C_data$Temp.C), to=max(C_data$Temp.C), by=0.071)) #manual interval if required

C.preds <- augment(C.fit, newdata = C.new_data)

#calculate TPC parameters
C.TCP.res <- calc_params(C.fit) %>%
  mutate_all(round, 2)   # round 

C.TCP.res 

```

Bootstrap the curve fit by resampling the original data with replacement. 
```{r}
# refit model using nlsLM
C.fit_nlsLM <- minpack.lm::nlsLM(R.nmol.org.min~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 28),
                        data = C_data,
                        start = coef(C.fit),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(C_data)))

# bootstrap using case resampling
C.boot1 <- Boot(C.fit_nlsLM, method = 'case')

# look at the data
head(C.boot1$t)


# create predictions of each bootstrapped model
C.boot1_preds <- C.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(C_data$Temp.C), max(C_data$Temp.C), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
C.boot1_conf_preds <- group_by(C.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
C.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), C.preds, col = 'orange') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), C.boot1_conf_preds, fill = 'orange', alpha = 0.3) +
  geom_point(aes(Temp.C, R.nmol.org.min), C_data, size = 2, alpha = 0.5,col = 'orange') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Respiration Rate (nmol O2/larva/min');C.CI.plot 

```

### Durusdinium TPC 

Run a model fit for Durusdinium 
```{r}
# CURVE FIT 
D_data<-tpc_data%>%
  filter(Symbiont=="Durusdinium")

#fit 
D.fit <- nls_multstart(R.nmol.org.min~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 28),
                                                     data = D_data,
                                                     iter = 500,
                                                     start_lower = start_vals - 1,
                                                     start_upper = start_vals + 1,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

D.fit

#generate the predicted data
D.new_data <- data.frame(temp = seq(from=min(D_data$Temp.C), to=max(D_data$Temp.C), by=((max(D_data$Temp.C)-min(D_data$Temp.C))/(nrow(D_data)-1)))) #set min and max temperatures and add prediction for number of rows in data frame 

#C.new_data <- data.frame(temp = seq(from=min(C_data$Temp.C), to=max(C_data$Temp.C), by=0.071)) #manual interval if required

D.preds <- augment(D.fit, newdata = D.new_data)

#calculate TPC parameters
D.TCP.res <- calc_params(D.fit) %>%
  mutate_all(round, 2)   # round 

D.TCP.res 

```

Bootstrap the curve fit by resampling the original data with replacement. 
```{r}
# refit model using nlsLM
D.fit_nlsLM <- minpack.lm::nlsLM(R.nmol.org.min~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 28),
                        data = D_data,
                        start = coef(D.fit),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(D_data)))

# bootstrap using case resampling
D.boot1 <- Boot(D.fit_nlsLM, method = 'case')

# look at the data
head(D.boot1$t)


# create predictions of each bootstrapped model
D.boot1_preds <- D.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(D_data$Temp.C), max(D_data$Temp.C), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
D.boot1_conf_preds <- group_by(D.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
D.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), D.preds, col = 'brown4') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), D.boot1_conf_preds, fill = 'brown4', alpha = 0.3) +
  geom_point(aes(Temp.C, R.nmol.org.min), D_data, size = 2, alpha = 0.5,col = 'brown4') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Respiration Rate (nmol O2/larva/min');D.CI.plot 

```

## Plot TPC parameters

```{r}
 #set plot colors
cols <- c("Cladocopium"="orange", "Durusdinium"="brown4")
  
# plot data and model fit
TPC.plot <- ggplot(data=tpc_data, aes(x=Temp.C)) +
   geom_point(aes(Temp.C, R.nmol.org.min, color="Cladocopium"), C_data, size = 2, alpha = 0.5) +
   geom_point(aes(Temp.C, R.nmol.org.min, color="Durusdinium"), D_data, size = 2, alpha = 0.5) +
   geom_line(aes(temp, .fitted), C.preds, col = 'orange', size=2) +
   geom_line(aes(temp, .fitted), D.preds, col = "brown4", size=2) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), C.boot1_conf_preds, fill = "orange", alpha = 0.3) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), D.boot1_conf_preds, fill = 'brown4', alpha = 0.3) +
   #xlim(21,33)+
   #scale_x_continuous(breaks=c(22,24,26,28,30,32,34,36))+
   theme_bw(base_size = 12) +
   scale_colour_manual(name="Symbiont",values=cols)+
   theme(legend.position = "right",
         panel.border = element_blank(), panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
   labs(x = 'Temperature (ºC)',
        y = expression("Respiration Rate"~nmol~O[2] ~larva^{-1}~min^{-1}));TPC.plot 

ggsave("figures/tpc_sdr/TPC_SharpSchool.png", TPC.plot, dpi=300, w=8, h=8, units="in")

```

## Confidence intervals of TPC parameters   

```{r}
broom::tidy(C.fit_nlsLM)
broom::tidy(D.fit_nlsLM)

```

Cladocopium parameters
```{r}
#Cladocopium
#calculate all the TPC parameters
C.extra_params <- calc_params(C.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
C.ci_extra_params <- Boot(C.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(C.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
C.ci_extra_params <- left_join(C.ci_extra_params, C.extra_params)
C.ci_extra_params$Symbiont <- "Cladocopium"
```

Durusdinium parameters 
```{r}
#Durusdinium
#calculate all the TPC parameters
D.extra_params <- calc_params(D.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
D.ci_extra_params <- Boot(D.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(D.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')
  
#join the parameters and CIs  
D.ci_extra_params <- left_join(D.ci_extra_params, D.extra_params)
D.ci_extra_params$Symbiont <- "Durusdinium"

#Join Ambient and High estimates and CIs
All_params <- rbind(C.ci_extra_params, D.ci_extra_params)
All_params <- All_params %>% 
 mutate_if(is.numeric, round, 2)

#Plot all of the estimates
estimate.plots <- ggplot(All_params, aes(Symbiont, estimate, color=Symbiont)) +
  geom_point(size = 2) +
  scale_color_manual(name="Symbiont", values=c("orange","brown4"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('');estimate.plots

#filter to only the most relavent and well characterized parameters
All_params <- All_params %>% 
  filter(!param=="ctmin") %>%
  filter(!param=="ctmax") %>%
  filter(!param=="skewness") %>%
  filter(!param=="thermal_safety_margin") %>%
  filter(!param=="thermal_tolerance") %>%
  filter(!param=="q10")%>%
  filter(!param=="breadth")
  
#view estimate plots
estimate.plots <- ggplot(All_params, aes(Symbiont, estimate, color=Symbiont)) +
  geom_point(size = 2) +
  scale_color_manual(name="Synmbiont", values=c("orange","brown4"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  labs(y = NULL)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        legend.position = "none",
        strip.background = element_blank(), 
        strip.placement = "outside") +
  facet_wrap(~param, scales = 'free_y', nrow=1, 
             labeller = as_labeller(c(e = "e (Energy)", eh = " eh (Energy)", rmax= "Rmax (~nmol~O[2] ~larva^{-1}~min^{-1})",topt="Topt (Temperature °C)")), strip.position = "left") +
  scale_x_discrete('');estimate.plots

ggsave("figures/tpc_sdr/TPC_estimates_SharpSchool.png", estimate.plots, dpi=300, w=6, h=2, units="in")
``` 


```{r}
#generate a combined figure of TPCs and estimate plots
figure <- ggarrange(TPC.plot, estimate.plots,
                    labels = c("A", "B"),
                    ncol = 1, nrow = 2,
                    heights=c(1,0.5)); figure

ggsave("figures/tpc_sdr/TPC_combined_plot.png", figure, dpi=300, w=6, h=8, units="in")
ggsave("figures/tpc_sdr/TPC_combined_plot.pdf", figure, dpi=300, w=6, h=8, units="in")

``` 








