---
title: "Larval PI Curve Plotting and Analysis - Hawaii 2023" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Note: Will need to add filter by date once we have multiple days represented in the dataset 

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
```

# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pi_data<-read.csv("output/larval_pi_curves/calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pi_data<-pi_data[!is.na(pi_data$Type),]

#format columns
pi_data$PAR<-as.factor(pi_data$PAR)
pi_data$Tank<-as.factor(pi_data$Tank)
pi_data$Symbiont<-as.factor(pi_data$Symbiont)
pi_data$SDR<-as.factor(pi_data$SDR)
pi_data$Run<-as.factor(pi_data$Run)
```

Look for outliers in the data.  

```{r}
boxplot(pi_data$P.nmol.org.min)
#pi_data<-pi_data%>%filter(P.nmol.org.min > -0.10)
#boxplot(pi_data$P.nmol.org.min)

boxplot(pi_data$P.nmol.org.min~pi_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pi.temps<-read.csv("output/larval_pi_curves/runs_temp.csv")
pi.temps = subset(pi.temps, select = -c(X) ) #remove empty column

#format "run" column
pi.temps<-pi.temps %>% #Format as "#" rather than "Run#"
  mutate(Run = str_sub(Run, 4, -1))
pi.temps$Run<-as.integer(pi.temps$Run) #format as integer
pi.temps$Run<-as.factor(pi.temps$Run) #format as factor
```

Add temperature data to master data frame.  
```{r}
pi_data<-left_join(pi_data,pi.temps)

#round to 0.1°C 
pi_data<-pi_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

# Plot photosynthesis over PAR treatments  

Plot data with means   
```{r}
pi_plot1<-pi_data %>%
    group_by(PAR, Symbiont)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot1

ggsave("figures/larval_pi_curves/pi_means.png", pi_plot1, dpi=300, w=7, h=5, units="in")
```

Plot data with a loess line   
```{r}
pi_plot2<-pi_data %>%
    
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=Symbiont)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=TRUE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot2

ggsave("figures/larval_pi_curves/pi_loess.png", pi_plot2, dpi=300, w=7, h=5, units="in")
```

# Apply PI curve calculations and models 

## Look for outliers 

View and remove outliers.   
```{r, warning=FALSE, message=FALSE}
#identify outliers by Temperature and Treatment groups
outlier.plot <- ggbetweenstats(pi_data, PAR, P.nmol.org.min, outlier.tagging = TRUE)
outlier.plot

ggsave("figures/larval_pi_curves/outliersbygroup.png", outlier.plot, dpi=300, w=8, h=8, units="in")

#set quantile values
q <- c(0.25, 0.75)

# calculate quantile values by PAR and Treatment groups
Quants <- pi_data %>%
  dplyr::group_by(PAR, Symbiont)%>%
  dplyr::summarise(quant25 = quantile(P.nmol.org.min, probs = q[1], na.rm=TRUE),
            quant75 = quantile(P.nmol.org.min, probs = q[2], na.rm=TRUE),
            IQRbyGroup=IQR(P.nmol.org.min, na.rm=TRUE))


# add a group name by pasting PAR and Treatment
Quants$group <-paste0(Quants$PAR,"_", Quants$Symbiont)

#Calculate Quantile upper and lower ranges 
Quants$upper <-  Quants$quant75+1.5*Quants$IQRbyGroup # Upper Range  
Quants$lower <- Quants$quant25-1.5*Quants$IQRbyGroup # Lower Range

#join outlier cutoffs with rate data
pi_data <- left_join(pi_data, Quants)

#remove outliers from rates
pi_data <- pi_data %>%
  filter(P.nmol.org.min < upper) %>%
  filter(P.nmol.org.min > lower) #%>%
  #filter(rate < 0.125)

# Visualize data following outlier removal
outlier.plot2 <- ggbetweenstats(pi_data, PAR, P.nmol.org.min, outlier.tagging = TRUE)
outlier.plot2

ggsave("figures/larval_pi_curves/removed_outliersbygroup.png", outlier.plot2, dpi=300, w=8, h=8, units="in")
```

## Fitting PI Curves 

Scripts for this section based on E5 PI curve analysis in the time series repository. 

Define data 
```{r}
#specify data
pi_data$PAR <- as.numeric(pi_data$PAR)
pi_data$Pc <- as.numeric(pi_data$P.nmol.org.min)
```

Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit

Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (Am from Bayesian script)  
alpha = quantum yeild (AQY from Bayesian script)  
I/E = irradiance (PAR from Bayesian script)  
Rd = dark respiration   

Next run NLS models.

Using fixed initial values (not using, but keeping for reference):  


Not working with current data, keep scripts to try with real data 

```{r}
nest <- nest_legacy
unnest <- unnest_legacy

nls_data <- pi_data %>% 
   mutate(new_Pc=Pc+1)%>%
   group_by(Tank) %>%
   nest(-Tank) %>%
   mutate(model1 = map(data, ~ 
                        nls(new_Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.05,  AQY=0.001, Rd=-0.02)) %>%
                             tidy %>%
                             dplyr::select(term, estimate) %>% 
                             spread(term, estimate))) %>%
  unnest(model1) %>%
  mutate(Ik = Am/AQY) 

#save

```

Using flexible initial values based on input data:  

```{r}
nest <- nest_legacy
unnest <- unnest_legacy

pi_data<-pi_data%>%
  mutate(new_Pc=Pc+1)

nls_data <- pi_data %>% 
   group_by(Symbiont) %>%
   nest(-Symbiont) %>%
   mutate(model1 = map(data, ~ 
                         nls(new_Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=(max(.$new_Pc)-min(.$new_Pc)),  AQY=0.001, Rd=-min(.$new_Pc)), trace=TRUE) %>%
                              tidy %>%
                              dplyr::select(term, estimate) %>% 
                              spread(term, estimate))) %>%
  unnest(model1) %>%
  mutate(Ik = Am/AQY) 

#save 

```

ERROR ABOVE THIS POINT NEED TO OPTIMIZE 


Plot curve over data points.  
```{r}
augmented <- pi_data %>% 
  nest(-Symbiont) %>% 
  mutate(
    fit = map(data, ~ nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)


#all tanks together
augmented %>%
  group_by(Symbiont)%>%
  qplot(PAR, Pc, data = ., geom = 'point', colour = Symbiont) +
  geom_line(aes(y=.fitted))+
  theme(legend.position="none")


#view individual plots
by(augmented,augmented$colony_id, function(i) {
  ggplot(i) +
          geom_point(aes(PAR, Pc, group=Symbiont)) + 
          geom_line(aes(y=.fitted, x=PAR)) + 
          theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste('Photosynthetic rate (nmol larva'^-1, 'min'^-1,")")),
               title = paste0("PIcurve_", augmented$Symbiont))
})

```


## Further plotting 

ADD HERE AS THE DATA BECOMES AVAILABLE 











