---
title: "Larval PI Curve Plotting and Analysis - Hawaii 2023" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
```

# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pr_data<-read.csv("output/pr_sdr/PR_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pr_data<-pr_data[!is.na(pr_data$Type),]

#format columns
pr_data$Symbiont<-as.factor(pr_data$Symbiont)
pr_data$SDR<-as.factor(pr_data$SDR)
pr_data$Plate<-as.factor(pr_data$Plate)
pr_data$Temperature<-as.factor(pr_data$Temperature)
```

Look for outliers in the data.  

```{r}
boxplot(pr_data$nmol.org.min)

boxplot(pr_data$nmol.org.min~pr_data$Light_Value)

pr_data$code<-paste(pr_data$Temperature, "_",pr_data$Light_Value)

boxplot(pr_data$nmol.org.min~pr_data$code)
```

Set values at 0 for light runs if <0 indicating no photosynthesis occurred. Set values at 0 for dark runs if >0 indicating no respiration occurred. 
```{r}
pr_data<-pr_data%>%
  mutate(nmol.org.min=if_else(nmol.org.min<0 & Light_Value==100, 0, 
                                if_else(nmol.org.min<0 & Light_Value==500, 0, 
                                        if_else(nmol.org.min>0 & Light_Value==0, 0, nmol.org.min))))
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/pr_sdr/mean_temperature_by_plate_and_light_interval.csv")

pr_data$Plate<-as.integer(pr_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pr.temps$SDR<-as.factor(as.character(pr.temps$SDR))
pr.temps$Plate <- as.integer(sub("(?i)^plate", "", pr.temps$Plate, perl = TRUE))  # removes leading "Plate"

pr_data<-left_join(pr_data,pr.temps)

#round to 0.1°C 
pr_data<-pr_data%>%
  mutate(Temp.C=round(Mean_T_internal_C,1))
```

Set levels for light. 
```{r}
pr_data$Light_Value<-factor(pr_data$Light_Value, levels=c("100", "500", "0"))
```

Filter out Plate 2 (values were off and blanks did not make sense). 

```{r}
pr_data<-pr_data%>%
  filter(!Plate==2)
```

# Plot and analyze: 20230624 

Plot data with means   
```{r}
pr_plot1<-pr_data %>%
    filter(Date=="20230624")%>%
    group_by(Light_Value, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(nmol.org.min, na.rm=TRUE), sd=sd(nmol.org.min, na.rm=TRUE), N=length(nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(Light_Value), y = mean, group=interaction(Light_Value, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Light_Value, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Light_Value, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1

ggsave("figures/pr_sdr/pr_means.png", pr_plot1, dpi=300, w=7, h=5, units="in")
```

Display with plate. 
```{r}
pr_plot1a<-pr_data %>%
    filter(Date=="20230624")%>%
    group_by(Light_Value, Symbiont, Plate)%>%
    dplyr::summarise(mean=mean(nmol.org.min, na.rm=TRUE), sd=sd(nmol.org.min, na.rm=TRUE), N=length(nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(Light_Value), y = mean, group=interaction(Light_Value, Symbiont))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Light_Value, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Light_Value, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1a
```

View by plate. 
```{r}
pr_plot2b<-pr_data %>%
    filter(Date=="20230624")%>%
    ggplot(., aes(x = as.factor(Light_Value), y = nmol.org.min, group=Symbiont)) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot2b
```

Plot data with dots   
```{r}
pr_plot3<-pr_data %>%
    filter(Date=="20230624")%>%
  
    ggplot(., aes(x = Temperature, y = nmol.org.min, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Light_Value)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Light_Value, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot3

ggsave("figures/pr_sdr/pr_dots.png", pr_plot3, dpi=300, w=7, h=5, units="in")
```

## Conduct anova 

Run anova on values at each temperature
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  aov(nmol.org.min~Light_Value*Symbiont*Temperature, data=.)

summary(model)

emm<-emmeans(model, ~ Temperature | Light_Value)
pairs(emm)

emm<-emmeans(model, ~ Symbiont | Light_Value)
pairs(emm)
```
Temperature differences present at 0, 100, and 100 PAR. 

Run model for dark 
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  filter(Light_Value=="0")%>%
  aov(nmol.org.min~Temperature*Symbiont, data=.)

summary(model)
```
Temp changes in dark. 

Run model for 100 PAR
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  filter(Light_Value=="100")%>%
  aov(nmol.org.min~Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~ Symbiont)
pairs(emm)
```
Change by temperature at 100 PAR, lower in NB compared to WT. 

Run model for 500 PAR
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  filter(Light_Value=="500")%>%
  aov(nmol.org.min~Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~ Temperature)
pairs(emm)
```
Peak at 30C at 500 PAR. 

# Calculate R, P, and P:R: 20230624 

Calculate gross photosynthesis and P:R ratio. 
```{r}
pr_data_calc<-pr_data%>%
  filter(Date=="20230624")%>%
  select(Date, Well, Plate, Run, Temperature, Symbiont, Light_Value, nmol.org.min)%>%
  pivot_wider(names_from=Light_Value, values_from=nmol.org.min)%>%
  mutate(ler=`0`)%>%
  mutate(ler_inverse=ler*-1)%>%
  mutate(p_net_low=`100`)%>%
  mutate(p_net_high=`500`)%>%
  mutate(p_gross_low=p_net_low-ler)%>%
  mutate(p_gross_high=p_net_high-ler)%>%
  mutate(ratio_low=p_gross_low/(ler_inverse))%>%
  mutate(ratio_high=p_gross_high/(ler_inverse))%>%
  pivot_longer(names_to="Metric", values_to="Value", cols=c(ler:ratio_high))%>%
  select(!c(`0`, `100`, `500`))
```

View P metrics. 
```{r}
metric_plot1<-pr_data_calc%>%
    filter(Date=="20230624")%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
     filter(!Metric=="ler_inverse")%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1

ggsave("figures/pr_sdr/pr_photometrics.png", metric_plot1, dpi=300, w=7, h=5, units="in")
```

View R metrics. 
```{r}
metric_plot1a<-pr_data_calc%>%
    filter(Date=="20230624")%>%
    filter(Metric %in% c("ler_inverse"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1a

ggsave("figures/pr_sdr/pr_resp.png", metric_plot1a, dpi=300, w=5, h=5, units="in")
```

View metrics for P:R ratio. 
```{r}
metric_plot2<-pr_data_calc%>%
   filter(Date=="20230624")%>%
    filter(Metric %in% c("ratio_high", "ratio_low"))%>%
  filter(Value>0)%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2

ggsave("figures/pr_sdr/pr_ratio.png", metric_plot2, dpi=300, w=7, h=5, units="in")
```

## Run anova on differences. 

Test P:R ratio at high light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ratio_high")%>%
  filter(Value>0)%>%
  #filter(!c(Value>4 & Symbiont=="C" & Temperature==33))%>% #remove outliers 
  #filter(!c(Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
  #filter(!Value<0)%>% #remove outliers 
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```

Test P:R ratios under low light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ratio_low")%>%
  filter(Value>0)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Symbiont")
```
Lower ratio in low light in NB. 

Test light enhanced respiration. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ler_inverse")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```

Test P gross at high light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_gross_high")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)
```

Test P gross at low light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_gross_low")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```

Test P net under low light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_net_low")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
TukeyHSD(model, "Symbiont")
```

```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_net_high")%>%
  aov(Value~Symbiont*Temperature, data=.)

summary(model)

```

## Test effect of light 

Test effect of light on P:R ratio. 
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("ratio_low", "ratio_high"))%>%
  filter(Date=="20230624")%>%
  filter(Value>0)%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Symbiont")
TukeyHSD(model, "Temperature")

```
Lower ratios in NB. 

Test effect of light on P gross.  
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("p_gross_low", "p_gross_high"))%>%
    filter(Date=="20230624")%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")
```

Test effect of light on P net.   
```{r}
model<-pr_data_calc%>%
  filter(Metric %in% c("p_net_low","p_net_high"))%>%
  filter(Date=="20230624")%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Temperature | Metric)
pairs(emm)
```


