---
title: "Larval PI Curve Plotting and Analysis - Hawaii 2023" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

NEED TO REMOVE/QC OUTLIERS FOR ALL PLOTS 

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
```


# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pr_data<-read.csv("output/pr_sdr/pr_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pr_data<-pr_data[!is.na(pr_data$Type),]

#format columns
pr_data$Tank<-as.factor(pr_data$Tank)
pr_data$Symbiont<-as.factor(pr_data$Symbiont)
pr_data$SDR<-as.factor(pr_data$SDR)
pr_data$Plate<-as.factor(pr_data$Plate)
pr_data$Temperature<-as.factor(pr_data$Temperature)
```

Remove plate that didnt work 

```{r}
#pr_data<-pr_data%>%
#  filter(!Plate==2)%>%
#  filter(!Plate==1)
```

Look for outliers in the data.  

```{r}
boxplot(pr_data$P.nmol.org.min)
pr_data<-pr_data%>%filter(P.nmol.org.min < 0.08)
boxplot(pr_data$P.nmol.org.min)

boxplot(pr_data$P.nmol.org.min~pr_data$PAR)

pr_data$code<-paste(pr_data$Temperature, "_",pr_data$PAR)

boxplot(pr_data$P.nmol.org.min~pr_data$code)
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/pr_sdr/runs_temp.csv")
pr.temps = subset(pr.temps, select = -c(X) ) #remove empty column

pr_data$Plate<-as.integer(pr_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pr_data<-left_join(pr_data,pr.temps)

#round to 0.1°C 
pr_data<-pr_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

# Plot and analyze: 20230624 

Plot data with means   
```{r}
pr_plot1<-pr_data %>%
    filter(Date=="20230624")%>%
    group_by(PAR, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1

ggsave("figures/pr_sdr/pr_means_0624.png", pr_plot1, dpi=300, w=7, h=5, units="in")
```

Display with plate. 
```{r}
pr_plot1a<-pr_data %>%
    filter(Date=="20230624")%>%
    group_by(PAR, Symbiont, Plate)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot1a
```

View by plate. 
```{r}
pr_plot2b<-pr_data %>%
    filter(Date=="20230624")%>%
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=Symbiont)) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot2b
```

Plot data with dots   
```{r}
pr_plot3<-pr_data %>%
    filter(Date=="20230624")%>%
  
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pr_plot3

ggsave("figures/pr_sdr/pr_dots_0624.png", pr_plot3, dpi=300, w=7, h=5, units="in")
```

## Conduct anova 

Run anova on values at each temperature
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  aov(P.nmol.org.min~PAR*Symbiont*Temperature, data=.)

summary(model)
```

Run model for 27 
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  filter(Temperature==27)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```

Run model for 30
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  filter(Temperature==30)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```

Run model for 33
```{r}
model<-pr_data%>%
  filter(Date=="20230624")%>%
  filter(Temperature==33)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```

# Calculate R, P, and P:R: 20230624 

Calculate gross photosynthesis and P:R ratio. 
```{r}
pr_data_calc<-pr_data%>%
  filter(Date=="20230624")%>%
  select(Date, Chamber.ID, Plate, Run, Temperature, Symbiont, PAR, P.nmol.org.min)%>%
  pivot_wider(names_from=PAR, values_from=P.nmol.org.min)%>%
  mutate(ler=`0`)%>%
  mutate(ler_inverse=ler*-1)%>%
  mutate(p_net_low=`100`)%>%
  mutate(p_net_high=`500`)%>%
  mutate(p_gross_low=p_net_low-ler)%>%
  mutate(p_gross_high=p_net_high-ler)%>%
  mutate(ratio_low=p_gross_low/(ler_inverse))%>%
  mutate(ratio_high=p_gross_high/(ler_inverse))%>%
  pivot_longer(names_to="Metric", values_to="Value", cols=c(ler:ratio_high))%>%
  select(!c(`0`, `100`, `500`))

boxplot(pr_data_calc$Value)

pr_data_calc<-pr_data_calc%>%
  filter(Value>-5)%>% #%>%
  filter(Value<9)

boxplot(pr_data_calc$Value)
```

View P metrics. 
```{r}
metric_plot1<-pr_data_calc%>%
    filter(Date=="20230624")%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
     filter(!Metric=="ler_inverse")%>%
    filter(!c(Value>0.05 & Metric=="p_net_low"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1

ggsave("figures/pr_sdr/pr_photometrics_0624.png", metric_plot1, dpi=300, w=7, h=5, units="in")
```

View R metrics. 
```{r}
metric_plot1a<-pr_data_calc%>%
    filter(Date=="20230624")%>%
    filter(Metric %in% c("ler_inverse"))%>%
    filter(!c(Value<0 & Metric=="ler_inverse"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1a

ggsave("figures/pr_sdr/pr_resp_0624.png", metric_plot1a, dpi=300, w=5, h=5, units="in")
```

View metrics for P:R ratio. 
```{r}
metric_plot2<-pr_data_calc%>%
   filter(Date=="20230624")%>%
    filter(Metric %in% c("ratio_high", "ratio_low"))%>%
  
    filter(!c(Metric=="ratio_high" & Value>4 & Symbiont=="C" & Temperature==33))%>% #remove outliers 
    filter(!c(Metric=="ratio_high" & Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
    filter(!c(Metric=="ratio_low" & Value< -1 & Symbiont=="WT" & Temperature==27))%>% #remove outliers 
    filter(!Value<0)%>% #remove outliers 
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2

ggsave("figures/pr_sdr/pr_ratio_0624.png", metric_plot2, dpi=300, w=7, h=5, units="in")
```

## Run anova on differences. 

Test P:R ratio at high light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ratio_high")%>%
  filter(!c(Value>4 & Symbiont=="C" & Temperature==33))%>% #remove outliers 
  filter(!c(Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
  filter(!Value<0)%>% #remove outliers 
  aov(Value~Temperature*Symbiont, data=.)

summary(model)
```
Almost significant effect of temp x symbiont. Strong effect of temperature with lower P:R ratio at high light and high temperature. 

Look at 30C under high light specifically. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ratio_high")%>%
  filter(Temperature==30)%>%
  filter(Value>-1)%>%
  aov(Value~Symbiont, data=.)

summary(model)

TukeyHSD(model, "Symbiont")
```
At high light under 30C, there is an effect of symbiont. C is lower than WT, but NB is not lower than WT. This suggests that C has lower P:R ratio at high temperature where as MIX is not as strongly affected. This matches our PI curve observation that MIX can handle higher light. 

Test P:R ratios under low light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ratio_low")%>%
  filter(Value>0)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Symbiont")
```
There is no effect of temperature on P:R ratios at low light. But there is an effect of symbiont. P:R ratios at low light are higher in WT than the MIX and are also lower in MIX than C across temperatures. This makes sense with what we saw in the PI curves with higher P in C at low light! 

Test light enhanced respiration. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="ler")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
There is an effect of temperature on LER. LER is lower at 27 than it is at 30 and 33. 

Test P gross at high light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_gross_high")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)
```
There is no effect of temp or symbiont. 

Test P gross at low light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_gross_low")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
There is an effect of temperature at low light. P net is higher as temperature increases. 

Test P net under low light. 
```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_net_low")%>%
  filter(!c(Value>0.05 & Metric=="p_net_low"))%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Temperature")
```
There is an effect of temperature on P net at low light. P net increases as temperature increases. 

```{r}
model<-pr_data_calc%>%
  filter(Date=="20230624")%>%
  filter(Metric=="p_net_high")%>%
  aov(Value~Symbiont*Temperature, data=.)

summary(model)

emm<-emmeans(model, ~ Symbiont | Temperature)
pairs(emm)
```
There is an effect of Symbiont x Temperature  on P net at high light. At 30C, P net in C is lower than WT. There is a trend for this in P gross as well above. This helps explain why P:R ratios are lower in C than WT. 

