---
title: "Lipidomics analysis"
author: "Ariana S Huffmyer"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(broom)
library(cowplot)
library(mixOmics)
library(dplyr)
library(conflicted)
library(vegan)
library(factoextra)
library(ggfortify)
library(RVAideMemoire)
library(ComplexHeatmap)
library(circlize)
library(tibble)
library(viridis)
library(genefilter)
library(UpSetR)

conflict_prefer("select", winner = "dplyr", quiet = FALSE)
conflict_prefer("filter", winner = "dplyr", quiet = FALSE)
conflict_prefer("rename", winner = "dplyr", quiet = FALSE)
```

# Data manipulation and preparations  

## Read in all files 

```{r}
metadata<-read_csv("data/lipids_metabolites/lipidomics/lipid_metadata_huffmyer_matthews.csv")%>%
  rename(sample=`Vial ID`, tube=`Sample ID`, temperature=Temperature, parent=Parent, type=Type, code=Code)

neg_values<-read_csv("data/lipids_metabolites/lipidomics/processed-data/negative/PeakValues_2_2024_09_19_13_01_19.csv")

neg_peaks<-read_csv("data/lipids_metabolites/lipidomics/processed-data/negative/PeakMaster_2_2024_09_19_13_01_19.csv")

pos_values<-read_csv("data/lipids_metabolites/lipidomics/processed-data/positive/PeakValues_3_2024_09_19_14_24_22.csv")

pos_peaks<-read_csv("data/lipids_metabolites/lipidomics/processed-data/positive/PeakMaster_3_2024_09_19_14_24_22.csv")

protein<-read_csv("data/lipids_metabolites/protein/protein_calculated.csv")
```

Add protein data to the metadata. 

```{r}
metadata$protein.ug<-protein$`Total Protein ug`[match(metadata$tube, protein$Sample)]
```

## Negative polarity 

How many peaks do we have?  

```{r}
length(neg_peaks$`Metabolite name`)
```

162 peaks. 

Reformat data set to be useful.  

```{r}
neg_peaks<-neg_peaks%>% 
  rename(peak=`Alignment ID`, lipid=`Metabolite name`, formula=Formula, ontology=Ontology)%>%
  select(peak, lipid, formula, ontology)%>%
  mutate(polarity=c("negative"))

neg_values<-neg_values%>%
  rename(peak=ID, file=File, group=Class, area=Area)%>%
  select(peak, file, group, area)%>%
  mutate(polarity=c("negative"))
```

The data frames now contain `peak` for a unique ID for each lipid detected. This number links the two files. Area represents the normalized area after internal standard normalization. We will further normalize this value below.  

Format the file names to correspond to sample IDs in the metadata sheet. 

```{r}
#first remove the test PBQCs ran at 5 and 10 uL loadings and keep the standard 20 uL loading 
neg_values<-neg_values%>%
  filter(!file=="20240423_Mcap_neg_PBQC1_5")%>%
  filter(!file=="20240423_Mcap_neg_PBQC1_10")%>%
  mutate(file = str_replace(file, "20240423_Mcap_neg_PBQC1_20", "20240423_Mcap_neg_PBQC1"))%>%
  mutate(sample = str_extract(file, "[^_]+$"))
```

Merge together all data from the metadata, values, and peak information to one data frame. 

```{r}
negative<-left_join(metadata, neg_values)

negative<-negative%>%
  select(sample, temperature, parent, type, peak, area, polarity, protein.ug)

negative<-left_join(negative, neg_peaks)
```

Next, add together the values for all peaks identified as the same metabolite for each sample.   

How many lipids have multiple peaks detected separately? 
```{r}
length(unique(neg_peaks$lipid))
```

There are 162 unique lipids and 162 peaks. 

Show an example of entries that have the same metabolite with multiple peak IDs. 
```{r}
filtered_values <- negative %>%
  group_by(lipid) %>%
  filter(n_distinct(peak) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, peak)

print(filtered_values)
```
No duplicates. 

Add together the values for each sample for each identified lipid. 
```{r}
negative<-negative%>%
  group_by(sample, lipid, temperature, parent, type, polarity, formula, ontology, protein.ug)%>%
  summarise(area=sum(area, na.rm=TRUE))
```

Remove PBQC's and Blanks. These were used for peak identification in MSDIAL and are not used downstream at this point. 
```{r}
negative<-negative%>%
  filter(!type=="PBQC")%>%
  filter(!type=="Blank")
```

Conduct internal standard normalization. 

Peak 16879 (PG) will be used as our internal standard for negative polarity. It was consistent across samples (low fold change), high signal:noise, and is mid-retention. 
```{r}
#pull out the name of the standard being used 
standard_neg<-neg_peaks%>%
  filter(peak=="16879")%>%
  pull(lipid)

#make a dataframe of standard values for each sample
standard_neg_value<-negative%>%
  filter(lipid %in% standard_neg)%>%
  group_by(sample)%>%
  summarise(area_standard=mean(area, na.rm=TRUE))

#add this value back into the master data frame for corrections. 
negative<-negative%>%
  left_join(.,standard_neg_value)
```

Remove sample 1, it did not have a signal for the internal standard. 

```{r}
negative<-negative%>%
  filter(!sample=="1")
```

Normalize to the standard. 
```{r}
negative<-negative%>%
   mutate(area_corrected=area/area_standard)
```

Finally, normalize corrected area to sample protein. This will generate units of area / ug protein. 
```{r}
negative<-negative%>%
  mutate(area_normalized=area_corrected/protein.ug)
```

Repeat for positive polarity.  

## Positive polarity 

How many peaks do we have?  

```{r}
length(pos_peaks$`Metabolite name`)
```

896 peaks 

Reformat data set to be useful.  

```{r}
pos_peaks<-pos_peaks%>% 
  rename(peak=`Alignment ID`, lipid=`Metabolite name`, formula=Formula, ontology=Ontology)%>%
  select(peak, lipid, formula, ontology)%>%
  mutate(polarity=c("positive"))

pos_values<-pos_values%>%
  rename(peak=ID, file=File, group=Class, area=Area)%>%
  select(peak, file, group, area)%>%
  mutate(polarity=c("positive"))
```

The data frames now contain `peak` for a unique ID for each lipid detected. This number links the two files. Area represents the normalized area after internal standard normalization. We will further normalize this value below.  

Format the file names to correspond to sample IDs in the metadata sheet. 

```{r}
pos_values<-pos_values%>%
  mutate(sample = str_extract(file, "[^_]+$"))
```

Merge together all data from the metadata, values, and peak information to one data frame. 

```{r}
positive<-left_join(metadata, pos_values)

positive<-positive%>%
  select(sample, temperature, parent, type, peak, area, polarity, protein.ug)

positive<-left_join(positive, pos_peaks)
```

Next, add together the values for all peaks identified as the same metabolite for each sample.   

How many lipids have multiple peaks detected separately? 
```{r}
length(unique(pos_peaks$lipid))
```

There are 896 unique lipids and 896 total peaks. 

Show an example of entries that have the same metabolite with multiple peak IDs. 
```{r}
filtered_values <- positive %>%
  group_by(lipid) %>%
  filter(n_distinct(peak) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, peak)

print(filtered_values)

levels(factor(filtered_values$lipid))
```

No duplicates

Remove PBQC's and blanks. These were used for peak identification in MSDIAL and are not used downstream at this point. 
```{r}
positive<-positive%>%
  filter(!type=="PBQC")%>%
  filter(!type=="Blank")
```

Conduct internal standard normalization. 

Peak 19029 (LPC) will be used as our internal standard for positive polarity. It was consistent across samples (low fold change), high signal:noise, and was the best representative from standards. 
```{r}
#pull out the name of the standard being used 
standard_pos<-pos_peaks%>%
  filter(peak=="19029")%>%
  pull(lipid)

#make a dataframe of standard values for each sample
standard_pos_value<-positive%>%
  filter(lipid %in% standard_pos)%>%
  group_by(sample)%>%
  summarise(area_standard=mean(area, na.rm=TRUE))

#add this value back into the master data frame for corrections. 
positive<-positive%>%
  left_join(.,standard_pos_value)
```

Remove sample 1, it did not have a signal for the internal standard for negative and therefore is being removed from the data analysis. 

```{r}
positive<-positive%>%
  filter(!sample=="1")
```

Normalize to the standard. 
```{r}
positive<-positive%>%
   mutate(area_corrected=area/area_standard)
```

Finally, normalize corrected area to sample protein. This will generate units of area / ug protein. 

```{r}
positive<-positive%>%
  mutate(area_normalized=area_corrected/protein.ug)
```

## Join polarities together 

Join positive and negative datasets.  

```{r}
length(negative$lipid)
length(positive$lipid)

lipids<-rbind(negative, positive)

length(lipids$lipid)
```

Select polarity for peaks in both polarity files. How many lipids are detected at both polarities? 

```{r}
filtered_values <- lipids %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, polarity)

print(filtered_values)

length(unique(filtered_values$lipid))
```

There are 40 lipids detected at both polarities. Select the polarity with the highest Area value for each lipid. 
```{r}
# Identify lipids with both polarities
lipids_with_both_polarities <- lipids %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) == 2)

# Filter lipids with both polarities to keep only the one with the highest area
filtered_lipids <- lipids_with_both_polarities %>%
  group_by(lipid) %>%
  filter(area == max(area)) %>%
  ungroup()

# Identify lipids that only have one entry (i.e., do not have both polarities)
single_entry_lipids <- lipids %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) == 1) %>%
  ungroup()

# Combine filtered lipids and single entry lipids
lipids_filtered <- bind_rows(filtered_lipids, single_entry_lipids)
```

How many lipids are detected at both polarities? 

```{r}
filtered_values <- lipids_filtered %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, polarity)

print(filtered_values)

length(unique(filtered_values$lipid))
```

Duplicates are now taken care of.  

View ontology categories available. 
```{r}
levels(as.factor(lipids_filtered$ontology))

ontology<-lipids_filtered%>%ungroup()%>%select(lipid, formula, ontology)%>%unique()
```
 [1] "AHexCAS"    "AHexCer"    "AHexCS"     "AHexSIS"    "ASM"        "BileAcid"   "BMP"       
 [8] "BRSE"       "CAR"        "CASE"       "CE"         "Cer_ADS"    "Cer_AP"     "Cer_AS"    
[15] "Cer_EOS"    "Cer_HDS"    "Cer_HS"     "Cer_NDS"    "Cer_NP"     "Cer_NS"     "CL"        
[22] "CoQ"        "DCAE"       "DEGSE"      "DG"         "DGDG"       "DMPE"       "EtherDG"   
[29] "EtherDGDG"  "EtherLPC"   "EtherLPE"   "EtherMGDG"  "EtherOxPC"  "EtherPC"    "EtherPE"   
[36] "EtherPI"    "EtherPS"    "EtherTG"    "FA"         "FAHFA"      "HBMP"       "HexCer_AP" 
[43] "HexCer_HS"  "HexCer_NDS" "KDCAE"      "KLCAE"      "LCAE"       "LPC"        "LPE"       
[50] "MG"         "MGDG"       "NAE"        "NAGly"      "NAGlySer"   "NAOrn"      "NATau"     
[57] "Others"     "OxFA"       "OxPC"       "OxPE"       "OxPI"       "OxTG"       "PA"        
[64] "PC"         "PE"         "PEtOH"      "PG"         "PI"         "PMeOH"      "PS"        
[71] "SHexCer"    "SISE"       "SL"         "SM"         "SSulfate"   "ST"         "TG"        
[78] "TG_EST"     "Unknown"    "VAE"  

```{r}
lipids_filtered<-lipids_filtered%>%
  select(sample, lipid, temperature, parent, type, polarity, formula, ontology, area_normalized)%>%
  group_by(sample, lipid, temperature, parent, type, formula, ontology)%>%
  summarize(area_normalized=sum(area_normalized, na.rm=TRUE))
```

Remove standard lipids - those with (d7) or (d9) in the lipid name.  

```{r}
lipid_d7 <- lipids_filtered %>%
  filter(str_detect(lipid, "\\(d7\\)"))%>%
  pull(lipid)%>%
  unique()

lipid_d9 <- lipids_filtered %>%
  filter(str_detect(lipid, "\\(d9\\)"))%>%
  pull(lipid)%>%
  unique()

lipids_filtered<-lipids_filtered%>%
  filter(!lipid %in% lipid_d7)%>%
  filter(!lipid %in% lipid_d9)
```

```{r}
length(unique(lipids_filtered$lipid))
```

We have 1009 lipids in our dataset.  

```{r}
hist(lipids_filtered$area_normalized)
```

Plot and remove outliers. 

```{r}
wide_data<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

Replace NA's with 0's. 

```{r}
wide_data[is.na(wide_data)] <- 0

wide_data%>%
    select_if(~ any(is.na(.)))
```

```{r}
str(wide_data)

wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data[c(4:1012)], scale=TRUE, center=TRUE) 
```

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = sample), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

Sample 23 and 48 are outliers, remove. 

```{r}
wide_data<-wide_data%>%
  filter(!sample=="23")%>%
  filter(!sample=="48")

lipids_filtered<-lipids_filtered%>%
  filter(!sample=="23")%>%
  filter(!sample=="48")
```

# Conduct filtering for relative standard deviation 

Make a dataframe that calculates the standard deviation of the lipid area within treatment relative to the mean (i.e., how "large" is the SD).  

The formula will be:  

(sd*100)/mean 

Calculate for each lipid in each treatment group. 
```{r}
test<-lipids_filtered%>%
  group_by(lipid, parent, temperature)%>%
  summarise(RSD=(sd(area_normalized)*100)/mean(area_normalized))

length(unique(test$lipid))

hist(test$RSD)
```
1009 lipids before 

Keep lipids with RSD <40 in all treatments.  
```{r}
test <- test %>%
  group_by(lipid, parent, temperature) %>%
  filter(all(RSD < 40))

length(unique(test$lipid))

keep_list<-c(unique(test$lipid))

hist(test$RSD)
```
843 metabolites after 

Filter the filtered dataset to include the selected metabolites after RSD filtering.  
```{r}
lipids_filtered<-lipids_filtered%>%
  filter(lipid %in% keep_list)
```

We are now ready to conduct analysis! 

Export the processed dataset. 

```{r}
write_csv(lipids_filtered, file="output/lipids_metabolites/lipids/processed_lipids.csv")
```

# Unsupervised analysis 

Conduct a PCA and PERMANOVA to examine the effects of parent and temperature on the lipidome of coral larvae. 

Convert to wide format. 

```{r}
wide_data<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
str(wide_data)

wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data[c(4:846)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(wide_data[c(4:846)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = wide_data, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = wide_data, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)  
parent              2     3925 0.09312 2.5252  0.024 *
temperature         2     3346 0.07938 2.1525  0.050 *
parent:temperature  4     2238 0.05309 0.7198  0.703  
Residual           42    32642 0.77441                
Total              50    42150 1.00000      

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = sample), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

```{r}
pca2<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2

pca2b<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2b

pca_plots<-plot_grid(pca2, pca2b, ncol=2, nrow=1)

ggsave(pca_plots, filename="figures/lipids/pca_plots.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-wide_data%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca3

ggsave(pca3, filename="figures/lipids/pca_all_groups.jpeg", width=8, height=6)
```

# Supervised PLS-DA analysis  

Script modified from the following tutorial: https://mixomicsteam.github.io/Bookdown/plsda.html

## PLS-DA for the effect of parent 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

Y <- as.factor(X$parent) #select treatment names
Y

X<-X[4:846] #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=2) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("gray", "orange", "brown4") 

pdf("figures/lipids/PLSDA_Parent.pdf", width=9, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_parent <- plsda(X,Y, ncomp=2) 

#view plsda model again
plotIndiv(MyResult.plsda_parent, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
parent_VIP <- PLSDA.VIP(MyResult.plsda_parent, graph=TRUE)

parent_VIP_DF <- as.data.frame(parent_VIP[["tab"]])
parent_VIP_DF

# Converting row names to column
parent_VIP_table <- rownames_to_column(parent_VIP_DF, var = "Lipid")

#filter for VIP > 1
parent_VIP_1 <- parent_VIP_table %>% 
  filter(VIP >= 1)

write_csv(parent_VIP_1, "output/lipids_metabolites/lipids/Overall_Parent_VIPs.csv")

#plot
parent_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Parent VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


pdf("figures/lipids/Parent_VIP.pdf", width=10, height=35)
parent_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Parent VIP > 1") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
dev.off() 
```

These lists represent VIPs that drive separation between parent groups.  

Make a list of vips. 
```{r}
parent_vip_list<-parent_VIP_1%>%
  pull(Lipid)
```
There are 237 VIPs. 

Plot heatmap of z score of lipids across life stages.  

```{r}
# Filter data for VIP metabolites
parent_vip_data <- lipids_filtered %>% filter(lipid %in% parent_vip_list)

# Convert counts to Z-scores
parent_vip_data <- parent_vip_data %>% 
  group_by(lipid) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(parent, lipid)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- parent_vip_data %>% 
  dplyr::select(lipid, parent, z_score) %>%
  spread(key = parent, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "lipid"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  #row_split=6,
  row_title = "Lipid", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/lipids/VIP_heatmap_parent.pdf", width=4, height=6)
draw(heatmap)
dev.off()
```

View a PCA of these VIPs. 

```{r}
vip_data_parent<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  filter(lipid %in% parent_vip_list)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
vip_data_parent%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_parent$temperature<-factor(vip_data_parent$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_parent$parent<-factor(vip_data_parent$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_parent[c(4:240)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_parent[c(4:240)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_parent, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_parent, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)    
parent              2   2631.8 0.22209 7.0674  0.001 ***
temperature         2    874.1 0.07376 2.3472  0.031 *  
parent:temperature  4    524.0 0.04422 0.7035  0.753    
Residual           42   7820.1 0.65993                  
Total              50  11850.0 1.00000         

```{r}
pca_vip_parent1<-ggplot2::autoplot(scaled.pca, data=vip_data_parent, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_parent1

pca_vip_parent2<-ggplot2::autoplot(scaled.pca, data=vip_data_parent, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_parent2

pca_plots_parent<-plot_grid(pca_vip_parent1, pca_vip_parent2, ncol=2, nrow=1)

ggsave(pca_plots_parent, filename="figures/lipids/pca_plots_parentVIP.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-vip_data_parent%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca_parent3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_parent3

ggsave(pca_parent3, filename="figures/lipids/pca_all_groups_parentVIP.jpeg", width=8, height=6)
```

## PLS-DA for the effect of temperature 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

Y <- as.factor(X$temperature) #select treatment names
Y

levels(Y)

X<-X[4:846] #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=2) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("blue", "orange", "red") 

pdf("figures/lipids/PLSDA_Temperature.pdf", width=9, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_temperature <- plsda(X,Y, ncomp=2) 

#view plsda model again
plotIndiv(MyResult.plsda_temperature, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
temperature_VIP <- PLSDA.VIP(MyResult.plsda_temperature, graph=TRUE)

temperature_VIP_DF <- as.data.frame(temperature_VIP[["tab"]])
temperature_VIP_DF

# Converting row names to column
temperature_VIP_table <- rownames_to_column(temperature_VIP_DF, var = "Lipid")

#filter for VIP > 1
temperature_VIP_1 <- temperature_VIP_table %>% 
  filter(VIP >= 1)

write_csv(temperature_VIP_1, "output/lipids_metabolites/lipids/Overall_Temperature_VIPs.csv")

#plot
temperature_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Temperature VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


pdf("figures/lipids/Temperature_VIP.pdf", width=10, height=35)
temperature_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Temperature VIP > 1") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
dev.off() 
```

These lists represent VIPs that drive separation between temperature groups.  

Make a list of vips. 
```{r}
temperature_vip_list<-temperature_VIP_1%>%
  pull(Lipid)
```
There are 251 VIPs. 

Plot heatmap of z score of lipids across temperature.  

```{r}
# Filter data for VIP metabolites
temperature_vip_data <- lipids_filtered %>% filter(lipid %in% temperature_vip_list)

# Convert counts to Z-scores
temperature_vip_data <- temperature_vip_data %>% 
  group_by(lipid) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(temperature, lipid)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

temperature_vip_data$temperature<-factor(temperature_vip_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- temperature_vip_data %>% 
  dplyr::select(lipid, temperature, z_score) %>%
  spread(key = temperature, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "lipid"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  #row_split=6,
  row_title = "Lipid", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/lipids/VIP_heatmap_temperature.pdf", width=4, height=6)
draw(heatmap)
dev.off()
```

View a PCA of these VIPs. 

```{r}
vip_data_temperature<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  filter(lipid %in% temperature_vip_list)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
vip_data_temperature%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_temperature$temperature<-factor(vip_data_temperature$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_temperature$parent<-factor(vip_data_temperature$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_temperature[c(4:254)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_temperature[c(4:254)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_temperature, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_temperature, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)  
parent              2   1228.1 0.09786 2.7436  0.018 *
temperature         2   1219.2 0.09715 2.7237  0.021 *
parent:temperature  4    702.6 0.05598 0.7848  0.681  
Residual           42   9400.1 0.74901                
Total              50  12550.0 1.00000          

```{r}
pca_vip_temperature1<-ggplot2::autoplot(scaled.pca, data=vip_data_temperature, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_temperature1

pca_vip_temperature2<-ggplot2::autoplot(scaled.pca, data=vip_data_temperature, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_temperature2

pca_plots_temperature<-plot_grid(pca_vip_temperature1, pca_vip_temperature2, ncol=2, nrow=1)

ggsave(pca_plots_temperature, filename="figures/lipids/pca_plots_temperatureVIP.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-vip_data_temperature%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca_temperature3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_temperature3

ggsave(pca_temperature3, filename="figures/lipids/pca_all_groups_temperatureVIP.jpeg", width=8, height=6)
```

## PLS-DA for the effect of parent x temperature 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

X$code<-paste(X$parent, X$temperature)

levels(as.factor(X$code))

Y <- as.factor(X$code) #select treatment names
Y

levels(Y)

X<-X[4:846] #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=5) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("orange", "orange3", "orange4", "brown", "brown2", "brown4", "gray", "lightgray", "darkgray") 

pdf("figures/lipids/PLSDA_interaction.pdf", width=9, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_interaction <- plsda(X,Y, ncomp=5) 

#view plsda model again
plotIndiv(MyResult.plsda_interaction, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent x Temperature", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
interaction_VIP <- PLSDA.VIP(MyResult.plsda_interaction, graph=TRUE)

interaction_VIP_DF <- as.data.frame(interaction_VIP[["tab"]])
interaction_VIP_DF

# Converting row names to column
interaction_VIP_table <- rownames_to_column(interaction_VIP_DF, var = "Lipid")

#filter for VIP > 1
interaction_VIP_1 <- interaction_VIP_table %>% 
  filter(VIP >= 1)

write_csv(interaction_VIP_1, "output/lipids_metabolites/lipids/Overall_Interaction_VIPs.csv")

#plot
interaction_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Parent x Temperature VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


pdf("figures/lipids/Interaction_VIP.pdf", width=10, height=35)
interaction_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Interaction VIP > 1") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
dev.off() 
```

These lists represent VIPs that drive separation between parent x temperature groups.  

Make a list of vips. 
```{r}
interaction_vip_list<-interaction_VIP_1%>%
  pull(Lipid)
```
There are 302 VIPs. 

Plot heatmap of z score of lipids across groups.   

```{r}
# Filter data for VIP metabolites
interaction_vip_data <- lipids_filtered %>% filter(lipid %in% interaction_vip_list)

# Convert counts to Z-scores
interaction_vip_data <- interaction_vip_data %>% 
  mutate(code=paste(parent, temperature))%>%
  group_by(lipid) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(code, lipid)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()


interaction_vip_data$code<-factor(interaction_vip_data$code, levels=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- interaction_vip_data %>% 
  dplyr::select(lipid, code, z_score) %>%
  spread(key = code, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "lipid"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  #row_split=4,
  row_title = "Lipid", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/lipids/VIP_heatmap_parentxtemperature.pdf", width=6, height=8)
draw(heatmap)
dev.off()
```

View a PCA of these VIPs. 

```{r}
vip_data_interaction<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  filter(lipid %in% interaction_vip_list)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
vip_data_interaction%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_interaction$temperature<-factor(vip_data_interaction$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_interaction$parent<-factor(vip_data_interaction$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_interaction[c(4:305)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_interaction[c(4:305)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_interaction, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_interaction, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)    
parent              2   2779.5 0.18407 5.8640  0.001 ***
temperature         2   1432.4 0.09486 3.0220  0.002 ** 
parent:temperature  4    934.5 0.06189 0.9858  0.450    
Residual           42   9953.7 0.65918                  
Total              50  15100.0 1.00000    

```{r}
pca_vip_interaction1<-ggplot2::autoplot(scaled.pca, data=vip_data_interaction, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_interaction1

pca_vip_interaction2<-ggplot2::autoplot(scaled.pca, data=vip_data_interaction, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_interaction2

pca_plots_interaction<-plot_grid(pca_vip_interaction1, pca_vip_interaction2, ncol=2, nrow=1)

ggsave(pca_plots_interaction, filename="figures/lipids/pca_plots_interactionVIP.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-vip_data_interaction%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca_interaction3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_interaction3

ggsave(pca_interaction3, filename="figures/lipids/pca_all_groups_interactionVIP.jpeg", width=8, height=6)
```

# Output lists for LION 

List of parental VIPS
```{r}
parent_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/parent_VIP_list.csv")
```

No significant enrichment of any terms on LION. 

List of temperature VIPS
```{r}
temperature_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/temperature_VIP_list.csv")
```

No significant enrichment of any terms on LION. 

List of interaction VIPS
```{r}
interaction_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/interaction_VIP_list.csv")
```

Full list of lipids
```{r}
full_lipid_list<-lipids_filtered%>%
  pull(lipid)%>%
  unique()%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/full_lipid_list.csv")
```

What is the overlap in these lists? Create an upset plot. 

```{r}
# Combine lists into a data frame indicating membership (1 = present, 0 = absent)
upset_data <- data.frame(
  Lipid = unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)),
  List1 = as.numeric(unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)) %in% parent_vip_list),
  List2 = as.numeric(unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)) %in% temperature_vip_list),
  List3 = as.numeric(unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)) %in% interaction_vip_list)
)

# Convert data to the format needed for the UpSetR plot
plot_data <- upset_data[, -1]
row.names(plot_data) <- upset_data$Lipid
names(plot_data) <- c("Parent", "Temperature", "PxT Group")

# Generate the UpSet plot
pdf("figures/lipids/upset_VIP_list_overlap.pdf", width=6, height=6)
upset(plot_data, sets = c("Parent", "Temperature", "PxT Group"), keep.order = TRUE)
dev.off()
```

# Compare height and area VIP lists 

Plot upset plot for each factor tested between height and area 

Parent VIP lists: 
```{r}
list1<-read_csv("output/lipids_metabolites/lipids/parent_VIP_list.csv")%>%pull(.)
list2<-read_csv("output/lipids_metabolites/lipids/parent_VIP_list_height.csv")%>%pull(.)

# Combine lists into a data frame indicating membership (1 = present, 0 = absent)
upset_data <- data.frame(
  Lipid = unique(c(list1, list2)),
  List1 = as.numeric(unique(c(list1, list2)) %in% list1),
  List2 = as.numeric(unique(c(list1, list2)) %in% list2))

# Convert data to the format needed for the UpSetR plot
plot_data <- upset_data[, -1]
row.names(plot_data) <- upset_data$Lipid
names(plot_data) <- c("Area", "Height")

# Generate the UpSet plot
upset(plot_data, sets = c("Area", "Height"), keep.order = TRUE)
```
218 shared, 31 unique to height, 19 unique to area data sets. 

Temperature VIP lists: 

```{r}
list1<-read_csv("output/lipids_metabolites/lipids/temperature_VIP_list.csv")%>%pull(.)
list2<-read_csv("output/lipids_metabolites/lipids/temperature_VIP_list_height.csv")%>%pull(.)

# Combine lists into a data frame indicating membership (1 = present, 0 = absent)
upset_data <- data.frame(
  Lipid = unique(c(list1, list2)),
  List1 = as.numeric(unique(c(list1, list2)) %in% list1),
  List2 = as.numeric(unique(c(list1, list2)) %in% list2))

# Convert data to the format needed for the UpSetR plot
plot_data <- upset_data[, -1]
row.names(plot_data) <- upset_data$Lipid
names(plot_data) <- c("Area", "Height")

# Generate the UpSet plot
upset(plot_data, sets = c("Area", "Height"), keep.order = TRUE)
```
213 shared, 38 unique to height, 38 unique to area data sets. 

Interaction VIP lists: 

```{r}
list1<-read_csv("output/lipids_metabolites/lipids/interaction_VIP_list.csv")%>%pull(.)
list2<-read_csv("output/lipids_metabolites/lipids/interaction_VIP_list_height.csv")%>%pull(.)

# Combine lists into a data frame indicating membership (1 = present, 0 = absent)
upset_data <- data.frame(
  Lipid = unique(c(list1, list2)),
  List1 = as.numeric(unique(c(list1, list2)) %in% list1),
  List2 = as.numeric(unique(c(list1, list2)) %in% list2))

# Convert data to the format needed for the UpSetR plot
plot_data <- upset_data[, -1]
row.names(plot_data) <- upset_data$Lipid
names(plot_data) <- c("Area", "Height")

# Generate the UpSet plot
upset(plot_data, sets = c("Area", "Height"), keep.order = TRUE)
```
266 shared, 39 unique to height, 36 unique to area data sets.

In each set, most lipids are shared by area and height data sets. 

# Plot enriched lipids in the group VIP data set 

Results from LION: 

- No enrichment in parent list 
- No enrichment in temperature list 
- fatty acids [FA], and fatty acids and conjugates [FA01] enriched in group specific list 

Ontology FA

Add in lipid ontology values. 

```{r}
lipids$ontology<-ontology$ontology[match(lipids$lipid, ontology$lipid)]
```

Plot a heatmap of the target ontologies in the group VIP list (parent x temperature). 

Fatty acids
```{r}
# Filter data for VIP metabolites
FA_data <- lipids %>% 
  filter(lipid %in% interaction_vip_list) %>% 
  filter(ontology=="FA") %>% 
  droplevels()

length(unique(FA_data$lipid))
#16 fatty acids in this data set 

# Convert counts to Z-scores
FA_data <- FA_data %>% 
  group_by(lipid) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(parent, temperature, lipid)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- FA_data %>% 
  mutate(code=paste(parent, temperature))%>%
  mutate(code=factor(code, levels=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)")))%>%
  dplyr::select(lipid, code, z_score) %>%
  spread(key = code, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "lipid"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  #row_split=split,
  #column_split=heatmap_data$code,
  row_title = "Fatty Acids", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/lipids/fatty_acids_VIP.pdf", width=6, height=8)
draw(heatmap)
dev.off()
```
