---
title: "Lipidomics analysis"
author: "Ariana S Huffmyer"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(broom)
library(cowplot)
library(mixOmics)
library(dplyr)
library(conflicted)
library(vegan)
library(factoextra)
library(ggfortify)
library(RVAideMemoire)
library(ComplexHeatmap)
library(circlize)
library(tibble)
library(viridis)
library(genefilter)
library(UpSetR)
library(emmeans)
library(car)
library(limma)

conflict_prefer("select", winner = "dplyr", quiet = FALSE)
conflict_prefer("filter", winner = "dplyr", quiet = FALSE)
conflict_prefer("rename", winner = "dplyr", quiet = FALSE)
```

# Data manipulation and preparations  

## Read in all files 

```{r}
metadata<-read_csv("data/lipids_metabolites/lipidomics/lipid_metadata_huffmyer_matthews.csv")%>%
  rename(sample=`Vial ID`, tube=`Sample ID`, temperature=Temperature, parent=Parent, type=Type, code=Code)

neg_values<-read_csv("data/lipids_metabolites/lipidomics/processed-data/negative/PeakValues_2_2024_09_19_13_01_19.csv")

neg_peaks<-read_csv("data/lipids_metabolites/lipidomics/processed-data/negative/PeakMaster_2_2024_09_19_13_01_19.csv")

pos_values<-read_csv("data/lipids_metabolites/lipidomics/processed-data/positive/PeakValues_3_2024_09_19_14_24_22.csv")

pos_peaks<-read_csv("data/lipids_metabolites/lipidomics/processed-data/positive/PeakMaster_3_2024_09_19_14_24_22.csv")

protein<-read_csv("data/lipids_metabolites/protein/protein_calculated.csv")
```

Add protein data to the metadata. 

```{r}
metadata$protein.ug<-protein$`Total Protein ug`[match(metadata$tube, protein$Sample)]
```

## Negative polarity 

How many peaks do we have?  

```{r}
length(neg_peaks$`Metabolite name`)
```

162 peaks. 

Reformat data set to be useful.  

```{r}
neg_peaks<-neg_peaks%>% 
  rename(peak=`Alignment ID`, lipid=`Metabolite name`, formula=Formula, ontology=Ontology)%>%
  select(peak, lipid, formula, ontology)%>%
  mutate(polarity=c("negative"))

neg_values<-neg_values%>%
  rename(peak=ID, file=File, group=Class, area=Area)%>%
  select(peak, file, group, area)%>%
  mutate(polarity=c("negative"))
```

The data frames now contain `peak` for a unique ID for each lipid detected. This number links the two files. Area represents the normalized area after internal standard normalization. We will further normalize this value below.  

Format the file names to correspond to sample IDs in the metadata sheet. 

```{r}
#first remove the test PBQCs ran at 5 and 10 uL loadings and keep the standard 20 uL loading 
neg_values<-neg_values%>%
  filter(!file=="20240423_Mcap_neg_PBQC1_5")%>%
  filter(!file=="20240423_Mcap_neg_PBQC1_10")%>%
  mutate(file = str_replace(file, "20240423_Mcap_neg_PBQC1_20", "20240423_Mcap_neg_PBQC1"))%>%
  mutate(sample = str_extract(file, "[^_]+$"))
```

Merge together all data from the metadata, values, and peak information to one data frame. 

```{r}
negative<-left_join(metadata, neg_values)

negative<-negative%>%
  select(sample, temperature, parent, type, peak, area, polarity, protein.ug)

negative<-left_join(negative, neg_peaks)
```

Next, add together the values for all peaks identified as the same metabolite for each sample.   

How many lipids have multiple peaks detected separately? 
```{r}
length(unique(neg_peaks$lipid))
```

There are 162 unique lipids and 162 peaks. 

Show an example of entries that have the same metabolite with multiple peak IDs. 
```{r}
filtered_values <- negative %>%
  group_by(lipid) %>%
  filter(n_distinct(peak) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, peak)

print(filtered_values)
```
No duplicates. 

Add together the values for each sample for each identified lipid. 
```{r}
negative<-negative%>%
  group_by(sample, lipid, temperature, parent, type, polarity, formula, ontology, protein.ug)%>%
  summarise(area=sum(area, na.rm=TRUE))
```

Remove PBQC's and Blanks. These were used for peak identification in MSDIAL and are not used downstream at this point. 
```{r}
negative<-negative%>%
  filter(!type=="PBQC")%>%
  filter(!type=="Blank")
```

Conduct internal standard normalization. 

Peak 16879 (PG) will be used as our internal standard for negative polarity. It was consistent across samples (low fold change), high signal:noise, and is mid-retention. 
```{r}
#pull out the name of the standard being used 
standard_neg<-neg_peaks%>%
  filter(peak=="16879")%>%
  pull(lipid)

#make a dataframe of standard values for each sample
standard_neg_value<-negative%>%
  filter(lipid %in% standard_neg)%>%
  group_by(sample)%>%
  summarise(area_standard=mean(area, na.rm=TRUE))

#add this value back into the master data frame for corrections. 
negative<-negative%>%
  left_join(.,standard_neg_value)
```

Remove sample 1, it did not have a signal for the internal standard. 

```{r}
negative<-negative%>%
  filter(!sample=="1")
```

Normalize to the standard. 
```{r}
negative<-negative%>%
   mutate(area_corrected=area/area_standard)
```

Finally, normalize corrected area to sample protein. This will generate units of area / ug protein. 
```{r}
negative<-negative%>%
  mutate(area_normalized=area_corrected/protein.ug)
```

Repeat for positive polarity.  

## Positive polarity 

How many peaks do we have?  

```{r}
length(pos_peaks$`Metabolite name`)
```

896 peaks 

Reformat data set to be useful.  

```{r}
pos_peaks<-pos_peaks%>% 
  rename(peak=`Alignment ID`, lipid=`Metabolite name`, formula=Formula, ontology=Ontology)%>%
  select(peak, lipid, formula, ontology)%>%
  mutate(polarity=c("positive"))

pos_values<-pos_values%>%
  rename(peak=ID, file=File, group=Class, area=Area)%>%
  select(peak, file, group, area)%>%
  mutate(polarity=c("positive"))
```

The data frames now contain `peak` for a unique ID for each lipid detected. This number links the two files. Area represents the normalized area after internal standard normalization. We will further normalize this value below.  

Format the file names to correspond to sample IDs in the metadata sheet. 

```{r}
pos_values<-pos_values%>%
  mutate(sample = str_extract(file, "[^_]+$"))
```

Merge together all data from the metadata, values, and peak information to one data frame. 

```{r}
positive<-left_join(metadata, pos_values)

positive<-positive%>%
  select(sample, temperature, parent, type, peak, area, polarity, protein.ug)

positive<-left_join(positive, pos_peaks)
```

Next, add together the values for all peaks identified as the same metabolite for each sample.   

How many lipids have multiple peaks detected separately? 
```{r}
length(unique(pos_peaks$lipid))
```

There are 896 unique lipids and 896 total peaks. 

Show an example of entries that have the same metabolite with multiple peak IDs. 
```{r}
filtered_values <- positive %>%
  group_by(lipid) %>%
  filter(n_distinct(peak) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, peak)

print(filtered_values)

levels(factor(filtered_values$lipid))
```

No duplicates

Remove PBQC's and blanks. These were used for peak identification in MSDIAL and are not used downstream at this point. 
```{r}
positive<-positive%>%
  filter(!type=="PBQC")%>%
  filter(!type=="Blank")
```

Conduct internal standard normalization. 

Peak 19029 (LPC) will be used as our internal standard for positive polarity. It was consistent across samples (low fold change), high signal:noise, and was the best representative from standards. 
```{r}
#pull out the name of the standard being used 
standard_pos<-pos_peaks%>%
  filter(peak=="19029")%>%
  pull(lipid)

#make a dataframe of standard values for each sample
standard_pos_value<-positive%>%
  filter(lipid %in% standard_pos)%>%
  group_by(sample)%>%
  summarise(area_standard=mean(area, na.rm=TRUE))

#add this value back into the master data frame for corrections. 
positive<-positive%>%
  left_join(.,standard_pos_value)
```

Remove sample 1, it did not have a signal for the internal standard for negative and therefore is being removed from the data analysis. 

```{r}
positive<-positive%>%
  filter(!sample=="1")
```

Normalize to the standard. 
```{r}
positive<-positive%>%
   mutate(area_corrected=area/area_standard)
```

Finally, normalize corrected area to sample protein. This will generate units of area / ug protein. 

```{r}
positive<-positive%>%
  mutate(area_normalized=area_corrected/protein.ug)
```

## Join polarities together 

Join positive and negative datasets.  

```{r}
length(negative$lipid)
length(positive$lipid)

lipids<-rbind(negative, positive)

length(lipids$lipid)
```

Select polarity for peaks in both polarity files. How many lipids are detected at both polarities? 

```{r}
filtered_values <- lipids %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, polarity)

print(filtered_values)

length(unique(filtered_values$lipid))
```

There are 40 lipids detected at both polarities. Select the polarity with the highest Area value for each lipid. 
```{r}
# Identify lipids with both polarities
lipids_with_both_polarities <- lipids %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) == 2)

# Filter lipids with both polarities to keep only the one with the highest area
filtered_lipids <- lipids_with_both_polarities %>%
  group_by(lipid) %>%
  filter(area == max(area)) %>%
  ungroup()

# Identify lipids that only have one entry (i.e., do not have both polarities)
single_entry_lipids <- lipids %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) == 1) %>%
  ungroup()

# Combine filtered lipids and single entry lipids
lipids_filtered <- bind_rows(filtered_lipids, single_entry_lipids)
```

How many lipids are detected at both polarities? 

```{r}
filtered_values <- lipids_filtered %>%
  group_by(lipid) %>%
  filter(n_distinct(polarity) > 1) %>%
  ungroup()%>%
  arrange(lipid, sample, polarity)

print(filtered_values)

length(unique(filtered_values$lipid))
```

Duplicates are now taken care of.  

View ontology categories available. 
```{r}
levels(as.factor(lipids_filtered$ontology))

ontology<-lipids_filtered%>%ungroup()%>%select(lipid, formula, ontology)%>%unique()
```
 [1] "AHexCAS"    "AHexCer"    "AHexCS"     "AHexSIS"    "ASM"        "BileAcid"   "BMP"       
 [8] "BRSE"       "CAR"        "CASE"       "CE"         "Cer_ADS"    "Cer_AP"     "Cer_AS"    
[15] "Cer_EOS"    "Cer_HDS"    "Cer_HS"     "Cer_NDS"    "Cer_NP"     "Cer_NS"     "CL"        
[22] "CoQ"        "DCAE"       "DEGSE"      "DG"         "DGDG"       "DMPE"       "EtherDG"   
[29] "EtherDGDG"  "EtherLPC"   "EtherLPE"   "EtherMGDG"  "EtherOxPC"  "EtherPC"    "EtherPE"   
[36] "EtherPI"    "EtherPS"    "EtherTG"    "FA"         "FAHFA"      "HBMP"       "HexCer_AP" 
[43] "HexCer_HS"  "HexCer_NDS" "KDCAE"      "KLCAE"      "LCAE"       "LPC"        "LPE"       
[50] "MG"         "MGDG"       "NAE"        "NAGly"      "NAGlySer"   "NAOrn"      "NATau"     
[57] "Others"     "OxFA"       "OxPC"       "OxPE"       "OxPI"       "OxTG"       "PA"        
[64] "PC"         "PE"         "PEtOH"      "PG"         "PI"         "PMeOH"      "PS"        
[71] "SHexCer"    "SISE"       "SL"         "SM"         "SSulfate"   "ST"         "TG"        
[78] "TG_EST"     "Unknown"    "VAE"  

```{r}
lipids_filtered<-lipids_filtered%>%
  select(sample, lipid, temperature, parent, type, polarity, formula, ontology, area_normalized)%>%
  group_by(sample, lipid, temperature, parent, type, formula, ontology)%>%
  summarize(area_normalized=sum(area_normalized, na.rm=TRUE))
```

Remove standard lipids - those with (d7) or (d9) in the lipid name.  

```{r}
lipid_d7 <- lipids_filtered %>%
  filter(str_detect(lipid, "\\(d7\\)"))%>%
  pull(lipid)%>%
  unique()

lipid_d9 <- lipids_filtered %>%
  filter(str_detect(lipid, "\\(d9\\)"))%>%
  pull(lipid)%>%
  unique()

lipids_filtered<-lipids_filtered%>%
  filter(!lipid %in% lipid_d7)%>%
  filter(!lipid %in% lipid_d9)
```

```{r}
length(unique(lipids_filtered$lipid))
```

We have 1009 lipids in our dataset.  

```{r}
hist(lipids_filtered$area_normalized)
summary(lipids_filtered$area_normalized)
max(lipids_filtered$area_normalized)
```

There is one value >10 with other values, remove this value. 
```{r}
lipids_filtered<-lipids_filtered%>%
  filter(area_normalized<10)

hist(lipids_filtered$area_normalized)
```

Plot and remove outliers. 

```{r}
wide_data<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

Replace NA's and -Inf with 0's. 

```{r}
wide_data[is.na(wide_data)] <- 0

wide_data%>%
    select_if(~ any(is.na(.)))

wide_data%>%
    select_if(~ any(is.infinite(.)))

wide_data <- wide_data %>%
  mutate(across(everything(), ~ replace(., is.infinite(.), 0)))

wide_data%>%
    select_if(~ any(is.infinite(.)))
```

```{r}
str(wide_data)

wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data%>%select(where(is.numeric)), scale=TRUE, center=TRUE) 
```

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = sample), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

Sample 23 and 48 is an outlier, remove. 

```{r}
wide_data<-wide_data%>%
  filter(!sample=="23")%>%
  filter(!sample=="48")

lipids_filtered<-lipids_filtered%>%
  filter(!sample=="23")%>%
  filter(!sample=="48")
```

# Conduct filtering for relative standard deviation 

Make a dataframe that calculates the standard deviation of the lipid area within treatment relative to the mean (i.e., how "large" is the SD).  

The formula will be:  

(sd*100)/mean 

Calculate for each lipid in each treatment group. 
```{r}
test<-lipids_filtered%>%
  group_by(lipid, parent, temperature)%>%
  summarise(RSD=(sd(area_normalized)*100)/mean(area_normalized))

length(unique(test$lipid))

hist(test$RSD)
```
1009 lipids before 

Keep lipids with RSD <40 in all treatments.  
```{r}
test <- test %>%
  group_by(lipid, parent, temperature) %>%
  filter(all(RSD < 40))

length(unique(test$lipid))

keep_list<-c(unique(test$lipid))

hist(test$RSD)

summary(test$RSD)
```
843 lipids after 

Filter the filtered dataset to include the selected metabolites after RSD filtering.  
```{r}
lipids_filtered<-lipids_filtered%>%
  filter(lipid %in% keep_list)
```

We are now ready to conduct analysis! 

Export the processed dataset. 

```{r}
write_csv(lipids_filtered, file="output/lipids_metabolites/lipids/processed_lipids.csv")
```

# Unsupervised analysis 

Conduct a PCA and PERMANOVA to examine the effects of parent and temperature on the lipidome of coral larvae. 

Convert to wide format. 

```{r}
wide_data<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
str(wide_data)

wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data%>%select(where(is.numeric)), scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(wide_data%>%select(where(is.numeric)))

# PerMANOVA 
set.seed(123)

permanova<-adonis2(vegan ~ parent*temperature, data = wide_data, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = wide_data, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)  
parent              2     3925 0.09312 2.5252  0.029 *
temperature         2     3346 0.07938 2.1525  0.042 *
parent:temperature  4     2238 0.05309 0.7198  0.727  
Residual           42    32642 0.77441                
Total              50    42150 1.00000      

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = sample), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

```{r}
parent_names<-c("Bleached" = "Susceptible (C)", "Nonbleached" = "Resistant (C/D)", "Wildtype" = "Wildtype (C/D)")

parent_colors<-c("Bleached" = "orange",   # calm blue
                 "Nonbleached" = "darkred", # stressed purple
                 "Wildtype" = "darkgray") 

temp_colors <- c("Ambient" = "#2c7bb6",   # calm blue
                 "Moderate (+3)" = "#7b3294", # stressed purple
                 "High (+6)" = "#d53e4f") 


pca2<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=parent_colors, labels=parent_names)+
  scale_fill_manual(name="Parental Phenotype", values=parent_colors, labels=parent_names)+
  theme_classic()+
  geom_text(x=0, y=-0.35, label="p=0.029", size=6)+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2

pca2b<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="temperature", loadings=FALSE,  colour="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=temp_colors)+
  scale_fill_manual(name="Temperature", values=temp_colors)+
  theme_classic()+
  geom_text(x=0, y=-0.35, label="p=0.042", size=6)+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2b

pca_plots<-plot_grid(pca2, pca2b, ncol=2, nrow=1, labels=c("A", "B"))

ggsave(pca_plots, filename="figures/lipids/pca_plots.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-wide_data%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca3

ggsave(pca3, filename="figures/lipids/pca_all_groups.jpeg", width=8, height=6)
```

Plot by each temperature colored by parent temperature. 

```{r}
pca4<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  facet_wrap(~temperature)+
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca4
```

# Supervised PLS-DA analysis - parent and temperature effects  

Script modified from the following tutorial: https://mixomicsteam.github.io/Bookdown/plsda.html

We will examine main effects of parent and temperature because there is no interactive effect present. 

## PLS-DA for the effect of parent 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

Y <- as.factor(X$parent) #select treatment names
Y

X<-X%>%select(where(is.numeric)) #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=2) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("gray", "orange", "brown4") 

pdf("figures/lipids/PLSDA_Parent.pdf", width=6, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=FALSE, ellipse = TRUE, title="A", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=FALSE, ellipse = TRUE, title="A", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_parent <- plsda(X,Y, ncomp=2) 

#view plsda model again
plotIndiv(MyResult.plsda_parent, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
parent_VIP <- PLSDA.VIP(MyResult.plsda_parent, graph=TRUE)

parent_VIP_DF <- as.data.frame(parent_VIP[["tab"]])
parent_VIP_DF

# Converting row names to column
parent_VIP_table <- rownames_to_column(parent_VIP_DF, var = "Lipid")

#filter for VIP > 1
parent_VIP_1 <- parent_VIP_table %>% 
  filter(VIP >= 2)

write_csv(parent_VIP_1, "output/lipids_metabolites/lipids/Overall_Parent_VIPs.csv")

#plot
parent_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Parent VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

These lists represent VIPs that drive separation between parent groups.  

Make a list of vips. 
```{r}
parent_vip_list<-parent_VIP_1%>%
  pull(Lipid)
```
There are 30 VIPs. 

Plot heatmap of z score of VIP lipids across life stages.  

```{r}
# Filter data for VIP metabolites
parent_vip_data <- lipids_filtered %>% filter(lipid %in% parent_vip_list)

# Convert counts to Z-scores
parent_vip_data <- parent_vip_data %>% 
  group_by(lipid) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(parent, lipid)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- parent_vip_data %>% 
  dplyr::select(lipid, parent, z_score) %>%
  spread(key = parent, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "lipid"))
colnames(heatmap_matrix)<-c("Susceptible (C)", "Resistant (C/D)", "Wildtype (C/D)")

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_split=6,
  row_title = "Lipid", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE), 
  row_names_max_width = unit(10, "cm")
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/lipids/VIP_heatmap_parent.pdf", width=8, height=6)
draw(heatmap)
dev.off()
```

Plot lollipop plot of VIP values for these compounds. 

```{r}
#Add lollipop plot of VIP importance (colored by class of lipid) - lipid on y, vip on x, color by class

lollipop_parent<-parent_VIP_1%>%
   mutate(class = str_extract(Lipid, "^[A-Za-z]+"))

library(RColorBrewer)

okabe_ito_12 <- c(
  "#E69F00", # orange
  "#56B4E9", # sky blue
  "#009E73", # bluish green
  "#F0E442", # yellow
  "#0072B2", # blue
  "#D55E00", # vermillion
  "#CC79A7", # reddish purple
  "#000000", # black
  "#999999", # grey
  "#117733", # dark green
  "#882255", # wine
  "#88CCEE"  # light teal
)

lolli_parent_plot<-lollipop_parent%>%
  
  ggplot(aes(x=VIP, y=reorder(Lipid, VIP), color=class))+
  geom_point()+
  geom_segment(aes(x = 2, xend = VIP, y = Lipid, yend = Lipid), linewidth = 0.8) +
  ylab("Lipid")+
  scale_color_manual(values = okabe_ito_12) +
  xlim(2, 2.8)+
  xlab("Variable Importance in Projection")+
  ggtitle("A")+
  theme_classic()+
  theme(
    legend.title=element_blank()
  );lolli_parent_plot

ggsave(lolli_parent_plot, filename="figures/lipids/VIP_scores_parent.png", width=7, height=5)
```

View a PCA of these VIPs. 

```{r}
vip_data_parent<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  filter(lipid %in% parent_vip_list)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
vip_data_parent%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_parent$temperature<-factor(vip_data_parent$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_parent$parent<-factor(vip_data_parent$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_parent%>%select(where(is.numeric)), scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_parent%>%select(where(is.numeric)))

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_parent, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_parent, method = "eu")
                   Df SumOfSqs      R2       F Pr(>F)    
parent              2   706.82 0.47121 23.1051  0.001 ***
temperature         2    97.80 0.06520  3.1970  0.012 *  
parent:temperature  4    52.96 0.03530  0.8655  0.562    
Residual           42   642.42 0.42828                   
Total              50  1500.00 1.00000       

```{r}
pca_vip_parent1<-ggplot2::autoplot(scaled.pca, data=vip_data_parent, frame.colour="parent", loadings=FALSE,  colour="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"), labels=parent_names)+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"), labels=parent_names)+
  theme_classic()+
  ggtitle("A")+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_parent1

ggsave(pca_vip_parent1, filename="figures/lipids/pca_plots_parentVIP.jpeg", width=6, height=4)
```

## PLS-DA for the effect of temperature 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

Y <- as.factor(X$temperature) #select treatment names
Y

levels(Y)

X<-X%>%select(where(is.numeric)) #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=2) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-temp_colors 

pdf("figures/lipids/PLSDA_Temperature.pdf", width=6, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=FALSE, legend.title = "B", ellipse = TRUE, title="B", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=FALSE, legend.title = "B", ellipse = TRUE, title="B", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_temperature <- plsda(X,Y, ncomp=2) 

#view plsda model again
plotIndiv(MyResult.plsda_temperature, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
temperature_VIP <- PLSDA.VIP(MyResult.plsda_temperature, graph=TRUE)

temperature_VIP_DF <- as.data.frame(temperature_VIP[["tab"]])
temperature_VIP_DF

# Converting row names to column
temperature_VIP_table <- rownames_to_column(temperature_VIP_DF, var = "Lipid")

#filter for VIP > 2
temperature_VIP_1 <- temperature_VIP_table %>% 
  filter(VIP >= 2)

write_csv(temperature_VIP_1, "output/lipids_metabolites/lipids/Overall_Temperature_VIPs.csv")

#plot
temperature_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Lipid,VIP,sum))) +
  geom_point() +
  ylab("Lipid") +
  xlab("VIP Score") +
  ggtitle("Temperature VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

These lists represent VIPs that drive separation between temperature groups.  

Make a list of vips. 
```{r}
temperature_vip_list<-temperature_VIP_1%>%
  pull(Lipid)
```
There are 23 VIPs. 

Plot heatmap of z score of lipids across temperature.  

```{r}
# Filter data for VIP metabolites
temperature_vip_data <- lipids_filtered %>% filter(lipid %in% temperature_vip_list)

# Convert counts to Z-scores
temperature_vip_data <- temperature_vip_data %>% 
  group_by(lipid) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(temperature, lipid)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

temperature_vip_data$temperature<-factor(temperature_vip_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- temperature_vip_data %>% 
  dplyr::select(lipid, temperature, z_score) %>%
  spread(key = temperature, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "lipid"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_split=6,
  row_title = "Lipid", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE), 
  row_names_max_width = unit(10, "cm")
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/lipids/VIP_heatmap_temperature.pdf", width=8, height=6)
draw(heatmap)
dev.off()
```

```{r}
#Add lollipop plot of VIP importance (colored by class of lipid) - lipid on y, vip on x, color by class

lollipop_temp<-temperature_VIP_1%>%
   mutate(class = str_extract(Lipid, "^[A-Za-z]+"))

okabe_ito_12 <- c(
  "#E69F00", # orange
  "#56B4E9", # sky blue
  "#009E73", # bluish green
  "#F0E442", # yellow
  "#0072B2", # blue
  "#D55E00", # vermillion
  "#CC79A7", # reddish purple
  "#000000", # black
  "#999999", # grey
  "#117733", # dark green
  "#882255", # wine
  "#88CCEE"  # light teal
)

lolli_temp_plot<-lollipop_temp%>%
  
  ggplot(aes(x=VIP, y=reorder(Lipid, VIP), color=class))+
  geom_point()+
  geom_segment(aes(x = 2, xend = VIP, y = Lipid, yend = Lipid), linewidth = 0.8) +
  ylab("Lipid")+
  scale_color_manual(values = okabe_ito_12) +
  xlim(2, 4.2)+
  xlab("Variable Importance in Projection")+
  ggtitle("C")+
  theme_classic()+
  theme(
    legend.title=element_blank()
  );lolli_temp_plot

ggsave(lolli_temp_plot, filename="figures/lipids/VIP_scores_temperature.png", width=7, height=5)
```

View a PCA of these VIPs. 

```{r}
vip_data_temperature<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  select(!ontology)%>%
  select(!type)%>%
  filter(lipid %in% temperature_vip_list)%>%
  pivot_wider(names_from=lipid, values_from=area_normalized)
```

```{r}
vip_data_temperature%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_temperature$temperature<-factor(vip_data_temperature$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_temperature$parent<-factor(vip_data_temperature$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_temperature%>%select(where(is.numeric)), scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_temperature%>%select(where(is.numeric)))

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_temperature, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_temperature, method = "eu")
                   Df SumOfSqs      R2       F Pr(>F)    
parent              2    85.37 0.07424  2.9998  0.010 ** 
temperature         2   374.96 0.32605 13.1754  0.001 ***
parent:temperature  4    92.03 0.08003  1.6169  0.061 .  
Residual           42   597.64 0.51969                   
Total              50  1150.00 1.00000          

```{r}
pca_vip_temperature2<-ggplot2::autoplot(scaled.pca, data=vip_data_temperature, frame.colour="temperature", loadings=FALSE,  colour="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  theme_classic()+
  ggtitle("B")+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_temperature2

ggsave(pca_vip_temperature2, filename="figures/lipids/pca_plots_temperatureVIP.jpeg", width=6, height=4)
```

# Output lists for LION and plot LION results  

List of parental VIPS
```{r}
parent_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/parent_VIP_list.csv")
```

Enrichment results saved in Output. 

List of temperature VIPS
```{r}
temperature_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/temperature_VIP_list.csv")
```

Enrichment results saved in Output. 

Full list of lipids
```{r}
full_lipid_list<-lipids_filtered%>%
  pull(lipid)%>%
  unique()%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/lipids/full_lipid_list.csv")
```

# Perform ANOVAs of lipid concentration between groups 

Perform an ANOVA with FDR corrections for lipids to detect which lipids are different between groups. 

Run model for each lipid and apply FDR p value correction. 
```{r}
name_list<-c(unique(lipids_filtered$lipid))

#Create an empty data frame to store the results
results_table <- data.frame(Lipid = character(),
                            Effect = character(),
                            DF = numeric(),
                            SumSq = numeric(),
                            F_value = numeric(),
                            p_value = numeric())

#Loop over each compound in compound_list
for(compound in name_list){
  
  #Fit the linear mixed effect model for each compound
  model <- lipids_filtered%>%
    filter(lipid==compound)%>%
    
    aov(area_normalized ~ temperature * parent, data = .)
  
  qqPlot(residuals(model))

  #Extract the relevant coefficients for interaction of treatment x position
  anova_output <- anova(model)
  coef_summary <- tidy(anova_output)
  coef_summary <- as.data.frame(coef_summary)
  
  #Store the results in the results_table data frame
  results_table <- rbind(results_table, data.frame(Lipid = compound,
                                                   Effect = coef_summary["term"],
                                                   DF = coef_summary["df"],
                                                   SumSq = coef_summary["sumsq"],
                                                   F_value = coef_summary["statistic"],
                                                   p_value = coef_summary["p.value"]))
}


results_table<-results_table%>%
  filter(!term=="Residuals")

#apply a p value correction (FDR) for multiple comparisons 
results_table$adj_pvalues <- p.adjust(results_table$p.value, method = "fdr")

#view only p<0.05
results_table_sign<-results_table%>%
  filter(adj_pvalues<=0.05)

head(results_table_sign)

write.csv(results_table, file="output/lipids_metabolites/lipids/anova_full_results.csv")
```

Residuals pass 

# View concentrations of lipids by class 

Summarize lipid concentration by class for each sample. 

```{r}
classes<-lipids_filtered%>%
  ungroup()%>%
  select(!type)%>%
  select(!formula)%>%
  group_by(sample, temperature, parent, ontology)%>%
  summarise(class_area=sum(area_normalized))
```

Run model for each ontology. 

```{r}
ontology_list<-c(unique(classes$ontology))

#Create an empty data frame to store the results
results_table <- data.frame(Ontology = character(),
                            Effect = character(),
                            DF = numeric(),
                            SumSq = numeric(),
                            F_value = numeric(),
                            p_value = numeric())

#Loop over each compound in compound_list
for(class in ontology_list){
  
  #Fit the linear mixed effect model for each compound
  model <- classes%>%
    filter(ontology==class)%>%
    
    aov(class_area ~ temperature * parent, data = .)
  
  qqPlot(residuals(model))

  #Extract the relevant coefficients for interaction of treatment x position
  anova_output <- anova(model)
  coef_summary <- tidy(anova_output)
  coef_summary <- as.data.frame(coef_summary)
  
  #Store the results in the results_table data frame
  results_table <- rbind(results_table, data.frame(Ontology = class,
                                                   Effect = coef_summary["term"],
                                                   DF = coef_summary["df"],
                                                   SumSq = coef_summary["sumsq"],
                                                   F_value = coef_summary["statistic"],
                                                   p_value = coef_summary["p.value"]))
}


results_table<-results_table%>%
  filter(!term=="Residuals")

#apply a p value correction (FDR) for multiple comparisons 
results_table$adj_pvalues <- p.adjust(results_table$p.value, method = "fdr")

#view only p<0.05
results_table_sign<-results_table%>%
  filter(adj_pvalues<=0.05)

results_table_sign
```

Residuals pass 

   Ontology        term df        sumsq statistic      p.value  adj_pvalues
1  BileAcid temperature  2 7.800681e-06  8.327628 8.990615e-04 0.0277424700
2        CL      parent  2 1.387019e-04 13.694806 2.636644e-05 0.0014237878
3   Cer_ADS      parent  2 1.960034e-04 10.344704 2.224116e-04 0.0080068191
4 EtherDGDG temperature  2 1.485662e-06 10.393233 2.153021e-04 0.0080068191
5 EtherMGDG      parent  2 7.278373e-06 14.517823 1.611506e-05 0.0011602844
6      LCAE      parent  2 1.187502e-05 16.440865 5.325273e-06 0.0007800822
7        SL      parent  2 5.213944e-04 15.901355 7.222983e-06 0.0007800822

Plot total class concentration for these classes.  

```{r}
plot_c1<-classes%>%
  filter(ontology %in% c("CL", "Cer_ADS", "EtherMGDG", "LCAE", "SL"))%>%
  
  ggplot(aes(x=paste(parent), y=class_area, color=parent))+
  facet_wrap(~ontology, scales="free")+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  theme_classic()+
  ylab("Class Concentration")+
  theme(legend.title=element_blank())+
  xlab("Parent Phenotype")+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  #scale_x_discrete(limits=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_c1
```

```{r}
plot_c2<-classes%>%
  filter(ontology %in% c("BileAcid", "EtherDGDG"))%>%
  
  ggplot(aes(x=paste(temperature), y=class_area, color=temperature))+
  facet_wrap(~ontology, scales="free")+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  ylab("Class Concentration")+
  xlab("Temperature")+
  theme_classic()+
  theme(legend.title=element_blank())+
  scale_color_manual(values=c("blue", "red", "orange"))+
  scale_x_discrete(limits=c("Ambient", "Moderate (+3)", "High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_c2


```

# View total lipids 

Add up total lipids - add up all classes

```{r}
totals<-classes%>%
  group_by(sample, temperature, parent)%>%
  summarise(total_area=sum(class_area))
```

View total lipids by treatment group. 

```{r}
plot_t1<-totals%>%
  
  ggplot(aes(x=paste(parent, temperature), y=total_area, color=parent))+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  theme_classic()+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  scale_x_discrete(limits=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_t1

ggsave(plot_t1, filename="figures/lipids/total_lipids.jpeg", width=8, height=8)
```

Run a model to analyze total lipids. 

```{r}
model<-totals%>%
  
  aov(total_area ~ temperature * parent, data=.)

summary(model)
```
No effect on total lipids, almost an effect of temperature but not quite.  

# View class composition as percentage of total lipids 

Calculate lipid proportion of each class as a function of the total lipids. 

```{r}
classes<-left_join(classes, totals)

classes$prop_class<-(classes$class_area/classes$total_area)*100
```

Most lipids (50-70%) are fatty acids. 

Calculate mean of each class. 
```{r}
summary<-classes%>%
  group_by(ontology)%>%
  summarise(mean=mean(prop_class))%>%
  arrange(desc(mean))%>%
  
  write.csv("output/lipids_metabolites/lipids/class_mean_percentage_composition.csv")
```


View percent lipids for each class. 

Fatty acids
```{r}
plot_t2<-classes%>%
  filter(ontology=="FA")%>%
  
  ggplot(aes(x=paste(parent, temperature), y=prop_class, color=parent))+
  facet_wrap(~ontology)+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  theme_classic()+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  scale_x_discrete(limits=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_t2

```

TGs
```{r}
plot_t2<-classes%>%
  filter(ontology=="TG")%>%
  
  ggplot(aes(x=paste(parent, temperature), y=prop_class, color=parent))+
  facet_wrap(~ontology)+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  theme_classic()+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  scale_x_discrete(limits=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_t2

```

Plot stacked bar plot by sample.  
```{r}
plot_t3<-classes%>%
  
  ggplot(aes(x=paste(sample), y=prop_class, fill=ontology))+
  geom_bar(stat="identity")+
  theme_classic()+
  #scale_x_discrete(limits=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_t3
```

Run a model
```{r}
model<-classes%>%
  filter(ontology=="FA")%>%
  
  aov(prop_class ~ temperature*parent, data=.)

summary(model)
```

```{r}
model<-classes%>%
  filter(ontology=="TG")%>%
  
  aov(prop_class ~ temperature*parent, data=.)

summary(model)
```

Run model for each ontology. 

```{r}
ontology_list<-c(unique(classes$ontology))

#Create an empty data frame to store the results
results_table <- data.frame(Ontology = character(),
                            Effect = character(),
                            DF = numeric(),
                            SumSq = numeric(),
                            F_value = numeric(),
                            p_value = numeric())

#Loop over each compound in compound_list
for(class in ontology_list){
  
  #Fit the linear mixed effect model for each compound
  model <- classes%>%
    filter(ontology==class)%>%
    
    aov(prop_class ~ temperature * parent, data = .)
  
  qqPlot(residuals(model))

  #Extract the relevant coefficients for interaction of treatment x position
  anova_output <- anova(model)
  coef_summary <- tidy(anova_output)
  coef_summary <- as.data.frame(coef_summary)
  
  #Store the results in the results_table data frame
  results_table <- rbind(results_table, data.frame(Ontology = class,
                                                   Effect = coef_summary["term"],
                                                   DF = coef_summary["df"],
                                                   SumSq = coef_summary["sumsq"],
                                                   F_value = coef_summary["statistic"],
                                                   p_value = coef_summary["p.value"]))
}


results_table<-results_table%>%
  filter(!term=="Residuals")

#apply a p value correction (FDR) for multiple comparisons 
results_table$adj_pvalues <- p.adjust(results_table$p.value, method = "fdr")

#view only p<0.05
results_table_sign<-results_table%>%
  filter(adj_pvalues<=0.05)

results_table_sign

list<-c(results_table_sign$Ontology)

#output
write.csv(results_table, "output/lipids_metabolites/lipids/class_anova_results_table.csv")
```

     Ontology        term df        sumsq statistic      p.value  adj_pvalues
1     AHexCAS temperature  2 0.0003345747 10.147535 2.539264e-04 5.484810e-03
2      AHexCS      parent  2 0.0185438708 10.286411 2.312780e-04 5.484810e-03
3          CL      parent  2 0.0218972383 28.010323 1.864517e-08 2.013678e-06
4     Cer_ADS      parent  2 0.0276896801 17.926508 2.352073e-06 1.270119e-04
5        DGDG      parent  2 0.0459064060  6.965348 2.441071e-03 3.882885e-02
6    EtherLPE      parent  2 0.0012377785 13.205228 3.553554e-05 1.096525e-03
7   EtherMGDG      parent  2 0.0008023878 21.970836 2.950905e-07 2.124652e-05
8  HexCer_NDS      parent  2 0.0003765059  6.924754 2.516684e-03 3.882885e-02
9        LCAE      parent  2 0.0018770011 12.765934 4.661772e-05 1.258678e-03
10        LPC      parent  2 0.0273479293  9.198133 4.864331e-04 9.551778e-03
11       OxFA      parent  2 0.0960743056 14.724233 1.426868e-05 5.313635e-04
12       SISE      parent  2 0.0001306493  8.984493 5.646436e-04 1.016358e-02
13         SL      parent  2 0.0516161870 43.822892 5.251806e-11 1.134390e-08
14         ST      parent  2 0.0060952440 14.666678 1.476010e-05 5.313635e-04

Plot all classes 
```{r}
plot_sign1<-classes%>%
  
  ggplot(aes(x=temperature, y=prop_class, color=parent))+
  facet_wrap(~ontology)+
  geom_boxplot(outliers=FALSE)+
  geom_point(position=position_dodge(0.75))+
  theme_classic()+
  ylab("Proportion Class Concentration")+
  xlab("Temperature")+
  theme(legend.position="none")+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  scale_x_discrete(limits=c("Ambient", "Moderate (+3)", "High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_sign1
```

Plot AHexCAS (sign by temperature) - higher at high temperature.  

```{r}
plot_sign1<-classes%>%
  filter(ontology=="AHexCAS")%>%
  
  ggplot(aes(x=paste(temperature), y=prop_class, color=temperature))+
  facet_wrap(~ontology)+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  theme_classic()+
  ylab("% Class Composition")+
  xlab("Temperature")+
  ylim(0,0.035)+
  theme(legend.position="none")+
  scale_color_manual(values=temp_colors)+
  scale_x_discrete(limits=c("Ambient", "Moderate (+3)", "High (+6)"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_sign1

ggsave(plot_sign1, filename="figures/lipids/sign_classes_temperature.jpeg", width=4, height=4)
```

- AHexCAS = other class of sterol lipids (acylhexosyl campesterol); higher at +6C temperature; this makes sense to maintain membrane stability under thermal stress 

```{r}
plot_sign2<-classes%>%
  filter(ontology %in% list)%>%
  filter(!ontology=="AHexCAS")%>%
  
  ggplot(aes(x=paste(parent), y=prop_class, color=parent))+
  facet_wrap(~ontology, nrow=3, scales="free_y")+
  geom_boxplot(outliers=FALSE)+
  geom_point()+
  theme_classic()+
    ylab("% Class Composition")+
  xlab("Phenotype")+
    theme(legend.position="none")+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  scale_x_discrete(labels=parent_names)+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot_sign2

ggsave(plot_sign2, filename="figures/lipids/sign_classes_parent.jpeg", width=9, height=6)
```

https://systemsomicslab.github.io/compms/msdial/lipidnomenclature.html

- AHexCS = other class of sterol lipids (acylhexosyl cholesterol); lower in bleached; higher in high temp 
- Cer_ADS = Ceramides Sphinogolipids (ceramide alpha-hydroxy fatty acid-dihydrosphingosine); lower in bleached 
- CL = glycerophospholipids (cardiolipin); lower in bleached 
- DGDG = glycerolipids (digalactosyldiacylglycerol); higher in bleached
- EtherLPE = Glycerophospholipids (Ether-linked lysophosphatidylethanolamine); higher in bleached
- EtherMGDG = Glycerolipids (Ether-linked monogalactosyldiacylglycerol); higher in bleached
- HexCer_NDS = Sphingolipids, neutral glycosphingolipids (Hexosylceramide non-hydroxyfatty acid-dihydrosphingosine); higher in bleached 
- LCAE = fatty acid (long-chain aliphatic ester); lower in bleached 
- LPC = Glycerophospholipid (Lysophophatidylcholine); lower in bleached 
- OxFA = Oxidized fatty acids; lower in wildtype, similar in bleached/non bleached
- SISE = Sterol lipids (Sitosterol ester); higher in wiltyp, similar in bleached/non bleached
- SL = Sphingolipids (sulfonolipid); higher in bleached, lowest in wildtype
- ST = Sterol lipids; lower in bleached

# Examine fatty acid and TAG saturation states between temperature and parent 

## Plot composition of free fatty acids 

FAs - polyunsaturated (":2" or more)
```{r}
fa_pufa<-lipids_filtered %>%
  mutate(class = str_extract(lipid, "^[A-Za-z]+"))%>%
  filter(ontology=="FA")

fa_pufa<-fa_pufa[grepl(":([2-9]|[1-9][0-9])$", fa_pufa$lipid), ]
```

```{r}
fa_plot1<-fa_pufa%>%
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");fa_plot1
```

Run a model
```{r}
model<-fa_pufa%>%
  
  aov(area_normalized ~ parent * temperature + lipid, data=.)

summary(model)
```

Total all PUFA for each sample. 
```{r}
fa_pufa<-fa_pufa%>%
  group_by(sample, temperature, parent)%>%
  summarise(total_pufa=sum(area_normalized))
```

```{r}
fa_plot1<-fa_pufa%>%
  ggplot(aes(x = temperature, y = total_pufa, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");fa_plot1
```

Run a model
```{r}
model<-fa_pufa%>%
  
  aov(total_pufa ~ parent * temperature, data=.)

summary(model)
```

Plot the VIP fatty acid of 28:7. 

```{r}
fa_pufa<-lipids_filtered %>%
  mutate(class = str_extract(lipid, "^[A-Za-z]+"))%>%
  filter(ontology=="FA")

fa_pufa<-fa_pufa[grepl(":([2-9]|[1-9][0-9])$", fa_pufa$lipid), ]
```

```{r}
fa_plot2<-fa_pufa%>%
  filter(lipid=="FA 28:7")%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"), labels=parent_names)+
  labs(x = "", y = "FFA 28:7 concentration") +
  ylim(0,0.028)+
  ggtitle("F")+
  geom_text(inherit.aes=FALSE, x=2, y=0.025, label="p<0.001", size=4)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.title=element_blank(),
        legend.position="right");fa_plot2

ggsave(fa_plot2, filename="figures/lipids/FA28_7_plot.png", width=5, height=5)
```

Run a model
```{r}
model<-fa_pufa%>%
  filter(lipid=="FA 28:7")%>%
  
  aov(area_normalized ~ parent * temperature, data=.)

summary(model)
```

Parent and temperature are significant. 

                   Df    Sum Sq   Mean Sq F value   Pr(>F)    
parent              2 0.0008263 0.0004132  66.378 9.93e-14 ***
temperature         2 0.0001176 0.0000588   9.450 0.000409 ***
parent:temperature  4 0.0000654 0.0000163   2.626 0.047927 *  
Residuals          42 0.0002614 0.0000062  

Also plot the top VIP: ST 27:2;O
```{r}
st_plot<-lipids_filtered%>%
  filter(lipid=="ST 27:2;O")%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"), labels=parent_names)+
  labs(x = "", y = "ST 27:2;O concentration") +
  ylim(0,0.0015)+
  ggtitle("E")+
  geom_text(inherit.aes=FALSE, x=2, y=0.0012, label="p<0.001", size=4)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.title=element_blank(),
        legend.position="right");st_plot

ggsave(st_plot, filename="figures/lipids/ST27_2_ox_plot.png", width=5, height=5)
```

Run a model
```{r}
model<-lipids_filtered%>%
  filter(lipid=="ST 27:2;O")%>%
  
  aov(area_normalized ~ parent * temperature, data=.)

summary(model)
```

FAs - monounsaturated (":1")
```{r}
fa_mufa<-lipids_filtered %>%
  mutate(class = str_extract(lipid, "^[A-Za-z]+"))%>%
  filter(ontology=="FA")

fa_mufa<-fa_mufa[grepl(":([1])$", fa_mufa$lipid), ]
```

```{r}
fa_plot3<-fa_mufa%>%
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  geom_boxplot()+
  facet_wrap(~lipid)+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");fa_plot3
```

Run a model
```{r}
model<-fa_mufa%>%
  
  aov(area_normalized ~ parent * temperature, data=.)

summary(model)
```

FAs - saturated (":0")

```{r}
fa_sat<-lipids_filtered %>%
  mutate(class = str_extract(lipid, "^[A-Za-z]+"))%>%
  filter(ontology=="FA")

fa_sat<-fa_sat[grepl(":([0])$", fa_sat$lipid), ]
```

```{r}
fa_plot4<-fa_sat%>%
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  geom_boxplot()+
  facet_wrap(~lipid, scales="free")+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");fa_plot4
```

Run a model
```{r}
model<-fa_sat%>%
  
  aov(area_normalized ~ parent * temperature + lipid, data=.)

summary(model)
```

Total all SFA for each sample. 
```{r}
fa_sat<-fa_sat%>%
  group_by(sample, temperature, parent)%>%
  summarise(total_sfa=sum(area_normalized))
```

```{r}
fa_plot1<-fa_sat%>%
  ggplot(aes(x = temperature, y = total_sfa, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");fa_plot1
```

Run a model
```{r}
model<-fa_sat%>%
  
  aov(total_sfa ~ parent * temperature, data=.)

summary(model)
```

## PUFA:SFA ratio

PUFA = 2 or more double bonds
SFA = 0 double bonds 

Calculate the sum percent composition of all PUFAs and all SFAs.  

```{r}
pufa_sfa <- lipids_filtered %>%
  filter(ontology=="FA")%>%
  
  # Extract the number after ":" and convert to numeric
  mutate(double_bonds = as.numeric(str_extract(lipid, "(?<=:)\\d+"))) %>%
  # Identify PUFA and SFA
  mutate(
    is_pufa = double_bonds >= 2,
    is_sfa = double_bonds == 0
  ) %>%
  # Group and summarize both PUFA and SFA
  group_by(sample, parent, temperature) %>%
  summarize(
    pufa = sum(area_normalized[is_pufa], na.rm = TRUE),
    sfa = sum(area_normalized[is_sfa], na.rm = TRUE)
  ) %>%
  ungroup()%>%
  mutate(ratio = pufa / sfa); pufa_sfa
```

Plot PUFA:SFA ratio. 

```{r}
pufa_plot<-pufa_sfa%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  group_by(sample, parent, temperature)%>%
  summarise(mean=mean(ratio, na.rm=TRUE))%>%
  
  ggplot(aes(x = temperature, y = mean, colour = parent)) +
  geom_point() +
  geom_boxplot() + 
  labs(x = "Temperature", y = "PUFA:SFA Ratio", title = "PUFA : SFA") +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");pufa_plot
```

Run a model
```{r}
model<-pufa_sfa%>%
  
  aov(ratio ~ parent * temperature, data=.)

summary(model)
```

## EPA (20:5):DHA (22:6)

Create a new data frame with EPA and DHA and calculate the ratio

```{r}
epa_dha<-lipids_filtered %>%
  filter(ontology=="FA")%>%
  filter(lipid %in% c("FA 20:5", "FA 22:6"))
```

Calculate the ratio. 

```{r}
epa_dha<-epa_dha%>%
  ungroup()%>%
  select(!formula)%>%
  pivot_wider(names_from = lipid, values_from = area_normalized)%>%
  mutate(ratio=`FA 20:5`/`FA 22:6`)
```

Plot EPA:DHA ratio. 

```{r}
ratio_plot1<-epa_dha%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  group_by(sample, parent, temperature)%>%
  summarise(mean=mean(ratio, na.rm=TRUE))%>%
  
  ggplot(aes(x = temperature, y = mean, colour = parent)) +
  geom_point(position=position_dodge(0.75)) +
  geom_boxplot() + 
  ggtitle("C")+
  labs(x = "Temperature", y = "EPA:DHA Ratio") +
  geom_text(inherit.aes=FALSE, x=2, y=3, label="p<0.001", size=4)+
  scale_colour_manual(values=c("orange", "darkred", "darkgray"), labels=parent_names)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right", 
        legend.title=element_blank());ratio_plot1

ggsave(ratio_plot1, filename="figures/lipids/epa_dha_ratio.png", width=5, height=5)
```

Run a model
```{r}
model<-epa_dha%>%
  
  aov(ratio ~ parent * temperature, data=.)

anova(model)
```

Significant effect of parent on EPA:DHA ratio with higher in bleached.  

                   Df  Sum Sq Mean Sq F value    Pr(>F)    
parent              2  9.2779  4.6389 12.7643 4.666e-05 ***
temperature         2  0.8674  0.4337  1.1933    0.3133    
parent:temperature  4  1.1426  0.2856  0.7860    0.5408    
Residuals          42 15.2640  0.3634 

```{r}
library(emmeans)

emm<-emmeans(model, ~parent | temperature)
pairs(emm)

emm<-emmeans(model, ~parent)
pairs(emm)
```

temperature = Ambient:
 contrast               estimate    SE df t.ratio p.value
 Bleached - Nonbleached    0.466 0.348 42   1.338  0.3823
 Bleached - Wildtype       0.701 0.348 42   2.014  0.1214
 Nonbleached - Wildtype    0.235 0.348 42   0.675  0.7790

temperature = High (+6):
 contrast               estimate    SE df t.ratio p.value
 Bleached - Nonbleached    0.887 0.348 42   2.549  0.0379
 Bleached - Wildtype       1.478 0.365 42   4.048  0.0006
 Nonbleached - Wildtype    0.591 0.365 42   1.618  0.2494

temperature = Moderate (+3):
 contrast               estimate    SE df t.ratio p.value
 Bleached - Nonbleached    0.305 0.365 42   0.837  0.6826
 Bleached - Wildtype       0.982 0.365 42   2.690  0.0270
 Nonbleached - Wildtype    0.677 0.381 42   1.774  0.1905

At high, ratio is higher in bleached than the other two groups. At moderate, bleached is higher than wildtype. No difference at ambient. 

 contrast               estimate    SE df t.ratio p.value
 Bleached - Nonbleached    0.553 0.204 42   2.706  0.0260
 Bleached - Wildtype       1.054 0.208 42   5.076  <.0001
 Nonbleached - Wildtype    0.501 0.211 42   2.376  0.0564

## Oxidized fatty acids 

Oxidized fatty acids 
```{r}
fa_ox<-lipids_filtered %>%
  mutate(class = str_extract(lipid, "^[A-Za-z]+"))%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  filter(ontology=="OxFA")
```

```{r}
xfa_plot1<-fa_ox%>%
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");xfa_plot1
```

Run a model
```{r}
model<-fa_ox%>%
  
  aov(area_normalized ~ parent * temperature * lipid, data=.)

summary(model)

emm<-emmeans(model, ~parent | temperature)
pairs(emm)
```

No clear effects. 

## Plot composition of TGs (ontology=TG) 

TG's are categorized by saturated, monounsaturated, polyunsaturated TGs and by the saturation of their FA.  

- Saturated FA in TG (0 double bonds)
- Monounsaturated FA in TG (1 double bond) 
- Polyunsaturated FA in TG (>1 double bond)

Classify TAGs. 
```{r}
# Function to classify
classify_TAG <- function(lipid_name) {
  # Extract the total double bonds from the first part (e.g., "TG 54:10")
  db <- as.numeric(str_match(lipid_name, "TG\\s*\\d+:(\\d+)")[,2])
  
  if (db == 0) {
    return("Saturated")
  } else if (db == 1) {
    return("Monounsaturated")
  } else {
    return("Polyunsaturated")
  }
}

# Apply function
tags <- lipids_filtered %>%
  filter(ontology=="TG")%>%
  mutate(classification = sapply(lipid, classify_TAG))%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))
```

Plot all TAGs that are in our VIP list. 
```{r}
tag_plot1a<-tags%>%
  filter(lipid %in% temperature_vip_list)%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");tag_plot1a
```

Plot total Monounsaturated TAGs

Monounsaturated TAGs 
```{r}
tag_plot1<-tags%>%
  filter(classification=="Monounsaturated")%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Monounsaturated TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");tag_plot1

tag_plot1b<-tags%>%
  filter(classification=="Monounsaturated")%>%
  group_by(sample, parent, temperature)%>%
  summarise(mono_tag=sum(area_normalized))%>%
  
  ggplot(aes(x = temperature, y = mono_tag, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Monounsaturated TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");tag_plot1b

#run model 
model<-tags%>%
  filter(classification=="Monounsaturated")%>%
  group_by(sample, parent, temperature)%>%
  summarise(sat_tag=sum(area_normalized))%>%
  
  aov(sat_tag ~ parent * temperature, data=.)

summary(model)
```

No effects on monounsaturated TAGs

Plot total Polyunsaturated TAGs

Polyunsaturated TAGs (none in VIP list, view totals only)
```{r}
tag_plot1b<-tags%>%
  filter(classification=="Polyunsaturated")%>%
  group_by(sample, parent, temperature)%>%
  summarise(poly_tag=sum(area_normalized))%>%
  
  ggplot(aes(x = temperature, y = poly_tag, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Polyunsaturated TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");tag_plot1b

#run model 
model<-tags%>%
  filter(classification=="Polyunsaturated")%>%
  group_by(sample, parent, temperature)%>%
  summarise(poly_tag=sum(area_normalized))%>%
  
  aov(poly_tag ~ parent * temperature, data=.)

summary(model)
```

No effects on polyunsaturated TAGs

## Plot oxidized TAGs 

Plot oxidized TAGs in our VIP list and total concentration of oxidized TAGS.

Plot all TAGs that are oxidized in our data set and total concentration. 
```{r}
#plot all 
ox_tags <- lipids_filtered %>%
  filter(ontology=="OxTG")%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))

oxtag_plot1<-ox_tags%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Oxidized TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");oxtag_plot1

oxtag_plot2<-ox_tags%>%
  group_by(sample, parent, temperature)%>%
  summarise(ox_tag=sum(area_normalized))%>%
  
  ggplot(aes(x = temperature, y = ox_tag, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Oxidized TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");oxtag_plot2

#run model 
model<-ox_tags%>%
  group_by(sample, parent, temperature)%>%
  summarise(ox_tag=sum(area_normalized))%>%
  
  aov(ox_tag ~ parent * temperature, data=.)

summary(model)
```

No effects. 

Plot only oxidized TAGs on our VIP list. 
```{r}
#plot all 
ox_tags <- lipids_filtered %>%
  filter(ontology=="OxTG")%>%
  filter(lipid %in% temperature_vip_list)%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))

oxtag_plot2<-ox_tags%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Oxidized TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");oxtag_plot2
```

## Plot ether TAGs

Plot all ether TAGs that are oxidized in our data set and total concentration. 
```{r}
#plot all 
ether_tags <- lipids_filtered %>%
  filter(ontology=="EtherTG")%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))

etag_plot1<-ether_tags%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Ether TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");oxtag_plot1

etag_plot2<-ether_tags%>%
  group_by(sample, parent, temperature)%>%
  summarise(e_tag=sum(area_normalized))%>%
  
  ggplot(aes(x = temperature, y = e_tag, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Ether TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");etag_plot2

#run model 
model<-ether_tags%>%
  group_by(sample, parent, temperature)%>%
  summarise(e_tag=sum(area_normalized))%>%
  
  aov(e_tag ~ parent * temperature, data=.)

summary(model)
```

No effects. 

Plot only ether TAGs on our VIP list. 
```{r}
#plot all 
ether_tags <- lipids_filtered %>%
  filter(ontology=="EtherTG")%>%
  filter(lipid %in% temperature_vip_list)%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))

etag_plot2<-ether_tags%>%
  
  ggplot(aes(x = temperature, y = area_normalized, colour = parent)) +
  facet_wrap(~lipid, scales="free")+
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "Ether TAG concentration") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");etag_plot2
```


## TAG:FA Ratios 

Calculate FFA:TAG ratio of percent composition of total lipid pool. A higher ratio indicates relative increases in FFA compared to TAGs. 

```{r}
ffa_tag<-lipids_filtered%>%
  ungroup()%>%
  select(!formula)%>%
  filter(ontology %in% c("FA", "TG"))%>%
  group_by(sample, parent, temperature, ontology)%>%
  summarise(total=sum(area_normalized))%>%
  pivot_wider(names_from=ontology, values_from=total)%>%
  mutate(ratio=FA/TG);ffa_tag
```

Plot FA:TAG ratio. 

```{r}
ratio_plot2<-ffa_tag%>%
  
  ggplot(aes(x = temperature, y = ratio, colour = parent)) +
  geom_boxplot()+
  geom_point(position=position_dodge(0.75)) +
  scale_colour_manual(values=c("orange", "darkred", "darkgray"))+
  labs(x = "", y = "FFA:TAG ratio") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position="right");ratio_plot2
```

Run a model
```{r}
model<-ffa_tag%>%
  
  aov(ratio ~ parent * temperature, data=.)

anova(model)
```

No effects 
