---
title: "Analyzing larval environmental data - Hawaii 2023"
author: "Ariana S Huffmyer"
date: "2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from Hobo tidbit loggers and daily measurements.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Tidbit analysis 

## Read in files  

Read in Hobo Tidbit files  
```{r}
tidbit.files <- list.files(path="data/environmental/tidbit_loggers", pattern="*.csv", full.names=TRUE)

#will need to change to HST time 
tidbits <- setNames(tidbit.files, tidbit.files) %>%
    map_dfr(read_csv, .id = "logger") %>%
    rename(DateTime=`Date-Time (HST)`, temp.C=`Ch: 1 - Temperature   (°C)`)%>%
    select(!`#`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

tidbits$DateTime<-as.POSIXct(tidbits$DateTime, format="%m/%d/%Y %H:%M:%S", tz="Pacific/Honolulu")
```

Remove time periods during logger read out.   
```{r}
#tidbits<-tidbits%>%filter(DateTime<as.POSIXct('2022-10-28 16:30:00', tz="Pacific/Honolulu"))
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(tidbits$logger))
```

There are 6 loggers as expected.  

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata<-read_csv("data/environmental/logger_metadata.csv")
metadata$logger<-as.character(metadata$logger)
```

Assign treatment information to logger data.  
```{r}
tidbits<-left_join(tidbits, metadata, by="logger")
tidbits$tank<-as.factor(tidbits$tank)
tidbits$group<-as.factor(tidbits$group)
tidbits$project<-as.factor(tidbits$project)
tidbits$type<-as.factor(tidbits$type)
```

# Developmental time series project 

Plot temperature by incubator colored by treatment. 

```{r}
temp_plot1<-tidbits%>%
  filter(project=="dev_timeseries")%>%
  
  ggplot(aes(x=DateTime, y=temp.C, colour=group))+
  geom_point()+
  #ylim(26, 30)+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  theme_classic(); temp_plot1
```

Summarize by treatment with error shading for standard error.  
```{r}
temp_plot_treatment1<-tidbits%>%
  filter(project=="dev_timeseries")%>%
  group_by(group, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE), sd=sd(temp.C, na.rm=TRUE), sem=sd/sqrt(6), upper=mean+sd, lower=mean-sd)%>%
  
  ggplot(aes(x=DateTime, y=mean, colour=group, fill=group))+
  #geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper, colour=group), alpha = 0.4, colour = NA, show.legend=FALSE) +
  geom_smooth(method="loess", span=0.1, se=FALSE, aes(colour=group), alpha=1, show.legend=TRUE) +
  #ylim(26, 30)+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  scale_colour_manual(values=c("blue", "red"), name="Temperature")+
  scale_fill_manual(values=c("blue", "red"), name="Temperature")+
  theme_classic(); temp_plot_treatment1

#ggsave("figures/environmental/tidbit_temperature_smooth.png", temp_plot_treatment, width=8, height=6)
```

Calculate the mean difference in temperature treatments between groups.  
```{r}
tidbits%>%
  filter(project=="dev_timeseries")%>%
  group_by(group, DateTime)%>%
  summarise(mean=mean(temp.C, na.rm=TRUE))%>%
  spread(key=group, value=mean)%>%
  reframe(difference=high-ambient)%>%
  reframe(mean=mean(difference))
```

Calculate mean temperature for each symbiont type.     
```{r}
tidbits %>%
  filter(project=="dev_timeseries")%>%
  mutate(day = floor_date(DateTime, "day")) %>%
           group_by(day, tank, group) %>%
           summarise(mean = mean(temp.C))%>%
          group_by(group) %>%
           summarise(mean=mean(mean))
```





# Black logger analysis (NEED TO ADD)

Add spots for black loggers for metadata for each tank and treatment 
Can do after Hawaii, just look over treatment information and do frequent measurements in the day of 
