---
title: "Metabolomics analysis"
author: "Ariana S Huffmyer"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(broom)
library(cowplot)
library(mixOmics)
library(dplyr)
library(conflicted)
library(vegan)
library(factoextra)
library(ggfortify)
library(RVAideMemoire)
library(ComplexHeatmap)
library(circlize)
library(tibble)
library(viridis)
library(genefilter)
library(UpSetR)
library(emmeans)
library(car)

conflict_prefer("select", winner = "dplyr", quiet = FALSE)
conflict_prefer("filter", winner = "dplyr", quiet = FALSE)
conflict_prefer("rename", winner = "dplyr", quiet = FALSE)
```

# Data manipulation and preparations  

## Read in all files 

```{r}
metadata<-read_csv("data/lipids_metabolites/metabolomics/Metabolite_Metadata_Huffmyer_Matthews.csv")%>%rename(tube=`Sample ID`, type=Type, temperature=Temperature, parent=Symbiont)

metabolites<-read_xlsx("data/lipids_metabolites/metabolomics/metabolite_data.xlsx")

protein<-read_csv("data/lipids_metabolites/protein/protein_calculated.csv")
```

Add protein data to the metadata. 

```{r}
metadata$protein.ug<-protein$`Total Protein ug`[match(metadata$tube, protein$Sample)]
```

## Format data and conduct protein normalization 

Format the metabolite data frame in long format to obtain sample names and normalize to total protein. 

```{r}
metabolites<-metabolites%>%
  pivot_longer(names_to = "file", values_to = "area", cols=c(2:74))%>%
  mutate(tube = str_extract(file, "(?<=_)[^_]+(?=_)"))%>%
  mutate(tube = str_remove_all(tube, "M"))

levels(as.factor(metabolites$tube))
levels(as.factor(metadata$tube))
```

Merge in metadata. 
```{r}
metadata<-metadata%>%
  select(tube, type, temperature, parent, protein.ug)

metabolites<-left_join(metabolites, metadata)%>%select(!file)
```

Replace NA's with 0's in the metabolite data frame. 

```{r}
metabolites %>%
  filter(is.na(area)) %>%
  nrow()
# there are many NA's, which need to be replaced with 0

metabolites <- metabolites %>%
  mutate(area = replace_na(area, 0))

metabolites %>%
  filter(is.na(area)) %>%
  nrow()
```

Filter out PBQC samples and blank samples, these were used for peak identification and are not needed now. 
```{r}
metabolites<-metabolites%>%
  filter(!type=="PBQC")%>%
  filter(!type=="Blank")
```

Now perform internal standard normalization to C13 sorbitol average for each sample corrections. 

```{r}
#make a dataframe of sorbitol values for each sample
sorbitol<-metabolites%>%
  filter(Name=="13C6-Sorbitol")%>%
  group_by(tube)%>%
  summarise(area_sorbitol=mean(area, na.rm=TRUE))

#add this value back into the master data frame for corrections. 
metabolites<-metabolites%>%
  left_join(.,sorbitol)
```

Normalize area by dividing area by sorbitol internal standard  value for each sample's peak value for each metabolite. 
```{r}
metabolites<-metabolites%>%
  mutate(area_corrected=area/area_sorbitol)
```

Filter to only keep metabolites that have a detectable value in every sample. 

Check for outliers at this point. 

```{r}
wide_data<-metabolites%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area)%>%
  pivot_wider(names_from=Name, values_from=area_corrected)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
#remove any column with all 0's 
wide_data <- wide_data[, colSums(wide_data != 0) > 0]

wide_data<-wide_data%>%
  select(!`13C6-Sorbitol`)

scaled.pca<-prcomp(wide_data%>%select(where(is.numeric)), scale=TRUE, center=TRUE) 
```

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = tube), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

No outliers before removing metabolites for presence/absence filtering. 

Keep only metabolites detected in all samples. 
```{r}
# Find metabolites that have a value > 0 for every tube
filtered_data <- metabolites %>%
  group_by(Name) %>%
  filter(all(area_corrected > 0))

# Find metabolites that were removed (i.e., have any value <= 0)
removed_metabolites <- metabolites %>%
  filter(!(Name %in% filtered_data$Name)) %>%
  distinct(Name)

kept_metabolites <- metabolites %>%
  filter((Name %in% filtered_data$Name)) %>%
  distinct(Name)

# do this instead by keeping those present at >0 value in at least 30% of samples (allowing for metabolites to remain if they are present in only one parent or temp group)
threshold <- 0.9  # 30% threshold

filtered_data2 <- metabolites %>%
  group_by(Name) %>%
  filter(mean(area_corrected > 0) >= threshold)

# Find metabolites that were removed (i.e., have any value <= 0)
removed_metabolites2 <- metabolites %>%
  filter(!(Name %in% filtered_data2$Name)) %>%
  distinct(Name)

kept_metabolites2 <- metabolites %>%
  filter((Name %in% filtered_data2$Name)) %>%
  distinct(Name) 

# Find the values unique to each list
unique_to_kept1 <- setdiff(kept_metabolites, kept_metabolites2) # Elements in kept1 but not in kept2
unique_to_kept2 <- setdiff(kept_metabolites2, kept_metabolites) # Elements in kept2 but not in kept1

# Combine the unique values
unique_values<- c(unique_to_kept1, unique_to_kept2)
unique_values

length(unique(filtered_data2$Name))

filtered_data<-filtered_data2
```

66 metabolites are kept after 90% of samples >0 filtering, many were removed. This makes sense, the data frame kept all metabolites in the library so we expect far fewer to be retained.   

Normalize corrected area to sample protein. This will generate units of normalized peak area / ug protein. 

```{r}
filtered_data<-filtered_data%>%
  mutate(area_normalized=area_corrected/protein.ug)
```

```{r}
hist(filtered_data$area_normalized)

quantile(filtered_data$area_normalized)
```

```{r}
length(unique(filtered_data$Name))
```

Plot to detect outliers. 

Convert to wide format. 

```{r}
wide_data<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data %>%
                      select(where(is.numeric)), 
                      scale = TRUE, 
                      center = TRUE)
```

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = tube), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

Sample L107 is a clear outlier. Remove. 

```{r}
filtered_data<-filtered_data%>%
  filter(!tube=="L107")

wide_data<-wide_data%>%
  filter(!tube=="L107")
```

Plot PCA again 

```{r}
scaled.pca<-prcomp(wide_data %>%
                      select(where(is.numeric)), 
                      scale = TRUE, 
                      center = TRUE)
```

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = tube), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

# Conduct filtering for relative standard deviation 

First make a heatmap of all samples and metabolites to view variability.  
```{r}
# Convert counts to Z-scores
test_heatmap <- filtered_data %>% 
  mutate(code=paste(temperature, parent))%>%
  #group_by(Name) %>% 
  #mutate(z_score = scale(area_normalized)) %>% 
  group_by(code, Name)%>%
  dplyr::summarise(RSD=(sd(area_normalized)*100)/mean(area_normalized))%>%
  ungroup()

# Pivot data to have metabolites as rows and sample as columns
heatmap_data <- test_heatmap %>% 
  dplyr::select(Name, code, RSD) %>%
  spread(key = code, value = RSD)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "RSD",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  #row_split=3,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

```

View metabolites that have duplicate TMS forms. 
```{r}
filtered_data%>%
  arrange(Name)%>%
  pull(Name)%>%
  unique()

# Extract core name by removing anything after the last dash and number/parentheses
filtered_data$core_name <- filtered_data$Name %>%
  str_replace_all("-[0-9]+TMS\\([0-9]+\\)", "") %>%  # Remove specific patterns like "-4TMS(1)"
  str_replace_all("-[0-9]+TMS", "") %>%  # Remove remaining patterns like "-2TMS"
  str_trim()  # Trim any whitespace

# Group by core_name to find metabolites with shared core names
shared_core_metabolites <- filtered_data %>%
  select(Name, core_name)%>%
  group_by(core_name) %>%
  unique()%>%
  filter(n() > 1) %>%
  ungroup()%>%
  pull(Name)%>%
  unique()
```

Plot  
```{r}
filtered_data%>%
  filter(Name %in% shared_core_metabolites)%>%
  
  ggplot(aes(x=parent, color=temperature, y=area_normalized))+
  facet_wrap(~Name, scales="free")+
  geom_point()+
  theme_classic()+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))
```

Show RSD values
```{r}
test_data<-test_heatmap%>%
  filter(Name %in% shared_core_metabolites)%>%
  group_by(Name)%>%
  summarise(mean=mean(RSD, na.rm=TRUE))%>%
  arrange(Name)
```

Write the metabolites to remove. 
```{r}
remove_list<-c("Glucose-meto-5TMS(2)", #lower concentration
               "Glycine-2TMS", #lower concentration
               "2-Deoxy-glucose-4TMS(1)", #higher RSD
               "Glucosamine-5TMS(1)", #higher RSD
               "Lyxose-meto-4TMS(1)", #higher RSD
               "Rhamnose-meto-4TMS(1)", #higher RSD
               "Sorbose-meto-5TMS(1)") #higher RSD
```

Filter these out. 
```{r}
filtered_data<-filtered_data%>%
  filter(!Name %in% remove_list)
```

View names again. 
```{r}
unique(filtered_data$Name)
```

Remove any metabolite standard ("13C"). 
```{r}
filtered_data<-filtered_data%>%
  filter(!grepl("13C", Name))
```

View names again. 
```{r}
length(unique(filtered_data$Name))
unique(filtered_data$Name)
```

There are now 62 metabolites.  

Make a dataframe that calculates the standard deviation of the metabolite area within treatment relative to the mean (i.e., how "large" is the SD).  

The formula will be:  

(sd*100)/mean 

Calculate for each metabolite in each treatment group. 
```{r}
test<-filtered_data%>%
  group_by(Name, parent, temperature)%>%
  #group_by(Name, parent)%>%
  dplyr::summarise(RSD=(sd(area_normalized)*100)/mean(area_normalized))

length(unique(test$Name))

hist(test$RSD)
```
62 metabolites before 

Remove metabolites that have RSD < threshold in all treatments (RSD values are high, come back to this).  
```{r}
test <- test %>%
  group_by(Name) %>%
  filter(all(RSD < 200))

length(unique(test$Name))

keep_list<-c(unique(test$Name))

length(unique(keep_list))
```
59 metabolites after 

Filter the filtered dataset to include the selected metabolites after RSD filtering.  
```{r}
filtered_data<-filtered_data%>%
  filter(Name %in% keep_list)
```

Filter out internal standards if present still. 
```{r}
filtered_data<-filtered_data%>%
  filter(!Name=="13C5,15N1-Valine")%>%
  filter(!Name=="13C6-Sorbitol")
```

```{r}
length(unique(filtered_data$Name))
```

59 metabolites kept 

# Unsupervised analysis 

Conduct a PCA and PERMANOVA to examine the effects of parent and temperature on the metabolome of coral larvae.

Convert to wide format. 

```{r}
wide_data<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  select(!core_name)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data %>%
                      select(where(is.numeric)), 
                      scale = TRUE, 
                      center = TRUE)

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(wide_data %>%
                      select(where(is.numeric)))

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = wide_data, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = wide_data, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)
parent              2    88.87 0.02897 0.7693  0.594
temperature         2    89.03 0.02902 0.7706  0.589
parent:temperature  4   348.48 0.11359 1.5082  0.100
Residual           44  2541.62 0.82843              
Total              52  3068.00 1.00000          

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = tube), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

```{r}
pca2<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2

pca2b<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="temperature", loadings=FALSE,  colour="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2b

pca_plots<-plot_grid(pca2, pca2b, ncol=2, nrow=1)

ggsave(pca_plots, filename="figures/metabolites/pca_plots.jpeg", width=12, height=4)
```

Plot heatmap of z scores of metabolites. 
```{r}
# Convert counts to Z-scores
test_heatmap <- filtered_data %>% 
  mutate(code=paste(temperature, parent))%>%
  group_by(Name) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(code, Name)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and sample as columns
heatmap_data <- test_heatmap %>% 
  dplyr::select(Name, code, z_score) %>%
  spread(key = code, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-Score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  #row_split=3,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

# save the heatmap 
pdf("figures/metabolites/treatment_zscores.pdf", width=8, heigh=10)
draw(heatmap)
dev.off()
```

By parent
```{r}
# Convert counts to Z-scores
test_heatmap <- filtered_data %>% 
  #mutate(code=paste(temperature, parent))%>%
  group_by(Name) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(parent, Name)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and sample as columns
heatmap_data <- test_heatmap %>% 
  dplyr::select(Name, parent, z_score) %>%
  spread(key = parent, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-Score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_split=3,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

# save the heatmap 
pdf("figures/metabolites/parent_zscores.pdf", width=8, heigh=10)
draw(heatmap)
dev.off()
```

By temperature
```{r}
# Convert counts to Z-scores
test_heatmap <- filtered_data %>% 
  #mutate(code=paste(temperature, parent))%>%
  group_by(Name) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(temperature, Name)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and sample as columns
heatmap_data <- test_heatmap %>% 
  dplyr::select(Name, temperature, z_score) %>%
  spread(key = temperature, value = z_score)%>%
  select(Name, Ambient, `Moderate (+3)`, `High (+6)`)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-Score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_split=3,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

# save the heatmap 
pdf("figures/metabolites/temperature_zscores.pdf", width=8, heigh=10)
draw(heatmap)
dev.off()
```

# Perform ANOVAs of metabolite concentration between groups 

Perform an ANOVA with FDR corrections for metabolites to detect which metabolites are different between groups. 

First run overall model. 
```{r}
model_data<- filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  select(!core_name)

str(model_data)

# Run ANOVA for each metabolite
metabolite_results <- model_data %>%
  group_by(Name) %>%
  nest() %>%
  mutate(
    model = purrr::map(data, ~ lm(area_normalized ~ temperature * parent, data = .x)),
    anova = purrr::map(model, ~ broom::tidy(anova(.x)))
  ) %>%
  select(-data, -model) %>%
  unnest(anova)%>%
  filter(!term=="Residuals")
```

Correct for multiple comparisons. 
```{r}
metabolite_results <- metabolite_results %>%
  group_by(term) %>%
  mutate(p.adj = p.adjust(p.value, method = "BH")) %>%
  ungroup()

significant_metabolites <- metabolite_results %>%
  filter(p.adj < 0.05)
```

```{r}
significant_metabolites
```

There are no significant metabolites. 

# Plot metabolites that have a trend for differences  

Trend for differences between parents 
```{r}
plot2<-filtered_data%>%
  filter(Name %in% c("Glucose-meto-5TMS(1)", "Urea-2TMS", "Docosahexaenoic acid-TMS", "Inositol-6TMS(2)-myo", "Inositol-6TMS(3)-Scyllo", "Glutaric acid-2TMS", "Palmitic acid-TMS"))%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  
  
  ggplot(aes(x=temperature, y=area_normalized, color=parent))+
  facet_wrap(~Name, scales="free")+
  geom_boxplot(outliers=FALSE)+
  geom_point(position=position_dodge(0.75))+
  theme_classic()+
  scale_color_manual(values=c("orange", "darkred", "darkgray"))+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1));plot2

ggsave(plot2, filename="figures/metabolites/test_metabolites.jpeg", width=12, height=8)
```
