---
title: "Metabolomics analysis"
author: "Ariana S Huffmyer"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(broom)
library(cowplot)
library(mixOmics)
library(dplyr)
library(conflicted)
library(vegan)
library(factoextra)
library(ggfortify)
library(RVAideMemoire)
library(ComplexHeatmap)
library(circlize)
library(tibble)
library(viridis)
library(genefilter)
library(UpSetR)

conflict_prefer("select", winner = "dplyr", quiet = FALSE)
conflict_prefer("filter", winner = "dplyr", quiet = FALSE)
conflict_prefer("rename", winner = "dplyr", quiet = FALSE)
```

# Data manipulation and preparations  

## Read in all files 

```{r}
metadata<-read_csv("data/lipids_metabolites/metabolomics/Metabolite_Metadata_Huffmyer_Matthews.csv")%>%rename(tube=`Sample ID`, type=Type, temperature=Temperature, parent=Symbiont)

metabolites<-read_xlsx("data/lipids_metabolites/metabolomics/metabolite_data.xlsx")

protein<-read_csv("data/lipids_metabolites/protein/protein_calculated.csv")
```

Add protein data to the metadata. 

```{r}
metadata$protein.ug<-protein$`Total Protein ug`[match(metadata$tube, protein$Sample)]
```

## Format data and conduct protein normalization 

Format the metabolite data frame in long format to obtain sample names and normalize to total protein. 

```{r}
metabolites<-metabolites%>%
  pivot_longer(names_to = "file", values_to = "area", cols=c(2:74))%>%
  mutate(tube = str_extract(file, "(?<=_)[^_]+(?=_)"))%>%
  mutate(tube = str_remove_all(tube, "M"))

levels(as.factor(metabolites$tube))
levels(as.factor(metadata$tube))
```

Merge in metadata. 
```{r}
metadata<-metadata%>%
  select(tube, type, temperature, parent, protein.ug)

metabolites<-left_join(metabolites, metadata)%>%select(!file)
```

Replace NA's with 0's in the metabolite data frame. 

```{r}
metabolites %>%
  filter(is.na(area)) %>%
  nrow()
# there are many NA's, which need to be replaced with 0

metabolites <- metabolites %>%
  mutate(area = replace_na(area, 0))

metabolites %>%
  filter(is.na(area)) %>%
  nrow()
```

Filter out PBQC samples and blank samples, these were used for peak identification and are not needed now. 
```{r}
metabolites<-metabolites%>%
  filter(!type=="PBQC")%>%
  filter(!type=="Blank")
```

Now perform internal standard normalization to C13 sorbitol average for each sample corrections. 

```{r}
#make a dataframe of sorbitol values for each sample
sorbitol<-metabolites%>%
  filter(Name=="13C6-Sorbitol")%>%
  group_by(tube)%>%
  summarise(area_sorbitol=mean(area, na.rm=TRUE))

#add this value back into the master data frame for corrections. 
metabolites<-metabolites%>%
  left_join(.,sorbitol)
```

Normalize area by dividing area by sorbitol internal standard  value for each sample's peak value for each metabolite. 
```{r}
metabolites<-metabolites%>%
  mutate(area_corrected=area/area_sorbitol)
```

Filter to only keep metabolites that have a detectable value in every sample. 

```{r}
# Find metabolites that have a value > 0 for every tube
filtered_data <- metabolites %>%
  group_by(Name) %>%
  filter(all(area_corrected > 0))

# Find metabolites that were removed (i.e., have any value <= 0)
removed_metabolites <- metabolites %>%
  filter(!(Name %in% filtered_data$Name)) %>%
  distinct(Name)

kept_metabolites <- metabolites %>%
  filter((Name %in% filtered_data$Name)) %>%
  distinct(Name)
```

57 metabolites are kept after this filtering, 569 were removed. This makes sense, the data frame kept all metabolites in the library.  

Normalize corrected area to sample protein. This will generate units of normalized peak area / ug protein. 

```{r}
filtered_data<-filtered_data%>%
  mutate(area_normalized=area_corrected/protein.ug)
```

```{r}
hist(filtered_data$area_normalized)

quantile(filtered_data$area_normalized)
```

```{r}
length(unique(filtered_data$Name))
```

Filter out internal standards if present still. 
```{r}
filtered_data<-filtered_data%>%
  filter(!Name=="13C5,15N1-Valine")%>%
  filter(!Name=="13C6-Sorbitol")
```

We are now ready for anaylsis! 

Plot to detect outliers. 

Convert to wide format. 

```{r}
wide_data<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data[c(4:58)], scale=TRUE, center=TRUE) 
```

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = tube), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

Sample L107 is a clear outlier. Remove. 

```{r}
filtered_data<-filtered_data%>%
  filter(!tube=="L107")

wide_data<-wide_data%>%
  filter(!tube=="L107")
```

# Conduct filtering for relative standard deviation 

Make a dataframe that calculates the standard deviation of the metabolite area within treatment relative to the mean (i.e., how "large" is the SD).  

The formula will be:  

(sd*100)/mean 

Calculate for each metabolite in each treatment group. 
```{r}
test<-filtered_data%>%
  group_by(Name, parent, temperature)%>%
  summarise(RSD=(sd(area_normalized)*100)/mean(area_normalized))

length(unique(test$Name))

hist(test$RSD)
```
55 metabolites before 

Remove metabolites that have RSD < 200 in all treatments (RSD values are high, come back to this).  
```{r}
test <- test %>%
  group_by(Name) %>%
  filter(all(RSD < 200))

length(unique(test$Name))

keep_list<-c(unique(test$Name))
```
52 metabolites after 

Filter the filtered dataset to include the selected metabolites after RSD filtering.  
```{r}
filtered_data<-filtered_data%>%
  filter(Name %in% keep_list)
```

# Unsupervised analysis 

Conduct a PCA and PERMANOVA to examine the effects of parent and temperature on the metabolome of coral larvae.

Convert to wide format. 

```{r}
wide_data<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
wide_data%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
wide_data$temperature<-factor(wide_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

wide_data$parent<-factor(wide_data$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(wide_data[c(4:55)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(wide_data[c(4:55)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = wide_data, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = wide_data, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)  
parent              2    82.93 0.03067 0.8205  0.551  
temperature         2    77.96 0.02883 0.7713  0.592  
parent:temperature  4   319.44 0.11814 1.5802  0.095 .
Residual           44  2223.66 0.82236                
Total              52  2704.00 1.00000      

```{r}
pca1<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  geom_text(aes(x = PC1, y = PC2, label = tube), vjust = -0.5)+
  theme_classic()+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pca1
```

```{r}
pca2<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2

pca2b<-ggplot2::autoplot(scaled.pca, data=wide_data, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca2b

pca_plots<-plot_grid(pca2, pca2b, ncol=2, nrow=1)

ggsave(pca_plots, filename="figures/metabolites/pca_plots.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-wide_data%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca3

ggsave(pca3, filename="figures/metabolites/pca_all_groups.jpeg", width=8, height=6)
```

# Supervised PLS-DA analysis  

Script modified from the following tutorial: https://mixomicsteam.github.io/Bookdown/plsda.html

## PLS-DA for the effect of parent 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

Y <- as.factor(X$parent) #select treatment names
Y

X<-X[4:55] #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=2) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("gray", "orange", "brown4") 

pdf("figures/metabolites/PLSDA_Parent.pdf", width=9, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_parent <- plsda(X,Y, ncomp=2) 

#view plsda model again
plotIndiv(MyResult.plsda_parent, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
parent_VIP <- PLSDA.VIP(MyResult.plsda_parent, graph=TRUE)

parent_VIP_DF <- as.data.frame(parent_VIP[["tab"]])
parent_VIP_DF

# Converting row names to column
parent_VIP_table <- rownames_to_column(parent_VIP_DF, var = "Metabolite")

#filter for VIP > 1
parent_VIP_1 <- parent_VIP_table %>% 
  filter(VIP >= 1)

write_csv(parent_VIP_1, "output/lipids_metabolites/metabolites/Overall_Parent_VIPs.csv")

#plot
parent_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Metabolite,VIP,sum))) +
  geom_point() +
  ylab("Metabolite") +
  xlab("VIP Score") +
  ggtitle("Parent VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


pdf("figures/metabolites/Parent_VIP.pdf", width=4, height=4)
parent_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Metabolite,VIP,sum))) +
  geom_point() +
  ylab("Metabolite") +
  xlab("VIP Score") +
  ggtitle("Parent VIP > 1") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
dev.off() 
```

These lists represent VIPs that drive separation between parent groups.  

Make a list of vips. 
```{r}
parent_vip_list<-parent_VIP_1%>%
  pull(Metabolite)
```
There are 15 VIPs. 

Plot heatmap of z score of lipids across life stages.  

```{r}
# Filter data for VIP metabolites
parent_vip_data <- filtered_data %>% filter(Name %in% parent_vip_list)

# Convert counts to Z-scores
parent_vip_data <- parent_vip_data %>% 
  group_by(Name) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(parent, Name)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- parent_vip_data %>% 
  dplyr::select(Name, parent, z_score) %>%
  spread(key = parent, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  #row_split=3,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/metabolites/VIP_heatmap_parent.pdf", width=5, height=6)
draw(heatmap)
dev.off()
```

View PCA for VIPs

```{r}
vip_data_parent<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  filter(Name %in% parent_vip_list)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
vip_data_parent%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_parent$temperature<-factor(vip_data_parent$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_parent$parent<-factor(vip_data_parent$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_parent[c(4:18)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_parent[c(4:18)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_parent, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_parent, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)
parent              2    44.50 0.05705 1.5591  0.182
temperature         2    22.15 0.02839 0.7760  0.548
parent:temperature  4    85.45 0.10955 1.4969  0.159
Residual           44   627.91 0.80501              
Total              52   780.00 1.00000      


```{r}
pca_vip_parent1<-ggplot2::autoplot(scaled.pca, data=vip_data_parent, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_parent1

pca_vip_parent2<-ggplot2::autoplot(scaled.pca, data=vip_data_parent, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_parent2

pca_plots_parent<-plot_grid(pca_vip_parent1, pca_vip_parent2, ncol=2, nrow=1)

ggsave(pca_plots_parent, filename="figures/metabolites/pca_plots_parentVIP.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-vip_data_parent%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca_parent3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_parent3

ggsave(pca_parent3, filename="figures/metabolites/pca_all_groups_parentVIP.jpeg", width=8, height=6)
```

## PLS-DA for the effect of temperature 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

Y <- as.factor(X$temperature) #select treatment names
Y

levels(Y)

X<-X[4:55] #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=2) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("blue", "orange", "red") 

pdf("figures/metabolites/PLSDA_Temperature.pdf", width=9, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_temperature <- plsda(X,Y, ncomp=2) 

#view plsda model again
plotIndiv(MyResult.plsda_temperature, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the lipids that are most highly differentiating lipids     

```{r}
#extract
temperature_VIP <- PLSDA.VIP(MyResult.plsda_temperature, graph=TRUE)

temperature_VIP_DF <- as.data.frame(temperature_VIP[["tab"]])
temperature_VIP_DF

# Converting row names to column
temperature_VIP_table <- rownames_to_column(temperature_VIP_DF, var = "Metabolite")

#filter for VIP > 1
temperature_VIP_1 <- temperature_VIP_table %>% 
  filter(VIP >= 1)

write_csv(temperature_VIP_1, "output/lipids_metabolites/metabolites/Overall_Temperature_VIPs.csv")

#plot
temperature_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Metabolite,VIP,sum))) +
  geom_point() +
  ylab("Metabolite") +
  xlab("VIP Score") +
  ggtitle("Temperature VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


pdf("figures/metabolites/Temperature_VIP.pdf", width=4, height=4)
temperature_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Metabolite,VIP,sum))) +
  geom_point() +
  ylab("Metabolite") +
  xlab("VIP Score") +
  ggtitle("Temperature VIP > 1") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
dev.off() 
```

These lists represent VIPs that drive separation between temperature groups.  

Make a list of vips. 
```{r}
temperature_vip_list<-temperature_VIP_1%>%
  pull(Metabolite)
```
There are 10 VIPs. 

Plot heatmap of z score of metabolites across temperature.  

```{r}
# Filter data for VIP metabolites
temperature_vip_data <- filtered_data %>% filter(Name %in% temperature_vip_list)

# Convert counts to Z-scores
temperature_vip_data <- temperature_vip_data %>% 
  group_by(Name) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(temperature, Name)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

temperature_vip_data$temperature<-factor(temperature_vip_data$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- temperature_vip_data %>% 
  dplyr::select(Name, temperature, z_score) %>%
  spread(key = temperature, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  #row_split=3,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/metabolites/VIP_heatmap_temperature.pdf", width=6, height=7)
draw(heatmap)
dev.off()
```

View PCA for VIPs

```{r}
vip_data_temperature<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  filter(Name %in% temperature_vip_list)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
vip_data_temperature%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_temperature$temperature<-factor(vip_data_temperature$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_temperature$parent<-factor(vip_data_temperature$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_temperature[c(4:13)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_temperature[c(4:13)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_temperature, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_temperature, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)
parent              2    28.82 0.05543 1.5314  0.208
temperature         2    23.68 0.04553 1.2579  0.263
parent:temperature  4    53.42 0.10273 1.4192  0.203
Residual           44   414.08 0.79630              
Total              52   520.00 1.00000   


```{r}
pca_vip_temperature1<-ggplot2::autoplot(scaled.pca, data=vip_data_temperature, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_temperature1

pca_vip_temperature2<-ggplot2::autoplot(scaled.pca, data=vip_data_temperature, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_temperature2

pca_plots_temperature<-plot_grid(pca_vip_temperature1, pca_vip_temperature2, ncol=2, nrow=1)

ggsave(pca_plots_temperature, filename="figures/metabolites/pca_plots_temperatureVIP.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-vip_data_temperature%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca_parent3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_parent3

ggsave(pca_parent3, filename="figures/metabolites/pca_all_groups_temperatureVIP.jpeg", width=8, height=6)
```

## PLS-DA for the effect of parent x temperature 

Generate a PLS-DA and plot.  
```{r, results=FALSE}
#assigning datasets 
X <- wide_data

levels(as.factor(X$parent))
levels(as.factor(X$temperature))

X$code<-paste(X$parent, X$temperature)

levels(as.factor(X$code))

Y <- as.factor(X$code) #select treatment names
Y

levels(Y)

X<-X[4:55] #pull only data columns

# run PLSDA 
MyResult.plsda <- plsda(X,Y, ncomp=5) # 1 Run the method with ncomp = groups -1
plotIndiv(MyResult.plsda)    # 2 Plot the samples

cols<-c("orange", "orange3", "orange4", "brown", "brown2", "brown4", "gray", "lightgray", "darkgray") 

pdf("figures/metabolites/PLSDA_interaction.pdf", width=9, height=6)
plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
dev.off() 

plotIndiv(MyResult.plsda, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Temperature", ellipse = FALSE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

View classification plot. 

```{r}
perf.plsda <- perf(MyResult.plsda, validation = 'Mfold', folds = 2, 
                  progressBar = FALSE,  # Set to TRUE to track progress
                  nrepeat = 50)         # We suggest nrepeat = 50

plot(perf.plsda, sd = TRUE, legend.position = 'horizontal')
```

Set model to determine VIPs.  

```{r, message=FALSE}
MyResult.plsda_interaction <- plsda(X,Y, ncomp=5) 

#view plsda model again
plotIndiv(MyResult.plsda_interaction, col=cols, ind.names = FALSE, legend=TRUE, legend.title = "Parent x Temperature", ellipse = TRUE, title="", style = "graphics", centroid=FALSE, point.lwd = 2, cex=2)
```

Extract VIP scores 

View the metabolites that are most highly differentiating metabolites     

```{r}
#extract
interaction_VIP <- PLSDA.VIP(MyResult.plsda_interaction, graph=TRUE)

interaction_VIP_DF <- as.data.frame(interaction_VIP[["tab"]])
interaction_VIP_DF

# Converting row names to column
interaction_VIP_table <- rownames_to_column(interaction_VIP_DF, var = "Metabolite")

#filter for VIP > 1
interaction_VIP_1 <- interaction_VIP_table %>% 
  filter(VIP >= 1)

write_csv(interaction_VIP_1, "output/lipids_metabolites/metabolites/Overall_Interaction_VIPs.csv")

#plot
interaction_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Metabolite,VIP,sum))) +
  geom_point() +
  ylab("Metabolite") +
  xlab("VIP Score") +
  ggtitle("Parent x Temperature VIP") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


pdf("figures/metabolites/Interaction_VIP.pdf", width=4, height=4)
interaction_VIP_1 %>%
  arrange(VIP) %>%
  ggplot( aes(x = VIP, y = reorder(Metabolite,VIP,sum))) +
  geom_point() +
  ylab("Metabolite") +
  xlab("VIP Score") +
  ggtitle("Interaction VIP > 1") +
  theme_bw() + theme(panel.border = element_rect(linetype = "solid", color = "black"), panel.grid.major = element_blank(), #Makes background theme white
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
dev.off() 
```

These lists represent VIPs that drive separation between parent x temperature groups.  

Make a list of vips. 
```{r}
interaction_vip_list<-interaction_VIP_1%>%
  pull(Metabolite)
```
There are 13 VIPs. 

Plot heatmap of z score of lipids across groups.   

```{r}
# Filter data for VIP metabolites
interaction_vip_data <- filtered_data %>% filter(Name %in% interaction_vip_list)

# Convert counts to Z-scores
interaction_vip_data <- interaction_vip_data %>% 
  mutate(code=paste(parent, temperature))%>%
  group_by(Name) %>% 
  mutate(z_score = scale(area_normalized)) %>% 
  group_by(code, Name)%>%
  summarise(z_score = mean(z_score, na.rm=TRUE))%>%
  ungroup()

interaction_vip_data$code<-factor(interaction_vip_data$code, levels=c("Bleached Ambient", "Bleached Moderate (+3)", "Bleached High (+6)", "Nonbleached Ambient", "Nonbleached Moderate (+3)", "Nonbleached High (+6)", "Wildtype Ambient", "Wildtype Moderate (+3)", "Wildtype High (+6)"))

# Pivot data to have metabolites as rows and lifestage as columns
heatmap_data <- interaction_vip_data %>% 
  dplyr::select(Name, code, z_score) %>%
  spread(key = code, value = z_score)

# Convert to matrix and set rownames
heatmap_matrix <- as.matrix(column_to_rownames(heatmap_data, var = "Name"))

# Create the heatmap
heatmap <- Heatmap(
  heatmap_matrix,
  name = "Z-score",
  col = inferno(10),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  #row_split=4,
  row_title = "Metabolite", 
  column_names_gp =  gpar(fontsize = 12, border=FALSE),
  column_names_rot = 45,
  row_gap = unit(1, "mm"), 
  border = TRUE,
  row_names_gp = gpar(fontsize = 12, alpha = 0.75, border = FALSE)
)

# Draw the heatmap
draw(heatmap)

dev.off()
pdf("figures/metabolites/VIP_heatmap_parentxtemperature.pdf", width=8, height=8)
draw(heatmap)
dev.off()
```

View PCA for VIPs

```{r}
vip_data_interaction<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  filter(Name %in% interaction_vip_list)%>%
  pivot_wider(names_from=Name, values_from=area_normalized)
```

```{r}
vip_data_interaction%>%
    select_if(~ any(is.na(.)))
```

No columns have NA's.  

```{r}
vip_data_interaction$temperature<-factor(vip_data_interaction$temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)"))

vip_data_interaction$parent<-factor(vip_data_interaction$parent, levels=c("Wildtype", "Bleached", "Nonbleached"))
```

All characteristics are renamed.  

View scree plot. 

```{r}
scaled.pca<-prcomp(vip_data_interaction[c(4:16)], scale=TRUE, center=TRUE) 

fviz_eig(scaled.pca)
```

Prepare a PCA plot
```{r}
# scale data
vegan <- scale(vip_data_interaction[c(4:16)])

# PerMANOVA 
permanova<-adonis2(vegan ~ parent*temperature, data = vip_data_interaction, method='eu')
permanova
```

adonis2(formula = vegan ~ parent * temperature, data = vip_data_interaction, method = "eu")
                   Df SumOfSqs      R2      F Pr(>F)  
parent              2    28.06 0.04151 1.1445  0.332  
temperature         2    27.70 0.04098 1.1299  0.315  
parent:temperature  4    80.85 0.11960 1.6489  0.099 .
Residual           44   539.39 0.79791                
Total              52   676.00 1.00000               

```{r}
pca_vip_interaction1<-ggplot2::autoplot(scaled.pca, data=vip_data_interaction, frame.colour="parent", loadings=FALSE,  colour="parent", shape="temperature", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_fill_manual(name="Parental Phenotype", values=c("Wildtype"="gray", "Bleached"="orange", "Nonbleached"="brown4"))+
  scale_shape_manual(name="Temperature", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_interaction1

pca_vip_interaction2<-ggplot2::autoplot(scaled.pca, data=vip_data_interaction, frame.colour="temperature", loadings=FALSE,  colour="temperature", shape="parent", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_fill_manual(name="Temperature", values=c("blue", "orange", "red"))+
  scale_shape_manual(name="Parental Phenotype", values=c(19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_vip_interaction2

pca_plots_interaction<-plot_grid(pca_vip_interaction1, pca_vip_interaction2, ncol=2, nrow=1)

ggsave(pca_plots_interaction, filename="figures/metabolites/pca_plots_interactionVIP.jpeg", width=12, height=4)
```

View a PCA with treatment code information. 
```{r}
test_wide_data<-vip_data_interaction%>%
  mutate(code=paste(parent, temperature))

levels(as.factor(test_wide_data$code))

pca_interaction3<-ggplot2::autoplot(scaled.pca, data=test_wide_data, frame.colour="code", loadings=FALSE,  colour="code", shape="code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=4) + 
  scale_color_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_fill_manual(name="Group", values=c("orange", "orange", "orange", "brown4", "brown4", "brown4", "gray", "gray", "gray"))+
  scale_shape_manual(name="Group", values=c(19,17,15,19,17,15,19,17,15))+
  theme_classic()+
   theme(legend.text = element_text(size=14), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=14, face="bold"), 
        axis.text = element_text(size=14), 
        axis.title = element_text(size=14,  face="bold"));pca_interaction3

ggsave(pca_interaction3, filename="figures/metabolites/pca_all_groups_interactionVIP.jpeg", width=8, height=6)
```

# Plot mean values of VIPs 

Plot "interaction" or parent x temperature VIPs as a mean. 

```{r}
interaction_means<-filtered_data%>%
  ungroup()%>%
  select(!type)%>%
  select(!protein.ug)%>%
  select(!area_sorbitol)%>%
  select(!area_corrected)%>%
  select(!area)%>%
  filter(Name %in% interaction_vip_list)%>%
  
  group_by(Name, parent, temperature)%>%
  summarise(mean=mean(area_normalized, na.rm=TRUE), se=sd(area_normalized, na.rm=TRUE)/sqrt(length(area_normalized)))%>%
  mutate(temperature=factor(temperature, levels=c("Ambient", "Moderate (+3)", "High (+6)")))%>%
  
  ggplot(aes(x = temperature, y = mean, color = parent)) +
  facet_wrap(~Name, scales="free")+
  geom_line(aes(group=parent), position=position_dodge(0.3))+
  geom_point(position=position_dodge(0.3))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, position=position_dodge(0.3))+
  scale_color_manual(values=c("orange", "brown4", "gray"))+
  labs(x = "",y = "Mean Normalized Area") +
  theme_classic()+
  theme(
    axis.text.x=element_text(angle=45, vjust=1, hjust=1)
  );interaction_means

ggsave(interaction_means, filename="figures/metabolites/interaction_VIP_means.jpeg", width=12, height=12)
```

# Output lists for Metaboanalyst 

List of parental VIPS
```{r}
parent_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/metabolites/parent_VIP_list.csv")
```

List of temperature VIPS
```{r}
temperature_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/metabolites/temperature_VIP_list.csv")
```

List of interaction VIPS
```{r}
interaction_vip_list%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/metabolites/interaction_VIP_list.csv")
```

Full list of lipids
```{r}
full_metabolite_list<-metabolites%>%
  pull(Name)%>%
  unique()%>%
  as.data.frame()%>%
  write_csv(., file="output/lipids_metabolites/metabolites/full_metabolite_list.csv")
```

What is the overlap in these lists? Create an upset plot. 

```{r}
# Combine lists into a data frame indicating membership (1 = present, 0 = absent)
upset_data <- data.frame(
  Name = unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)),
  List1 = as.numeric(unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)) %in% parent_vip_list),
  List2 = as.numeric(unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)) %in% temperature_vip_list),
  List3 = as.numeric(unique(c(parent_vip_list, temperature_vip_list, interaction_vip_list)) %in% interaction_vip_list)
)

# Convert data to the format needed for the UpSetR plot
plot_data <- upset_data[, -1]
row.names(plot_data) <- upset_data$Name
names(plot_data) <- c("Parent", "Temperature", "PxT Group")

# Generate the UpSet plot
pdf("figures/metabolites/upset_VIP_list_overlap.pdf", width=6, height=6)
upset(plot_data, sets = c("Parent", "Temperature", "PxT Group"), keep.order = TRUE)
dev.off()

```

No signature of changes in metabolome by parent or temperature treatment. 

