---
title: "Larval Phenoplate Analysis" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

1. Dark 26°C
- Respiration (O2 Fm)
- QY (relative quantum yield)
- NPQ (non photochemical quenching under dark)

2. Sub saturating Light 26°C 
- Net Photosynthesis (O2 Fm)
- Gross Photosynthesis (Net P + Respiration from Dark 26°C) (O2 Fm)
- NPQ (non photochemical quenching under light)

3. Dark at temperature gradient 
- Respiration (O2 Fm) under temperatures in dark 
- QY (relative quantum yield) under temperatures in dark
- NPQ (non photochemical quenching under dark) under temperatures in dark

Use the last 5 min averaged for QY and NPQ. 
Extract slopes excluding first 5 min from R and P O2 Fm measurements 

O2 - Fm = measured after saturating pulse. Use this metric for O2, it is measured under standardized light pulse consistent light. Lower flourescence = more O2; Higher fluorescence = less O2. 02 quenches the fluorescence. 

Analyze within a protocol - shifts between phases are meaningless. 

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("readxl" %in% rownames(installed.packages()) == 'FALSE') install.packages('readxl')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
library('readxl')
```

# Load data 

Load each data frame to extract O2 Fm, FvFm, and NPQ values. Also extract metadata for light and temperature treatment profiles.  

Turn each data frame to long form, add identifying columns, and join all data together. Prepare each data frame and join together in a running list.  

## C Cold Plate 

Read in data files for C Cold plate 
```{r}
c_cold_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="df_O2_Fm")
c_cold_QY<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="df_QY")
c_cold_NPQ<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="df_NPQ")
c_cold_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
c_cold_O2_fm<-c_cold_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., c_cold_well_metadata)

c_cold_QY<-c_cold_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., c_cold_well_metadata)

c_cold_NPQ<-c_cold_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., c_cold_well_metadata)

c_cold<- c_cold_O2_fm%>%
  left_join(c_cold_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(c_cold_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

c_cold$Plate<-"C-Cold"
```

## C Warm Plate


Read in data files for C Warm plate 
```{r}
c_warm_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="df_O2_Fm")
c_warm_QY<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="df_QY")
c_warm_NPQ<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="df_NPQ")
c_warm_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric.

```{r}
c_warm_O2_fm<-c_warm_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., c_warm_well_metadata)

c_warm_QY<-c_warm_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., c_warm_well_metadata)

c_warm_NPQ<-c_warm_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., c_warm_well_metadata)

c_warm<- c_warm_O2_fm%>%
  left_join(c_warm_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(c_warm_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

c_warm$Plate<-"C-Warm"
```

## MIX Cold Plate

Read in data files for MIX Cold plate 
```{r}
mix_cold_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="df_O2_Fm")
mix_cold_QY<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="df_QY")
mix_cold_NPQ<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="df_NPQ")
mix_cold_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
mix_cold_O2_fm<-mix_cold_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., mix_cold_well_metadata)

mix_cold_QY<-mix_cold_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., mix_cold_well_metadata)

mix_cold_NPQ<-mix_cold_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., mix_cold_well_metadata)

mix_cold<- mix_cold_O2_fm%>%
  left_join(mix_cold_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(mix_cold_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

mix_cold$Plate<-"MIX-Cold"
```

## MIX Warm Plate

Read in data files for MIX Warm plate 
```{r}
mix_warm_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="df_O2_Fm")
mix_warm_QY<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="df_QY")
mix_warm_NPQ<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="df_NPQ")
mix_warm_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
mix_warm_O2_fm<-mix_warm_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., mix_warm_well_metadata)

mix_warm_QY<-mix_warm_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., mix_warm_well_metadata)

mix_warm_NPQ<-mix_warm_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., mix_warm_well_metadata)

mix_warm<- mix_warm_O2_fm%>%
  left_join(mix_warm_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(mix_warm_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

mix_warm$Plate<-"MIX-Warm"
```

## WT Cold Plate

Read in data files for WT Cold plate 
```{r}
wt_cold_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="df_O2_Fm")
wt_cold_QY<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="df_QY")
wt_cold_NPQ<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="df_NPQ")
wt_cold_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
wt_cold_O2_fm<-wt_cold_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., wt_cold_well_metadata)

wt_cold_QY<-wt_cold_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., wt_cold_well_metadata)

wt_cold_NPQ<-wt_cold_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., wt_cold_well_metadata)

wt_cold<- wt_cold_O2_fm%>%
  left_join(wt_cold_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(wt_cold_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

wt_cold$Plate<-"WT-Cold"
```

## WT Warm Plate

Read in data files for WT Warm plate 
```{r}
wt_warm_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="df_O2_Fm")
wt_warm_QY<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="df_QY")
wt_warm_NPQ<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="df_NPQ")
wt_warm_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
wt_warm_O2_fm<-wt_warm_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., wt_warm_well_metadata)

wt_warm_QY<-wt_warm_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., wt_warm_well_metadata)

wt_warm_NPQ<-wt_warm_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., wt_warm_well_metadata)

wt_warm<- wt_warm_O2_fm%>%
  left_join(wt_warm_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(wt_warm_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

wt_warm$Plate<-"WT-Warm"
```

## Join all files together 

Join together. 
```{r}
df<-full_join(c_cold, c_warm)
df<-full_join(df, mix_cold)
df<-full_join(df, mix_warm)
df<-full_join(df, wt_cold)
df<-full_join(df, wt_warm)

df$code<-paste(df$Plate, df$well)
```

Change to factors. 
```{r}
df$Gradient_Temperature<-as.factor(df$Gradient_Temperature)
df$Symbiont<-as.factor(df$Symbiont)
```

# Perform calculations 

## O2 Fm

Normalize each O2 observation to the first measurement at Time 0 for each well. 
```{r}
df<-df%>%
  group_by(Plate, well)%>%
  mutate(O2_Fm_norm=O2_Fm_raw-first(O2_Fm_raw))
```

## QY

QY (relative quantum yield) values do not require any normalization. 

## NPQ

First calculate Fm for each well as average of values from the first 15 min dark period. Then calculate NPQ as (Fm-value)/value. 

```{r}
df <- df %>%
  # Filter rows where Time is within the first 15 minutes
  filter(Time <= 14.99) %>%
  # Group by 'well'
  group_by(Plate, well, code) %>%
  # Calculate the average of 'NPQ_raw' for each 'well'
  summarize(Fm = mean(NPQ_raw, na.rm = TRUE)) %>%
  # Join the calculated averages back to the original data
  left_join(df, by = c("Plate", "well", "code")) %>%
  # Calculate the 'NPQ_norm' column
  mutate(NPQ_norm = (Fm - NPQ_raw) / NPQ_raw)
```

# Visualize data over entire protocol 

Create colour palette for temperatures. 
```{r}
# Define the number of colors in the palette
num_colors <- 12

# Create a color palette from dark blue to dark red
palette <- colorRampPalette(c("darkblue", "lightblue", "red1", "darkred"))(num_colors)
```

## C larvae 

Plot a time series plot of each response over the measurement period. 

Set custom theme for panel plots. 
```{r}
custom_theme<-theme(text=element_text(size=20, face="bold", color="black"), 
        axis.text=element_text(size=18, color="black"), 
        legend.position="none", 
        )
```

Make a light plot to add onto the bottom of each plot for reference. 
```{r}
light_plot<-df%>%
  select(Time, Light)%>%
  group_by(Time, Light)%>%
  summarise(Light=mean(Light, na.rm=TRUE))%>%
  
  ggplot(aes(y=Light, x=Time))+
  geom_line(colour="black", linewidth=2)+
  geom_text(label="Dark (26°C)", x=8, y=250, size=4)+
  geom_text(label="Light (26°C)", x=22, y=600, size=4)+
  geom_text(label="Dark \n[T Gradient]", x=36, y=250, size=4)+
  geom_text(label="Low Light \n[T Gradient]", x=45, y=250, size=4)+
  geom_text(label="Rapid Light Curve \n[T Gradient]", x=58, y=1400, size=4)+
  geom_text(label="Low Light \nRecovery \n(26°C)", x=70, y=400, size=4)+
  theme_classic()+
  custom_theme+
  ylim(0, 1650)+
  xlab("Time (min)")+
  ylab("Light (PAR)");light_plot
```

O2 Fm
```{r}
c_o2_fm_plot<-df%>%
  filter(Symbiont=="C")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(O2_Fm_norm, na.rm=TRUE), se=sd(O2_Fm_norm, na.rm=TRUE)/sqrt(length(O2_Fm_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("Change in O2 fluorescence (Fm)")+
  ggtitle("Cladocopium - Bleached Parents")+
  
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  ylim(-13000, 13000)+
  theme(legend.position="none")+
  custom_theme; c_o2_fm_plot

c_o2_fm_grid<-plot_grid(c_o2_fm_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

QY
```{r}
c_o2_QY_plot<-df%>%
  filter(Symbiont=="C")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(QY, na.rm=TRUE), se=sd(QY, na.rm=TRUE)/sqrt(length(QY)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("QY")+
  ggtitle("Cladocopium - Bleached Parents")+
  ylim(0,1)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none")+
  custom_theme; c_o2_QY_plot

c_o2_QY_grid<-plot_grid(c_o2_QY_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))

```

NPQ
```{r}
c_o2_npq_plot<-df%>%
  filter(Symbiont=="C")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Cladocopium - Bleached Parents")+
  ylim(-0.5,1.8)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none")+
 custom_theme; c_o2_npq_plot

c_o2_npq_grid<-plot_grid(c_o2_npq_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))

```

## MIX larvae 

Plot a time series plot of each response over the measurement period. 

O2 Fm
```{r}
mix_o2_fm_plot<-df%>%
  filter(Symbiont=="MIX")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(O2_Fm_norm, na.rm=TRUE), se=sd(O2_Fm_norm, na.rm=TRUE)/sqrt(length(O2_Fm_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("Change in O2 fluorescence (Fm)")+
  ggtitle("Mixed - Non-Bleached Parents")+
  
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  ylim(-13000, 13000)+
  theme(legend.position="none")+
  custom_theme; mix_o2_fm_plot

mix_o2_fm_grid<-plot_grid(mix_o2_fm_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

QY
```{r}
mix_o2_QY_plot<-df%>%
  filter(Symbiont=="MIX")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(QY, na.rm=TRUE), se=sd(QY, na.rm=TRUE)/sqrt(length(QY)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("QY")+
  ggtitle("Mixed - Non-Bleached Parents")+
  ylim(0,1)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none")+
  custom_theme; mix_o2_QY_plot

mix_o2_QY_grid<-plot_grid(mix_o2_QY_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

NPQ
```{r}
mix_o2_npq_plot<-df%>%
  filter(Symbiont=="MIX")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Mixed - Non-Bleached Parents")+
  ylim(-0.5,1.8)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none")+
  custom_theme; mix_o2_npq_plot

mix_o2_npq_grid<-plot_grid(mix_o2_npq_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

## WT larvae 

Plot a time series plot of each response over the measurement period. 

O2 Fm
```{r}
wt_o2_fm_plot<-df%>%
  filter(Symbiont=="WT")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(O2_Fm_norm, na.rm=TRUE), se=sd(O2_Fm_norm, na.rm=TRUE)/sqrt(length(O2_Fm_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("Change in O2 fluorescence (Fm)")+
  ggtitle("Wildtype")+
  ylim(-13000, 13000)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none")+
  custom_theme; wt_o2_fm_plot

wt_o2_fm_grid<-plot_grid(wt_o2_fm_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

QY
```{r}
wt_o2_QY_plot<-df%>%
  filter(Symbiont=="WT")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(QY, na.rm=TRUE), se=sd(QY, na.rm=TRUE)/sqrt(length(QY)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("QY")+
  ggtitle("Wildtype")+
  ylim(0,1)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  custom_theme; wt_o2_QY_plot

wt_o2_QY_grid<-plot_grid(wt_o2_QY_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

NPQ
```{r}
wt_o2_npq_plot<-df%>%
  filter(Symbiont=="WT")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Wildtype")+
  ylim(-0.5,1.8)+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none")+
  custom_theme; wt_o2_npq_plot

wt_o2_npq_grid<-plot_grid(wt_o2_npq_plot, light_plot, ncol=1, nrow=2, rel_heights=c(1, 0.3))
```

## Make a panel of all symbiont types together 

Grab the legend from the plot. 
```{r}
my_plot <- df%>%
  filter(Symbiont=="WT")%>%
  group_by(Symbiont, Gradient_Temperature, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=1)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Wildtype")+

  scale_colour_manual(values=palette, name="Gradient \nTemperature °C")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature °C")+
  theme_classic()+
  theme(legend.position="right")+
  theme(legend.text=element_text(size=20, face="bold"))+
  theme(legend.title=element_text(size=20, face="bold"))+
  theme(legend.key.size = unit(1.7, 'cm'), #change legend key size
        legend.key.height = unit(1.7, 'cm'), #change legend key height
        legend.key.width = unit(1.7, 'cm'))

# Extract the legend
legend <- get_legend(my_plot)
```

Assemble full time series plots. 
```{r}
fm_grid<-plot_grid(c_o2_fm_plot, mix_o2_fm_plot, wt_o2_fm_plot, 
                    light_plot, light_plot, light_plot, 
                    ncol=3, nrow=2, rel_heights = c(1, 0.7), align="vh", axis = "lr")

fm_grid_l<-plot_grid(fm_grid, legend, ncol=2, nrow=1, rel_widths = c(1, 0.1), align="vh", axis = "lr")


ggsave(filename="figures/phenoplate/timeseries_O2_Fm.jpeg", fm_grid_l, width=35, height=10)
```

```{r}
QY_grid<-plot_grid(c_o2_QY_plot, mix_o2_QY_plot, wt_o2_QY_plot, 
                    light_plot, light_plot, light_plot, 
                    ncol=3, nrow=2, rel_heights = c(1, 0.7), align="vh", axis = "lr")

QY_grid_l<-plot_grid(QY_grid, legend, ncol=2, nrow=1, rel_widths = c(1, 0.1), align="vh", axis = "lr")

ggsave(filename="figures/phenoplate/timeseries_O2_QY.jpeg", QY_grid_l, width=35, height=10)
```

```{r}
npq_grid<-plot_grid(c_o2_npq_plot, mix_o2_npq_plot, wt_o2_npq_plot, 
                    light_plot, light_plot, light_plot, 
                    ncol=3, nrow=2, rel_heights = c(1, 0.7), align="vh", axis = "lr")

npq_grid_l<-plot_grid(npq_grid, legend, ncol=2, nrow=1, rel_widths = c(1, 0.1), align="vh", axis = "lr")

ggsave(filename="figures/phenoplate/timeseries_O2_NPQ.jpeg", npq_grid_l, width=35, height=10)
```

# Phase specific analyses  

Load a dataframe of start and end times for each phase from the treatment metadata. 
```{r}
treatments<-read_csv(file="data/phenoplate_oxygen/treatment_metadata.csv")
```

Add a column for the phase of each observation in our data frame. 

```{r}
get_phase <- function(Time) {
  matching_phase <- treatments$Phase[Time >= treatments$Time.Start & Time <= treatments$Time.End]
  if (length(matching_phase) > 0) {
    return(matching_phase[1])
  } else {
    return(NA)
  }
}

df$Phase <- sapply(df$Time, get_phase)
```

Select only required columns. 

```{r}
df<-df%>%
  select(Plate, well, code, Time, Light, Gradient_Temperature, Symbiont, O2_Fm_norm, NPQ_norm, QY, Phase)
```

## 1. Dark 26°C: Dark respiration, NPQ, and QY ~ Phenotype 

### Respiration - O2 Fm

Plot the data from the first phase (dark respiration at 26°C) faceted by symbiont type. 

Higher value on Y = more oxygen depleted. 
```{r}
dark_26c_plot2<-df%>%
  filter(Phase=="Dark 26C")%>%
  group_by(Symbiont, Phase, Time, Light)%>%
  mutate(inv_02_Fm_norm=-1*O2_Fm_norm)%>%
  summarise(mean=mean(inv_02_Fm_norm, na.rm=TRUE), se=sd(inv_02_Fm_norm, na.rm=TRUE)/sqrt(length(inv_02_Fm_norm)))%>%
  
  ggplot(aes(x=Time, y=(mean), colour=Symbiont, fill=Symbiont))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("Oxygen Concentration [-1*rel Fm]")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(-800, 200)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_plot2
```
There appears to be a higher respiration rate in wildtype and perhaps a lower rate in tolerant compared to sensitive parental phenotypes. 

Calculate respiration rate as the mean total change in relative O2 Fm values divided by the total time (15 min). Use the last 3 minutes as the interval from which to summarize total   change in normalized (normalized to starting value) oxygen concentration.  

Summarize oxygen values in last 3 minutes. Calculate as relative change in Fm per min. 
```{r}
resp_dark26<-df%>%
  filter(Phase=="Dark 26C")%>%
  filter(Time>=13)%>%
  group_by(Plate, well, code, Symbiont, Phase)%>%
  summarise(Rdark26=mean(O2_Fm_norm, na.rm=TRUE)/15)
```

Plot values. 

```{r}
dark_26c_plot3<-resp_dark26%>%
  
  ggplot(aes(x=Symbiont, y=Rdark26, colour=Symbiont))+
  geom_boxplot()+
  geom_point(position=position_jitterdodge())+
  xlab("Symbiont")+
  ylab("O2 per min [-1*norm. Fm]")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  #ylim(-800, 200)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_plot3

dark_26c_plot4<-resp_dark26%>%
  group_by(Symbiont)%>%
  summarise(mean=mean(Rdark26, na.rm=TRUE), se=sd(Rdark26, na.rm=TRUE)/sqrt(length(Rdark26)))%>%
  
  ggplot(aes(x=Symbiont, y=mean, colour=Symbiont, fill=Symbiont))+
  geom_point(size=4)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  xlab("Parental Phenotype")+
  ylab(expression(bold(paste("Oxygen Consumed [-1*rel Fm min" ^-1, "]"))))+
  ggtitle(expression(bold(paste("R" [D], " 26°C"))))+
  geom_text(x=2, y=28, label="p=0.040", size=6, color="black")+
  geom_text(x=3, y=23, label="*", size=6, color="black")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_x_discrete(labels=c("Sensitive", "Tolerant", "Wildtype"))+
  ylim(0, 30)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=14, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        title = element_text(size=12, color="black", face="bold")); dark_26c_plot4

ggsave("figures/phenoplate/respiration_dark26.png", width=4.5, heigh=4.5)
```

Run ANOVA model on these values. 
```{r}
model<-resp_dark26%>%
  
  aov(Rdark26~Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont)
pairs(emm)
```
Slightly significant differences detected (P=0.0403). Lower respiration rates in MIX compared to WT, but no differences between C and other groups (P=0.0334).  

Create a new master data frame to populate and add in dark respiration data. 
```{r}
master<-df%>%
  select(Plate, well, code, Light, Gradient_Temperature, Symbiont, Phase)%>%
  unique()

master<-left_join(master, resp_dark26)
```

### Relative quantum yield - QY

Plot the data from the first phase (dark at 26°C) faceted by symbiont type. 

```{r}
dark_26c_QY_plot2<-df%>%
  filter(Phase=="Dark 26C")%>%
  group_by(Symbiont, Phase, Time, Light)%>%
  summarise(mean=mean(QY, na.rm=TRUE), se=sd(QY, na.rm=TRUE)/sqrt(length(QY)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Symbiont, fill=Symbiont))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("Relative Quantum Yield")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(0.2, 0.75)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_QY_plot2
```
There appears to be lower QY in wildtype larvae. 

Calculate mean QY using the last 5 minutes. 

Summarize oxygen values in last 5 minutes. 
```{r}
QY_dark26<-df%>%
  filter(Phase=="Dark 26C")%>%
  filter(Time>=10)%>%
  group_by(Plate, well, code, Symbiont, Phase)%>%
  summarise(QYdark26=mean(QY, na.rm=TRUE))
```

Plot values. 

```{r}
dark_26c_QY_plot3<-QY_dark26%>%
  
  ggplot(aes(x=Symbiont, y=QYdark26, colour=Symbiont))+
  geom_boxplot()+
  geom_point(position=position_jitterdodge())+
  xlab("Symbiont")+
  ylab("Relative QY")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(0.2, 0.8)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_QY_plot3

dark_26c_QY_plot4<-QY_dark26%>%
  group_by(Symbiont)%>%
  summarise(mean=mean(QYdark26, na.rm=TRUE), se=sd(QYdark26, na.rm=TRUE)/sqrt(length(QYdark26)))%>%
  
  ggplot(aes(x=Symbiont, y=mean, colour=Symbiont, fill=Symbiont))+
  geom_point(size=4)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  xlab("Parental Phenotype")+
  ylab(expression(bold(paste("Relative Quantum Yield"))))+
  ggtitle(expression(bold(paste("QY 26°C"))))+
  geom_text(x=2, y=0.68, label="p<0.001", size=6, color="black")+
  geom_text(x=3, y=0.62, label="*", size=6, color="black")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_x_discrete(labels=c("Sensitive", "Tolerant", "Wildtype"))+
  ylim(0.55, 0.7)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=14, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        title = element_text(size=12, color="black", face="bold")); dark_26c_QY_plot4

ggsave("figures/phenoplate/QY_dark26.png", width=4.5, heigh=4.5)
```

Run ANOVA model on these values. 
```{r}
model<-QY_dark26%>%
  
  aov(QYdark26~Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont)
pairs(emm)
```
Lower QY in wildtype (P<0.001).  

Create a new master data frame to populate and add in dark QY data. 
```{r}
master<-left_join(master, QY_dark26)
```

### Non-photochemical quenching (NPQ)

Plot the data from the first phase (dark at 26°C) faceted by symbiont type. 

```{r}
dark_26c_NPQ_plot2<-df%>%
  filter(Phase=="Dark 26C")%>%
  group_by(Symbiont, Phase, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Symbiont, fill=Symbiont))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  #ylim(0.2, 0.75)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_NPQ_plot2
```
There appears to be lower NPQ in wildtype larvae. 

Calculate mean NPQ using the last 5 minutes. 

Summarize values in last 5 minutes.
```{r}
NPQ_dark26<-df%>%
  filter(Phase=="Dark 26C")%>%
  filter(Time>=10)%>%
  group_by(Plate, well, code, Symbiont, Phase)%>%
  summarise(NPQdark26=mean(NPQ_norm, na.rm=TRUE))
```

Plot values. 

```{r}
dark_26c_NPQ_plot3<-NPQ_dark26%>%
  
  ggplot(aes(x=Symbiont, y=NPQdark26, colour=Symbiont))+
  geom_boxplot()+
  geom_point(position=position_jitterdodge())+
  xlab("Symbiont")+
  ylab("NPQ")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(-0.1, 0.2)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_NPQ_plot3

dark_26c_NPQ_plot4<-NPQ_dark26%>%
  group_by(Symbiont)%>%
  summarise(mean=mean(NPQdark26, na.rm=TRUE), se=sd(NPQdark26, na.rm=TRUE)/sqrt(length(NPQdark26)))%>%
  
  ggplot(aes(x=Symbiont, y=mean, colour=Symbiont, fill=Symbiont))+
  geom_point(size=4)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  xlab("Parental Phenotype")+
  ylab(expression(bold(paste("Non-Photochemical Quenching"))))+
  ggtitle(expression(bold(paste("NPQ 26°C"))))+
  geom_text(x=2, y=0.09, label="p=0.007", size=6, color="black")+
  geom_text(x=3, y=0.051, label="*", size=6, color="black")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_x_discrete(labels=c("Sensitive", "Tolerant", "Wildtype"))+
  ylim(0, 0.1)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=14, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        title = element_text(size=12, color="black", face="bold")); dark_26c_NPQ_plot4

ggsave("figures/phenoplate/NPQ_dark26.png", width=4.5, heigh=4.5)
```

Run ANOVA model on these values. 
```{r}
model<-NPQ_dark26%>%
  
  aov(NPQdark26~Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont)
pairs(emm)
```
Lower QY in wildtype (P=0.007).  

Create a new master data frame to populate and add in dark NPQ data. 
```{r}
master<-left_join(master, NPQ_dark26)

levels(as.factor(master$Phase))

master<-master%>%
  filter(Phase %in% c("Dark 26C", "Light 26C", "Dark - Temp Gradient"))
```












FvFm
```{r}
dark_26c_plot3<-df%>%
  filter(Phase=="Dark 26C")%>%
  group_by(Symbiont, Phase, Time, Light)%>%
  summarise(mean=mean(FvFm, na.rm=TRUE), se=sd(FvFm, na.rm=TRUE)/sqrt(length(FvFm)))%>%
  
  ggplot(aes(x=Time, y=mean, fill=Symbiont, colour=Symbiont))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("FvFm")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(0.5,0.75)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_plot3

model<-df%>%
  filter(Phase=="Dark 26C")%>%
  
  aov(FvFm~Symbiont*Time, data=.)

summary(model)
```
There are symbiont and symbiont x time effects.

NPQ
```{r}
dark_26c_plot4<-df%>%
  filter(Phase=="Dark 26C")%>%
  group_by(Symbiont, Phase, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, fill=Symbiont, colour=Symbiont))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(-0.3,0.3)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); dark_26c_plot4

model<-df%>%
  filter(Phase=="Dark 26C")%>%
  
  aov(NPQ_norm~Symbiont*Time, data=.)

summary(model)
```
There are time and symbiont x time effects. 

Make a legend for symbiont type. 
```{r}
my_plot_sym <- df%>%
  filter(Phase=="Dark 26C")%>%
  group_by(Symbiont, Phase, Time, Light)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, fill=Symbiont, colour=Symbiont))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("Dark 26°C")+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  ylim(-0.3,0.3)+
  theme_classic()+
  theme(text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold"))+
  theme(legend.position="right")+
  theme(legend.text=element_text(size=15, face="bold"))+
  theme(legend.title=element_text(size=15, face="bold"))+
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'))

# Extract the legend
legend_symbiont <- get_legend(my_plot_sym)
```

Plot all together. 
```{r}
dark26_grid<-plot_grid(dark_26c_plot1, dark_26c_plot2, dark_26c_plot3, dark_26c_plot4,
                    ncol=4, nrow=1, align="vh", axis = "lr")

dark26_grid_l<-plot_grid(dark26_grid, legend_symbiont, ncol=2, nrow=1, rel_widths = c(1, 0.1), align="vh", axis = "lr")

ggsave(filename="figures/phenoplate/1_dark_26c_metrics.jpeg", dark26_grid_l, width=20, height=5)
```



## 2. Light 26°C: Net photosynthesis, Gross photosynthesis, NPQ, and QY ~ Phenotype





## 2. Dark [Temperature Gradient]: Light enhanced respiration, NPQ, and QY ~ Temperature and phenotype














