---
title: "Larval Phenoplate Analysis" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---


Look at temperature induced NPQ responses (Suggett paper) + symbiont/bacteria co culture phenoplate paper
- Look at NPQ levels before RLC to see if NPQ is starting at a different point in RLC from previous treatment, could explain why NPQ starts at different points
- Look at NPQ at end of dark t gradient and end of low light t gradient 
- Then calculate relative change in NPQ within the light curve 

Low light t gradient is to keep photosynthesis going so that its ready for light curve, otherwise dark to light would hvae lag time 

Add statistically different points to all figures to visualize 

Differences are temperature specific - difference in parents are dependent by temperatures 

Plot a heatmap for all metrics - put each metri and temp as own column

Higher NPQ = experiencing stress and or ability to protect self? Decline in NPQ at high temp x light could be sign of damage not repairable 

Bimodal temperature responses of Rd and rETR - expected, optimum in middle with high stress at end. NPQ increase at low temps iss insteresting - Andrei will look into this more. High stress at low and high temps, optimum in mid temperatures 

Andrei - environmental plasticity of phenoplate data? Will look into this 

Think about making conceptual model of biological response with phenotype with PHenoplate as performance metric

Dont use O2 data - maybe use for respiration 





1. Fluorescent light curve = RLC 

Subset RLC analyses 
Calculate rETR as QY (effective quantum yield) / PAR 
Plot y as rETR against x as PAR 

Calculate:
- Saturating irradiance
- rETR max 
- Alpha 
- Beta

2. Oxygen light curve (PI curve)

Plot y as O2 Fm against x as PAR 
Calculate: 
- Pmax
- Saturating irradiance
- Alpha 





# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("readxl" %in% rownames(installed.packages()) == 'FALSE') install.packages('readxl')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
library('readxl')
library('multcomp')
library('dplyr')
```

# Load data 

Load each data frame to extract O2 Fm, FvFm, and NPQ values. Also extract metadata for light and temperature treatment profiles.  

Turn each data frame to long form, add identifying columns, and join all data together. Prepare each data frame and join together in a running list.  

## C Cold Plate 

Read in data files for C Cold plate 
```{r}
c_cold_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="df_O2_Fm")
c_cold_QY<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="df_QY")
c_cold_NPQ<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="df_NPQ")
c_cold_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/C-Cold-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
c_cold_O2_fm<-c_cold_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., c_cold_well_metadata)

c_cold_QY<-c_cold_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., c_cold_well_metadata)

c_cold_NPQ<-c_cold_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., c_cold_well_metadata)

c_cold<- c_cold_O2_fm%>%
  left_join(c_cold_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(c_cold_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

c_cold$Plate<-"C-Cold"
```

## C Warm Plate


Read in data files for C Warm plate 
```{r}
c_warm_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="df_O2_Fm")
c_warm_QY<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="df_QY")
c_warm_NPQ<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="df_NPQ")
c_warm_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/C-Warm-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric.

```{r}
c_warm_O2_fm<-c_warm_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., c_warm_well_metadata)

c_warm_QY<-c_warm_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., c_warm_well_metadata)

c_warm_NPQ<-c_warm_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., c_warm_well_metadata)

c_warm<- c_warm_O2_fm%>%
  left_join(c_warm_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(c_warm_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

c_warm$Plate<-"C-Warm"
```

## MIX Cold Plate

Read in data files for MIX Cold plate 
```{r}
mix_cold_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="df_O2_Fm")
mix_cold_QY<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="df_QY")
mix_cold_NPQ<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="df_NPQ")
mix_cold_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/Mix-Cold-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
mix_cold_O2_fm<-mix_cold_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., mix_cold_well_metadata)

mix_cold_QY<-mix_cold_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., mix_cold_well_metadata)

mix_cold_NPQ<-mix_cold_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., mix_cold_well_metadata)

mix_cold<- mix_cold_O2_fm%>%
  left_join(mix_cold_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(mix_cold_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

mix_cold$Plate<-"MIX-Cold"
```

## MIX Warm Plate

Read in data files for MIX Warm plate 
```{r}
mix_warm_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="df_O2_Fm")
mix_warm_QY<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="df_QY")
mix_warm_NPQ<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="df_NPQ")
mix_warm_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/Mix-Warm-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
mix_warm_O2_fm<-mix_warm_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., mix_warm_well_metadata)

mix_warm_QY<-mix_warm_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., mix_warm_well_metadata)

mix_warm_NPQ<-mix_warm_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., mix_warm_well_metadata)

mix_warm<- mix_warm_O2_fm%>%
  left_join(mix_warm_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(mix_warm_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

mix_warm$Plate<-"MIX-Warm"
```

## WT Cold Plate

Read in data files for WT Cold plate 
```{r}
wt_cold_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="df_O2_Fm")
wt_cold_QY<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="df_QY")
wt_cold_NPQ<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="df_NPQ")
wt_cold_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/WT-Cold-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
wt_cold_O2_fm<-wt_cold_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., wt_cold_well_metadata)

wt_cold_QY<-wt_cold_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., wt_cold_well_metadata)

wt_cold_NPQ<-wt_cold_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., wt_cold_well_metadata)

wt_cold<- wt_cold_O2_fm%>%
  left_join(wt_cold_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(wt_cold_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

wt_cold$Plate<-"WT-Cold"
```

## WT Warm Plate

Read in data files for WT Warm plate 
```{r}
wt_warm_O2_fm<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="df_O2_Fm")
wt_warm_QY<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="df_QY")
wt_warm_NPQ<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="df_NPQ")
wt_warm_well_metadata<-read_xlsx(path="data/phenoplate_oxygen/WT-Warm-Plate.xlsx", sheet="metadata_wells")
```

Prepare data frames. Make the data frame in a long version, attach metadata, and add column specifying the metric. 
```{r}
wt_warm_O2_fm<-wt_warm_O2_fm%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="O2_Fm_raw")%>%
  left_join(., wt_warm_well_metadata)

wt_warm_QY<-wt_warm_QY%>%
  rename(Time=`Time (min)`)%>%
  pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="QY")%>%
  left_join(., wt_warm_well_metadata)

wt_warm_NPQ<-wt_warm_NPQ%>%
  rename(Time=`Time (min)`)%>%
pivot_longer(names_to="well", cols=`Area 1`:`Area 24`, values_to="NPQ_raw")%>%
  left_join(., wt_warm_well_metadata)

wt_warm<- wt_warm_O2_fm%>%
  left_join(wt_warm_QY, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont")) %>%
  left_join(wt_warm_NPQ, by = c("well", "Time", "Light", "Gradient_Temperature", "Symbiont"))

wt_warm$Plate<-"WT-Warm"
```

## Join all files together 

Join together. 
```{r}
df<-full_join(c_cold, c_warm)
df<-full_join(df, mix_cold)
df<-full_join(df, mix_warm)
df<-full_join(df, wt_cold)
df<-full_join(df, wt_warm)

df$code<-paste(df$Plate, df$well)
```

Change to factors. 
```{r}
df$Gradient_Temperature<-as.factor(df$Gradient_Temperature)
df$Symbiont<-as.factor(df$Symbiont)
```

# Perform calculations 

## O2 Fm

Normalize each O2 observation to the first measurement at Time 0 for each well. Calculate the inverse value such that a higher Y axis value indicates more oxygen. This will make it easier to interpret. 
```{r}
df<-df%>%
  group_by(Plate, well)%>%
  mutate(O2_Fm_norm=O2_Fm_raw-first(O2_Fm_raw))%>%
  mutate(O2_Fm_norm=-1*O2_Fm_norm)
```

## QY

QY (relative quantum yield) values do not require any normalization. 

## NPQ

First calculate Fm for each well as average of values from the first 15 min dark period. Then calculate NPQ as (Fm-value)/value. 

```{r}
df <- df %>%
  # Filter rows where Time is within the first 15 minutes
  filter(Time <= 14.99) %>%
  # Group by 'well'
  group_by(Plate, well, code) %>%
  # Calculate the average of 'NPQ_raw' for each 'well'
  summarize(Fm = mean(NPQ_raw, na.rm = TRUE)) %>%
  # Join the calculated averages back to the original data
  left_join(df, by = c("Plate", "well", "code")) %>%
  # Calculate the 'NPQ_norm' column
  mutate(NPQ_norm = (Fm - NPQ_raw) / NPQ_raw)
```

# Set color palettes

Create colour palette for temperatures. 
```{r}
# Define the number of colors in the palette
num_colors <- 12

# Create a color palette from dark blue to dark red
palette <- colorRampPalette(c("darkblue", "lightblue", "red1", "darkred"))(num_colors)
```

Set custom theme for panel plots. 
```{r}
custom_theme<-theme(text=element_text(size=20, face="bold", color="black"), 
        axis.text=element_text(size=18, color="black"), 
        legend.position="none", 
        )
```

Make a light plot to add onto the bottom of each plot for reference. 
```{r}
light_plot<-df%>%
  dplyr::select(Time, Light)%>%
  group_by(Time, Light)%>%
  summarise(Light=mean(Light, na.rm=TRUE))%>%
  
  ggplot(aes(y=Light, x=Time))+
  geom_line(colour="black", linewidth=2)+
  geom_text(label="Dark (26°C)", x=8, y=250, size=4)+
  geom_text(label="Light (26°C)", x=22, y=600, size=4)+
  geom_text(label="Dark \n[T Gradient]", x=36, y=250, size=4)+
  geom_text(label="Low Light \n[T Gradient]", x=45, y=250, size=4)+
  geom_text(label="Rapid Light Curve \n[T Gradient]", x=58, y=1400, size=4)+
  geom_text(label="Low Light \nRecovery \n(26°C)", x=70, y=400, size=4)+
  theme_classic()+
  custom_theme+
  ylim(0, 1650)+
  xlab("Time (min)")+
  ylab("Light (PAR)");light_plot
```

Set names for parent phenotypes. 
```{r}
parent_names<-c("C"="Bleached", "MIX"="Nonbleached", "WT"="Wildtype")
```

# Oxygen PI-Curve Summary   

Load a dataframe of start and end times for each phase from the treatment metadata. 
```{r}
treatments<-read_csv(file="data/phenoplate_oxygen/treatment_metadata.csv")
```

Add a column for the phase of each observation in our data frame. 

```{r}
get_phase <- function(Time) {
  matching_phase <- treatments$Phase[Time >= treatments$Time.Start & Time <= treatments$Time.End]
  if (length(matching_phase) > 0) {
    return(matching_phase[1])
  } else {
    return(NA)
  }
}

df$Phase <- sapply(df$Time, get_phase)
```

Select only required columns. 

```{r}
df<-df%>%
  dplyr::select(Plate, well, code, Time, Light, Gradient_Temperature, Symbiont, O2_Fm_norm, NPQ_norm, QY, Phase)
```

Plot the data from the PI Curve faceted by parent type.

```{r}
pi_plot1<-df%>%
  filter(Phase=="Rapid light curve - Temp Gradient")%>%
  group_by(Symbiont, Phase, Time, Light, Gradient_Temperature)%>%
  summarise(mean=mean(O2_Fm_norm, na.rm=TRUE), se=sd(O2_Fm_norm, na.rm=TRUE)/sqrt(length(O2_Fm_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  facet_grid(~Symbiont, labeller = as_labeller(parent_names))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("Oxygen Concentration [-1*rel Fm]")+
  ggtitle("PI Curve")+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  ylim(1000, 14000)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); pi_plot1

ggsave(filename="figures/phenoplate/RLC_O2_summary.jpeg", pi_plot1, width=10, height=7)
```


Does not look like photosynthesis is being captured/measured. How do we want to proceed? 

# Fluorescence Summary 

## Summarise data before equation calculations  

### QY 

Plot the QY data from the RLC faceted by parent type.

```{r}
rlc_QY_plot1<-df%>%
  filter(Phase=="Rapid light curve - Temp Gradient")%>%
  group_by(Symbiont, Phase, Time, Light, Gradient_Temperature)%>%
  summarise(mean=mean(QY, na.rm=TRUE), se=sd(QY, na.rm=TRUE)/sqrt(length(QY)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  facet_grid(~Symbiont, labeller = as_labeller(parent_names))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("QY")+
  ggtitle("RLC: QY")+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  ylim(0, 0.7)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); rlc_QY_plot1

ggsave(filename="figures/phenoplate/RLC_QY_summary.jpeg", rlc_QY_plot1, width=10, height=7)
```


### NPQ 

Plot the NPQ data from the RLC faceted by parent type.

```{r}
rlc_NPQ_plot1<-df%>%
  filter(Phase=="Rapid light curve - Temp Gradient")%>%
  group_by(Symbiont, Phase, Time, Light, Gradient_Temperature)%>%
  summarise(mean=mean(NPQ_norm, na.rm=TRUE), se=sd(NPQ_norm, na.rm=TRUE)/sqrt(length(NPQ_norm)))%>%
  
  ggplot(aes(x=Time, y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  facet_grid(~Symbiont, labeller = as_labeller(parent_names))+
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), alpha=0.2, linewidth=0) +
  geom_line(aes(y = mean), linewidth=2)+
  xlab("Time (min)")+
  ylab("NPQ")+
  ggtitle("RLC: NPQ")+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  #ylim(0, 0.7)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); rlc_NPQ_plot1

ggsave(filename="figures/phenoplate/RLC_NPQ_summary.jpeg", rlc_NPQ_plot1, width=10, height=7)
```

Summarize NPQ values in last 5 minutes. Calculate as relative change in Fm per min. 
```{r}
NPQ_RLC<-df%>%
  filter(Phase=="Rapid light curve - Temp Gradient")%>%
  group_by(Plate, well, code, Symbiont, Gradient_Temperature, Light)%>%
  summarise(NPQrlc=mean(NPQ_norm, na.rm=TRUE))
```

Plot values. 
```{r}
NPQ_RLC_plot2<-NPQ_RLC%>%
  group_by(Symbiont, Gradient_Temperature, Light)%>%
  summarise(mean=mean(NPQrlc, na.rm=TRUE), se=sd(NPQrlc, na.rm=TRUE)/sqrt(length(NPQrlc)))%>%
  
  ggplot(aes(x=as.integer(as.character(Light)), y=mean, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  facet_grid(~Symbiont)+
  geom_point(size=2)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  geom_smooth(aes(group=Gradient_Temperature, fill=Gradient_Temperature, color=Gradient_Temperature), linewidth=2, alpha=0.1)+
  xlab("Light")+
  ylab(expression(bold(paste("Non-Photochemical Quenching"))))+
  ggtitle(expression(bold(paste("NPQ during RLC"))))+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  #ylim(-0.01,0.055)+
  theme_classic()+
  theme(legend.position="right", 
        text=element_text(size=14, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        title = element_text(size=12, color="black", face="bold")); NPQ_RLC_plot2

ggsave(filename="figures/phenoplate/RLC_NPQ_summary2.jpeg", NPQ_RLC_plot2, width=10, height=7)
```

Run ANOVA model on these values. 
```{r}
model<-NPQ_RLC%>%
  
  aov(NPQrlc~Symbiont*Gradient_Temperature*Light, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont)
pairs(emm)
cld(emm, Letters = letters)

emm<-emmeans(model, ~Gradient_Temperature)
pairs(emm)
cld(emm, Letters = letters)

emm<-emmeans(model, ~Symbiont | Gradient_Temperature)
pairs(emm)
cld(emm, Letters = letters)

emm<-emmeans(model, ~Symbiont * Light | Gradient_Temperature )
pairs(emm)
cld(emm, Letters = letters)
```

                                      Df Sum Sq Mean Sq  F value   Pr(>F)    
Symbiont                               2  16.94    8.47  621.992  < 2e-16 ***
Gradient_Temperature                  11  14.71    1.34   98.214  < 2e-16 ***
Light                                  1 120.17  120.17 8823.973  < 2e-16 ***
Symbiont:Gradient_Temperature         22   5.64    0.26   18.832  < 2e-16 ***
Symbiont:Light                         2   0.26    0.13    9.486 7.93e-05 ***
Gradient_Temperature:Light            11   6.01    0.55   40.091  < 2e-16 ***
Symbiont:Gradient_Temperature:Light   22   0.71    0.03    2.364 0.000353 ***
Residuals                           2088  28.44    0.01  

Plot values faceted by temperature. 
```{r}
NPQ_RLC_plot3<-NPQ_RLC%>%
  group_by(Symbiont, Gradient_Temperature, Light)%>%
  summarise(mean=mean(NPQrlc, na.rm=TRUE), se=sd(NPQrlc, na.rm=TRUE)/sqrt(length(NPQrlc)))%>%
  
  ggplot(aes(x=as.integer(as.character(Light)), y=mean, colour=Symbiont, fill=Symbiont))+
  facet_grid(~Gradient_Temperature)+
  geom_point(size=2)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  geom_smooth(aes(group=Symbiont, fill=Symbiont, color=Symbiont), linewidth=2, alpha=0.1)+
  xlab("Light")+
  ylab(expression(bold(paste("Non-Photochemical Quenching"))))+
  ggtitle(expression(bold(paste("NPQ during RLC"))))+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Symbiont")+
  #ylim(-0.01,0.055)+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=14, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        title = element_text(size=12, color="black", face="bold")); NPQ_RLC_plot3

ggsave(filename="figures/phenoplate/RLC_NPQ_summary3.jpeg", NPQ_RLC_plot3, width=20, height=7)
```

# Rapid Light Curve 

## Calculate rETR 

First, calculate rETR as QY * PAR. 
```{r}
rlc<-df%>%
  filter(Phase=="Rapid light curve - Temp Gradient")%>%
  mutate(rETR=QY*Light)
```

Plot rETR across PAR with line for each temperature faceted by parent type.    
```{r}
rlc_plot1<-rlc%>%

  ggplot(aes(x=Light, y=rETR, colour=Gradient_Temperature, fill=Gradient_Temperature))+
  facet_grid(~Symbiont, labeller = as_labeller(parent_names))+
  geom_point()+
  geom_smooth(aes(group=Gradient_Temperature), se=FALSE, linewidth=2)+
  xlab("Light (PAR)")+
  ylab("rETR")+
  scale_colour_manual(values=palette, name="Gradient \nTemperature")+
  scale_fill_manual(values=palette, name="Gradient \nTemperature")+
  theme_classic()+
  theme(legend.position="none", 
        text=element_text(size=18, color="black"), 
        axis.title=element_text(size=20, color="black", face="bold"), 
        title = element_text(size=20, color="black", face="bold")); rlc_plot1
```

## Fit nls model 

Fit a model to the data following Platt 1976, Platt 1980 using light dependent model of photosynthesis with photoinhibition because we see the downward trend in rETR at the highest irradiance values.

From this we will extract parameters: 
- alpha (slope under light limitation)
- beta (photoinhibition)
- max rETR (maximal rETR)
- Ek (saturating irradiance)

Reference: https://enveurope.springeropen.com/articles/10.1186/s12302-020-00306-9 

This model runs, but doesn't account for beta currently 



For now, running - 

Nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   

```{r}
library(purrr)
library(minpack.lm)

nest <- nest_legacy
unnest <- unnest_legacy

rlc$group <- paste(rlc$Symbiont, rlc$Gradient_Temperature)

#test starting values
-min(rlc$rETR)
max(rlc$rETR) - min(rlc$rETR)
as.numeric(as.character(rlc$Light))

nls_data <- rlc %>%
  group_by(code) %>%
  nest() %>%
  mutate(model1 = map(data, ~ {
    # Calculate I within the map function for each group
    I <- as.numeric(as.character(.x$Light))
    
    # Fit the model with adjusted starting values and control parameters
    #nls_fit <- nlsLM(rETR ~ (Ps*(1-exp(-(alpha * I)/Ps))*exp(-(beta * I)/Ps) - Rd),
    #                 data = .x, 
    #                 start = list(Ps = 200, alpha = 5, beta = 1, Rd = 10))
    
    nls_fit <- nlsLM(rETR ~ (Am*((AQY*I)/(sqrt(Am^2 + (AQY*I)^2)))-Rd), 
                     data=., 
                     start=list(Am=200,  AQY=5, Rd=10))
    
    # Return tidy results
    tidy(nls_fit) %>%
      dplyr::select(term, estimate) %>%
      spread(term, estimate)
  })) %>%
  unnest(model1) %>%
  unnest(data) %>%
  group_by(code) %>%
  summarise(rETRmax=mean(Am), AQY=mean(AQY), dark_rETR=mean(Rd))%>%
  mutate(Ik = rETRmax/AQY) %>%
  mutate(Ic=(rETRmax*dark_rETR)/(AQY*(sqrt(rETRmax^2-dark_rETR^2))))
  #summarise(Ps = mean(Ps), alpha = mean(alpha), Rd = mean(Rd), beta = mean(beta)) %>%
  #mutate(Ik = (Ps/alpha)* log((alpha + beta)/beta),
  #       rETRmax = Ps*(alpha/(alpha + beta))*(beta/(alpha+beta))^(beta/alpha) - Rd,
  #       Ic = Rd/alpha)
```

Add in metadata and generate summaries. 

```{r}
results<-nls_data
results$Gradient_Temperature<-rlc$Gradient_Temperature[match(results$code, rlc$code)]
results$Symbiont<-rlc$Symbiont[match(results$code, rlc$code)]

results<-results%>%
  pivot_longer(names_to="Metric", values_to=("Value"), cols=rETRmax:Ic)
```

## Plot results 

```{r}
rlc_metrics_plot1<-results%>%
  group_by(Symbiont, Gradient_Temperature, Metric)%>%
  summarise(mean=mean(Value, na.rm=TRUE), se=sd(Value, na.rm=TRUE)/sqrt(length(Value)))%>%
  
  ggplot(aes(x=Gradient_Temperature, y=mean, colour=Symbiont, fill=Symbiont))+
  facet_grid(Metric ~., scales = "free_y")+
  geom_point(size=2)+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.1)+
  geom_smooth(aes(group=Symbiont, fill=Symbiont, color=Symbiont), linewidth=2, alpha=0.1)+
  xlab("Temperature")+
  ylab(expression(bold(paste("Value"))))+
  scale_colour_manual(values=c("orange", "brown4", "gray"), name="Parental \nPhenotype", labels=c("Bleached", "Nonbleached", "Wildtype"))+
  scale_fill_manual(values=c("orange", "brown4", "gray"), name="Parental \nPhenotype", labels=c("Bleached", "Nonbleached", "Wildtype"))+
  #ylim(-0.01,0.055)+
  theme_classic()+
  theme(legend.position="right", 
        text=element_text(size=14, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        title = element_text(size=12, color="black", face="bold")); rlc_metrics_plot1

ggsave(filename="figures/phenoplate/rlc_metrics.jpeg", rlc_metrics_plot1, width=7, height=10)
```

Ic isn't very meaningful to be below 0 though? What about Rd/dark rETR? 

## ANOVA test

AQY. 
```{r}
model<-results%>%
  filter(Metric=="AQY")%>%
  
  aov(Value~Gradient_Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont | Gradient_Temperature)
pairs(emm)
```
Significant temperature and interaction term. At low temperatures, alpha is higher in MIX than C. 

Ic 
```{r}
model<-results%>%
  filter(Metric=="Ic")%>%
  
  aov(Value~Gradient_Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont | Gradient_Temperature)
pairs(emm)
```
Significant main effects and interaction. Ic is lower in C than other groups. At low temperatures C is lower than other groups. 

Ic 
```{r}
model<-results%>%
  filter(Metric=="Ik")%>%
  
  aov(Value~Gradient_Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont | Gradient_Temperature)
pairs(emm)
```
Significant main effects and interaction. Ik is higher in C at 38, but barely. C is higher than WT at 31. C is higher than other groups at low temperatures. 

Rd (dark rETR essentially)
```{r}
model<-results%>%
  filter(Metric=="dark_rETR")%>%
  
  aov(Value~Gradient_Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont | Gradient_Temperature)
pairs(emm)
```
Significant main effects and interaction. Dark rETR is lower in C at 31 than other groups. C is lower than other groups at low temperatures. 

rETR max 
```{r}
model<-results%>%
  filter(Metric=="rETRmax")%>%
  
  aov(Value~Gradient_Temperature*Symbiont, data=.)

summary(model)

emm<-emmeans(model, ~Symbiont)
pairs(emm)

emm<-emmeans(model, ~Gradient_Temperature)
cld(emm, Letters=letters)
```
Significant main effects of symbiont and gradient temp. rETR is higher in mix than WT, but no difference in MIX and C. Bimodal temperature effects. 
