---
title: Protein analysis for lipid and metabolite extractions
author: "AS Huffmyer"
date: '2024'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes protein data for normalization for lipid and metabolite extractions. 

# Set up 

```{r}
if (!require("dplyr")) install.packages("dplyr")
if (!require("readr")) install.packages("readr")
if (!require("stringr")) install.packages("stringr")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("grid")) install.packages("grid")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("lattice")) install.packages("lattice")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lsmeans")) install.packages("lsmeans")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("car")) install.packages("car")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("corrr")) install.packages("corrr")
if (!require("GGally")) install.packages("GGally")
if (!require("cowplot")) install.packages("cowplot")

library(dplyr)
library(readr)
library(stringr)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(Rmisc)
library(ggpubr)
library(lsmeans)
library(car)
library(kableExtra)
library(ggcorrplot)
library(corrr)
library(GGally)
library(cowplot)
library(tidyverse)
```

# Set plot theme 

```{r}
custom_theme<-theme_classic() + 
  theme(axis.text = element_text(size=12, color="black"), 
        axis.title=element_text(size=12, color="black", face="bold"), 
        legend.text=element_text(size=12, color="black"), 
        legend.title=element_text(size=12, face="bold", color="black"))
```

# Protein

## Load data  

Read in datafiles

```{r}
protein <- read.csv("data/lipids_metabolites/protein/test_protein_data.csv")
metadata <- read.csv("data/lipids_metabolites/test_metadata.csv")
```

## Prepare dataframes  

Merging Files and renaming columns

```{r}
protein <- merge(metadata, protein, by = c("Well", "Run"))
#protein <- left_join(protein, volumes, by=c("Tube.ID"))

# Blank correction for each run separately

Blank <- protein %>% 
  filter(Sample.Type == "Blank") %>%
  summarise(blk.avg = mean(X562))

protein$abs.corr <- protein$X562 - Blank$blk.avg
```

Plot standard curve regressions for each plate.  

```{r}
protein %>% 
  filter(Sample.Type == "Standard")%>%

  ggplot(aes(x=Concentration, y=abs.corr))+
  facet_wrap(~ Run)+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  ggtitle("Run")+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = after_stat(eq.label))) +
  custom_theme
```

Run a loop to extract regression for each plate and calculate protein concentration for samples within each plate.  

```{r}
run_list<-c("1", "2", "3")

df_protein = data.frame()

for (run in run_list) {

  #subset data
Standard <- protein %>% 
  filter(Sample.Type == "Standard")%>%
  filter(Run==run)
  
#generate standard curve equation
lmstandard <- lm (Concentration ~ abs.corr, data = Standard)
lmsummary <- summary(lmstandard) 

#select samples
Samples <- protein %>% #subsetting Samples
  filter(Sample.Type == "Sample")%>%
  filter(Run==run)

#calculate concentration
Samples$Concentration <- predict(lmstandard, newdata = Samples) #using model to get concentration

#add run column
Samples$Run<-run

#normalize to homogenate volume - used because samples were diluted - adjust as needed 
Samples$Protein.ug.mL <- Samples$Concentration * (25/Samples$Homo_vol)

#join dataframes
df <- data.frame(Samples)
df_protein <- rbind(df_protein,df)

} 
```

Plot results 

```{r}
Protein.Plot <- df_protein%>%
  filter(!Treatment=="Astrangia")%>%
  filter(Fraction=="Host")%>%
  ggplot(aes(x=Sample.ID, y=Protein.ug.mL)) +
  geom_point(pch = 21, position=position_dodge(0.4)) +
  xlab("Temperature") + ylab(expression(bold("Protein " (ug/mL)))) + #Axis titles
  custom_theme; Protein.Plot
```

Calculate total protein per unit volume.

## Output file of volume needed for each sample to obtain a total of 50 µg protein   

```{r}
volume.needed <- 50 

df_protein%>%
  filter(!Treatment=="Astrangia")%>%
  filter(Fraction=="Host")%>%
  dplyr::select(Tube.ID, Sample.ID, Protein.ug.mL)%>%
  group_by(Tube.ID, Sample.ID)%>%
  dplyr::summarise(Protein.ug.mL=mean(Protein.ug.mL))%>%
  dplyr::mutate(uL.50ug.protein=(volume.needed/Protein.ug.mL)*1000)%>%
  write_csv(., "output/lipids_metabolites/protein/volume_50ug_protein.csv")
```

