---
title: "Larval PI Curve Plotting and Analysis - Hawaii 2023 - Temp x PI curve analysis" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('emmeans')
```


# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pi_temp_data<-read.csv("output/temp_pi_curves/temp_PIcurves_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pi_temp_data<-pi_temp_data[!is.na(pi_temp_data$Type),]

#format columns
pi_temp_data$Tank<-as.factor(pi_temp_data$Tank)
pi_temp_data$Symbiont<-as.factor(pi_temp_data$Symbiont)
pi_temp_data$SDR<-as.factor(pi_temp_data$SDR)
pi_temp_data$Plate<-as.factor(pi_temp_data$Plate)
pi_temp_data$Temperature<-as.factor(pi_temp_data$Temperature)
```

Remove plate that didnt work 

Removing plate 13 at light 3 (100 PAR) for now, the data are super variable and weird due to probe issue. Need to revisit and trim this section. 

NEED TO FIX THIS - ITS ASSIGNED AS PLATE 5 RATHER THAN 13

```{r}
#pi_temp_data<-pi_temp_data%>%
#  filter(!c(Plate==5 & Light==3))
```

Look for outliers in the data.  

```{r}
boxplot(pi_temp_data$P.nmol.org.min)
#pi_temp_data<-pi_temp_data%>%filter(P.nmol.org.min < 0.10)
#boxplot(pi_temp_data$P.nmol.org.min)

boxplot(pi_temp_data$P.nmol.org.min~pi_temp_data$PAR)
```

Set values at 0 for light runs if <0 indicating no photosynthesis occurred. Set values at 0 for dark runs if >0 indicating no respiration occurred. 
```{r}
pi_temp_data<-pi_temp_data%>%
  mutate(P.nmol.org.min=if_else(P.nmol.org.min<0 & PAR==50, 0, 
                      if_else(P.nmol.org.min<0 & PAR==100, 0, 
                              if_else(P.nmol.org.min<0 & PAR==500, 0, 
                                      if_else(P.nmol.org.min>0 & PAR==0.1, 0, 
                                              if_else(P.nmol.org.min>0 & PAR==0.2, 0, P.nmol.org.min))))))

boxplot(pi_temp_data$P.nmol.org.min~pi_temp_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/temp_pi_curves/runs_temp.csv")
pr.temps = subset(pr.temps, select = -c(X) ) #remove empty column

pi_temp_data$Plate<-as.integer(pi_temp_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pi_temp_data<-left_join(pi_temp_data,pr.temps)

#round to 0.1°C 
pi_temp_data<-pi_temp_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

Reorder factor. 
```{r}
pi_temp_data$PAR<-factor(pi_temp_data$PAR, levels=c("0.1", "50", "100", "500", "0.2"))
```

# Plot and analyze 

Plot data with means   
```{r}
pi_temp_plot1<-pi_temp_data %>%
    group_by(PAR, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot1

ggsave("figures/temp_pi_curves/pi_means.png", pi_temp_plot1, dpi=300, w=7, h=5, units="in")
```

Display with plate. 
```{r}
pi_temp_plot1a<-pi_temp_data %>%
    group_by(PAR, Symbiont, Plate)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot1a
```

View by plate. 
```{r}
pi_temp_plot2b<-pi_temp_data %>%
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=Symbiont)) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot2b
```

Plot data with dots   
```{r}
pi_temp_plot3<-pi_temp_data %>%
  
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot3

ggsave("figures/temp_pi_curves/pi_temp_dots.png", pi_temp_plot3, dpi=300, w=7, h=5, units="in")
```

## Conduct anova 

Run anova on values at each temperature
```{r}
model<-pi_temp_data%>%
  aov(P.nmol.org.min~PAR*Symbiont*Temperature, data=.)

summary(model)
```
There is an effect of temperature x light. 

Run model for 27 
```{r}
model<-pi_temp_data%>%
  filter(Temperature==27)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```
There is an effect of light at 27C. 

Run model for 30
```{r}
model<-pi_temp_data%>%
  filter(Temperature==30)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```
There is an effect of light at 30C. 

Run model for 33
```{r}
model<-pi_temp_data%>%
  filter(Temperature==33)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```
There is an effect of light at 33C. 

Run model for 36
```{r}
model<-pi_temp_data%>%
  filter(Temperature==36)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```
There is an effect of light at 36C. 

# Calculate R, P, and P:R 

Calculate gross photosynthesis and P:R ratio. 
```{r}
pi_temp_data_calc<-pi_temp_data%>%
  select(Date, Chamber.ID, Plate, Run, Temperature, Symbiont, PAR, P.nmol.org.min)%>%
  pivot_wider(names_from=PAR, values_from=P.nmol.org.min)%>%
  mutate(ler=`0.2`)%>%
  mutate(rd=`0.1`)%>%
  mutate(ler_inverse=ler*-1)%>%
  mutate(rd_inverse=rd*-1)%>%
  mutate(p_net_low=`50`)%>%
  mutate(p_net_medium=`100`)%>%
  mutate(p_net_high=`500`)%>%
  mutate(p_gross_low=p_net_low-ler)%>%
  mutate(p_gross_medium=p_net_medium-ler)%>%
  mutate(p_gross_high=p_net_high-ler)%>%
  mutate(ratio_low=p_gross_low/ler_inverse)%>%
  mutate(ratio_medium=p_gross_medium/ler_inverse)%>%
  mutate(ratio_high=p_gross_high/ler_inverse)%>%
  pivot_longer(names_to="Metric", values_to="Value", cols=c(ler:ratio_high))%>%
  select(!c(`0.1`, `50`, `100`, `500`, `0.2`))

boxplot(pi_temp_data_calc$Value)

pi_temp_data_calc<-pi_temp_data_calc%>%
 filter(Value<15)

boxplot(pi_temp_data_calc$Value)
```

Set level of factors. 
```{r}
levels(as.factor(pi_temp_data_calc$Metric))

pi_temp_data_calc$Metric<-factor(pi_temp_data_calc$Metric, levels=c("p_gross_low", "p_gross_medium", "p_gross_high", "p_net_low", "p_net_medium", "p_net_high", "ratio_low", "ratio_medium", "ratio_high", "rd", "rd_inverse", "ler", "ler_inverse"))
```

View P metrics by dots. 
```{r}
metric_plot1<-pi_temp_data_calc%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_medium")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
    filter(!Metric=="rd")%>%
    filter(!Metric=="ler_inverse")%>%
    filter(!Metric=="rd_inverse")%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1

ggsave("figures/temp_pi_curves/pi_metrics.png", metric_plot1, dpi=300, w=7, h=5, units="in")
```

View P metrics by means. 
```{r}
metric_plot1_means<-pi_temp_data_calc%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_medium")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
    filter(!Metric=="rd")%>%
    filter(!Metric=="ler_inverse")%>%
    filter(!Metric=="rd_inverse")%>%

    group_by(Metric, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
  
  
    ggplot(., aes(x = Temperature, y = mean)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=4, fill="white", position = position_dodge(0.4), shape=19) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Metric, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.6, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1_means

ggsave("figures/temp_pi_curves/pi_metrics_mean.png", metric_plot1_means, dpi=300, w=7, h=5, units="in")
```

View Rd and LER metrics. 
```{r}
metric_plot1a<-pi_temp_data_calc%>%
    filter(Metric %in% c("ler_inverse", "rd_inverse"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1a

ggsave("figures/temp_pi_curves/pi_rd_ler.png", metric_plot1a, dpi=300, w=7, h=5, units="in")
```

LER is higher than Rd because photosynthesis stimulates respiration in the host.  

View Rd and LER metrics by means. 
```{r}
metric_plot1a_means<-pi_temp_data_calc%>%
    filter(Metric %in% c("ler_inverse", "rd_inverse"))%>%
  
    group_by(Metric, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
  
    ggplot(., aes(x = Temperature, y = mean)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_dodge(0.4), shape=19) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Metric, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.6, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1a_means

ggsave("figures/temp_pi_curves/pi_rd_ler_means.png", metric_plot1a_means, dpi=300, w=7, h=5, units="in")
```

View P:R metrics. 
```{r}
metric_plot2<-pi_temp_data_calc%>%
    filter(Metric %in% c("ratio_high", "ratio_medium", "ratio_low"))%>%
    filter(Value>0)%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2

ggsave("figures/temp_pi_curves/pi_ratio.png", metric_plot2, dpi=300, w=7, h=5, units="in")
```

View P:R ratios by means. 
```{r}
metric_plot2_means<-pi_temp_data_calc%>%
    filter(Metric %in% c("ratio_high", "ratio_medium", "ratio_low"))%>%
    filter(Value>0)%>%
  
    group_by(Metric, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%

    ggplot(., aes(x = Temperature, y = mean)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_dodge(0.4), shape=19) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Metric, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.6, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2_means

ggsave("figures/temp_pi_curves/pi_ratio_means.png", metric_plot2_means, dpi=300, w=7, h=5, units="in")
```

# Run anova tests

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ratio_high")%>%
  filter(Value>0)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
Ratios are lower at 36 C under low light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ratio_medium")%>%
  filter(Value>0)%>% 
  aov(Value~Temperature*Symbiont, data=.)

summary(model)
```
Ratios decrease with temperature under medium light.  

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ratio_low")%>%
  filter(Value>0)%>% 
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
Ratios decrease with temperature at high light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ler_inverse")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
TukeyHSD(model, c("Symbiont"))
```
LER is affected by temperature and symbiont. LER is higher in WT than Mix.

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="rd_inverse")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Symbiont"))
TukeyHSD(model, c("Temperature"))
```
Rd is affected by temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_gross_high")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P gross decreases with temperature when at high light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_gross_medium")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P gross is affected by temperature at medium light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_gross_low")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P gross at low light decreases with temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_net_low")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P net at low light decreases with temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_net_medium")%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P net at medium light decreases with temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_net_high")%>%
  aov(Value~Symbiont*Temperature, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P net at high light decreases with temperature. 

## Test effect of light 

Test effect of light on P:R ratio. 
```{r}
model<-pi_temp_data_calc%>%
  filter(Metric %in% c("ratio_low", "ratio_medium", "ratio_high"))%>%
  filter(Value>0)%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")
TukeyHSD(model, "Symbiont")

emm<-emmeans(model, ~ Metric | Temperature)
pairs(emm)

emm<-emmeans(model, ~ Symbiont | Temperature)
pairs(emm)

emm<-emmeans(model, ~ Symbiont | Temperature | Metric)
pairs(emm)
```
There are light and temperature effects on PR ratios. 

Test effect of light on P gross.  
```{r}
model<-pi_temp_data_calc%>%
  filter(Metric %in% c("p_gross_low", "p_gross_medium", "p_gross_high"))%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Metric | Temperature)
pairs(emm)
```
Interaction between light and temperature for P gross. There is almost a temp x symbiont effect (p=0.08). 

Test effect of light on P net.   
```{r}
model<-pi_temp_data_calc%>%
  filter(Metric %in% c("p_net_low", "p_net_medium", "p_net_high"))%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Metric | Temperature)
pairs(emm)
```
Interaction between temperature and light on P net. 