---
title: "Larval PI Curve Plotting and Analysis - Hawaii 2023 - Temp x PI curve analysis" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

NEED TO DO ANOTHER ROUND OF QC ON THESE MEASUREMENTS 

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
```


# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pi_temp_data<-read.csv("output/temp_pi_curves/temp_PIcurves_calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pi_temp_data<-pi_temp_data[!is.na(pi_temp_data$Type),]

#format columns
pi_temp_data$Tank<-as.factor(pi_temp_data$Tank)
pi_temp_data$Symbiont<-as.factor(pi_temp_data$Symbiont)
pi_temp_data$SDR<-as.factor(pi_temp_data$SDR)
pi_temp_data$Plate<-as.factor(pi_temp_data$Plate)
pi_temp_data$Temperature<-as.factor(pi_temp_data$Temperature)
```

Remove plate that didnt work 

Removing plate 13 at light 3 (100 PAR) for now, the data are super variable and weird due to probe issue. Need to revisit and trim this section. 

NEED TO FIX THIS - ITS ASSIGNED AS PLATE 5 RATHER THAN 13

```{r}
pi_temp_data<-pi_temp_data%>%
  filter(!c(Plate==5 & Light==3))
```

Look for outliers in the data.  

```{r}
boxplot(pi_temp_data$P.nmol.org.min)
#pi_temp_data<-pi_temp_data%>%filter(P.nmol.org.min < 0.10)
#boxplot(pi_temp_data$P.nmol.org.min)

boxplot(pi_temp_data$P.nmol.org.min~pi_temp_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pr.temps<-read.csv("output/temp_pi_curves/runs_temp.csv")
pr.temps = subset(pr.temps, select = -c(X) ) #remove empty column

pi_temp_data$Plate<-as.integer(pi_temp_data$Plate)
```

Add temperature data to master data frame.  
```{r}
pi_temp_data<-left_join(pi_temp_data,pr.temps)

#round to 0.1°C 
pi_temp_data<-pi_temp_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

Reorder factor. 
```{r}
pi_temp_data$PAR<-factor(pi_temp_data$PAR, levels=c("0.1", "50", "100", "500", "0.2"))
```

# Plot and analyze 

Plot data with means   
```{r}
pi_temp_plot1<-pi_temp_data %>%
    group_by(PAR, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot1

ggsave("figures/temp_pi_curves/pi_means.png", pi_temp_plot1, dpi=300, w=7, h=5, units="in")
```

Display with plate. 
```{r}
pi_temp_plot1a<-pi_temp_data %>%
    group_by(PAR, Symbiont, Plate)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(PAR, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot1a
```

View by plate. 
```{r}
pi_temp_plot2b<-pi_temp_data %>%
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=Symbiont)) +
    facet_wrap(~Plate)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot2b
```

Plot data with dots   
```{r}
pi_temp_plot3<-pi_temp_data %>%
  
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=interaction(PAR, Symbiont))) +
    facet_wrap(~Temperature)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(PAR, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Light (PAR)") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_temp_plot3

ggsave("figures/temp_pi_curves/pi_temp_dots.png", pi_temp_plot3, dpi=300, w=7, h=5, units="in")
```

## Conduct anova 

Run anova on values at each temperature
```{r}
model<-pi_temp_data%>%
  aov(P.nmol.org.min~PAR*Symbiont*Temperature, data=.)

summary(model)
```

Run model for 27 
```{r}
model<-pi_temp_data%>%
  filter(Temperature==27)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```

Run model for 30
```{r}
model<-pi_temp_data%>%
  filter(Temperature==30)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```

Run model for 33
```{r}
model<-pi_temp_data%>%
  filter(Temperature==33)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```

Run model for 36
```{r}
model<-pi_temp_data%>%
  filter(Temperature==36)%>%
  aov(P.nmol.org.min~PAR*Symbiont, data=.)

summary(model)
```
Almost an interaction here. 

# Calculate R, P, and P:R 

Calculate gross photosynthesis and P:R ratio. 
```{r}
pi_temp_data_calc<-pi_temp_data%>%
  select(Date, Chamber.ID, Plate, Run, Temperature, Symbiont, PAR, P.nmol.org.min)%>%
  pivot_wider(names_from=PAR, values_from=P.nmol.org.min)%>%
  mutate(ler=`0.1`)%>%
  mutate(rd=`0.2`)%>%
  mutate(ler_inverse=ler*-1)%>%
  mutate(rd_inverse=rd*-1)%>%
  mutate(p_net_low=`50`)%>%
  mutate(p_net_medium=`100`)%>%
  mutate(p_net_high=`500`)%>%
  mutate(p_gross_low=p_net_low-ler)%>%
  mutate(p_gross_medium=p_net_medium-ler)%>%
  mutate(p_gross_high=p_net_high-ler)%>%
  mutate(ratio_low=p_gross_low/(ler_inverse))%>%
  mutate(ratio_medium=p_gross_medium/(ler_inverse))%>%
  mutate(ratio_high=p_gross_high/(ler_inverse))%>%
  pivot_longer(names_to="Metric", values_to="Value", cols=c(ler:ratio_high))%>%
  select(!c(`0.1`, `50`, `100`, `500`, `0.2`))

boxplot(pi_temp_data_calc$Value)

pi_temp_data_calc<-pi_temp_data_calc%>%
 filter(Value>-10) %>% 
 filter(Value<20)

boxplot(pi_temp_data_calc$Value)
```

Set level of factors. 
```{r}
levels(as.factor(pi_temp_data_calc$Metric))

pi_temp_data_calc$Metric<-factor(pi_temp_data_calc$Metric, levels=c("p_gross_low", "p_gross_medium", "p_gross_high", "p_net_low", "p_net_medium", "p_net_high", "ratio_low", "ratio_medium", "ratio_high", "rd", "rd_inverse", "ler", "ler_inverse"))
```

View P metrics by dots. 
```{r}
metric_plot1<-pi_temp_data_calc%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_medium")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
    filter(!Metric=="rd")%>%
    filter(!Metric=="ler_inverse")%>%
    filter(!Metric=="rd_inverse")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
    #filter(!Value< -0.07)%>%
    #filter(!Value> 0.07)%>%
    #filter(!c(Value>0.05 & Metric=="p_net_low"))%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1

ggsave("figures/temp_pi_curves/pi_metrics.png", metric_plot1, dpi=300, w=7, h=5, units="in")
```

View P metrics by means. 
```{r}
metric_plot1_means<-pi_temp_data_calc%>%
    filter(!Metric=="ratio_high")%>%
    filter(!Metric=="ratio_medium")%>%
    filter(!Metric=="ratio_low")%>%
    filter(!Metric=="ler")%>%
    filter(!Metric=="rd")%>%
    filter(!Metric=="ler_inverse")%>%
    filter(!Metric=="rd_inverse")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%

    group_by(Metric, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
  
  
    ggplot(., aes(x = Temperature, y = mean)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=4, fill="white", position = position_dodge(0.4), shape=19) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Metric, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.6, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1_means

ggsave("figures/temp_pi_curves/pi_metrics_mean.png", metric_plot1_means, dpi=300, w=7, h=5, units="in")
```

View Rd and LER metrics. 
```{r}
metric_plot1a<-pi_temp_data_calc%>%
    filter(Metric %in% c("ler_inverse", "rd_inverse"))%>%
    #filter(!c(Value<0 & Metric=="ler_inverse"))%>%
    #filter(!c(Value>0.05 & Metric=="p_net_low"))%>%
    filter(!Value< -0.02)%>%
    filter(Value>0)%>%
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1a

ggsave("figures/temp_pi_curves/pi_rd_ler.png", metric_plot1a, dpi=300, w=7, h=5, units="in")
```

View Rd and LER metrics by means. 
```{r}
metric_plot1a_means<-pi_temp_data_calc%>%
    filter(Metric %in% c("ler_inverse", "rd_inverse"))%>%
    #filter(!c(Value<0 & Metric=="ler_inverse"))%>%
    #filter(!c(Value>0.05 & Metric=="p_net_low"))%>%
    filter(!Value< -0.02)%>%
    filter(Value>0)%>%
  
    group_by(Metric, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
  
    ggplot(., aes(x = Temperature, y = mean)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_dodge(0.4), shape=19) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Metric, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.6, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot1a_means

ggsave("figures/temp_pi_curves/pi_rd_ler_means.png", metric_plot1a_means, dpi=300, w=7, h=5, units="in")
```

View P:R metrics. 
```{r}
metric_plot2<-pi_temp_data_calc%>%
    filter(Metric %in% c("ratio_high", "ratio_medium", "ratio_low"))%>%
    filter(Value<15)%>%
    filter(Value>-7)%>%
  
    #filter(!c(Metric=="ratio_high" & Value>4 & Symbiont=="C" & Temperature==33))%>% #remove outliers 
    #filter(!c(Metric=="ratio_high" & Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
    #filter(!c(Metric=="ratio_low" & Value< -1 & Symbiont=="WT" & Temperature==27))%>% #remove outliers 
    #filter(!Value<0)%>% #remove outliers 
  
    ggplot(., aes(x = Temperature, y = Value)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_jitterdodge(0.4), shape=21) + 
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", size=2, alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2

ggsave("figures/temp_pi_curves/pi_ratio.png", metric_plot2, dpi=300, w=7, h=5, units="in")
```

View P:R ratios by means. 
```{r}
metric_plot2_means<-pi_temp_data_calc%>%
    filter(Metric %in% c("ratio_high", "ratio_medium", "ratio_low"))%>%
    filter(Value<15)%>%
    filter(Value>-7)%>%
  
    group_by(Metric, Symbiont, Temperature)%>%
    dplyr::summarise(mean=mean(Value, na.rm=TRUE), sd=sd(Value, na.rm=TRUE), N=length(Value), se=sd/sqrt(N))%>%
  
    #filter(!c(Metric=="ratio_high" & Value>4 & Symbiont=="C" & Temperature==33))%>% #remove outliers 
    #filter(!c(Metric=="ratio_high" & Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
    #filter(!c(Metric=="ratio_low" & Value< -1 & Symbiont=="WT" & Temperature==27))%>% #remove outliers 
    #filter(!Value<0)%>% #remove outliers 
  
    ggplot(., aes(x = Temperature, y = mean)) +
    facet_wrap(~Metric)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temperature, Symbiont), color=Symbiont), size=3, fill="white", position = position_dodge(0.4), shape=19) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Metric, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.6, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("Value")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); metric_plot2_means

ggsave("figures/temp_pi_curves/pi_ratio_means.png", metric_plot2_means, dpi=300, w=7, h=5, units="in")
```

# Run anova tests

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ratio_high")%>%
  filter(Value<15)%>%
  filter(Value>-7)%>%
  #filter(!c(Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
  #filter(!Value<0)%>% #remove outliers 
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
Ratios decrease with temperature and are lower at 33 than at 27, but level off after that. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ratio_medium")%>%
  filter(Value<15)%>%
  filter(Value>-7)%>%
  #filter(!c(Value>4 & Symbiont=="C" & Temperature==33))%>% #remove outliers 
  #filter(!c(Value<0 & Symbiont=="WT" & Temperature==30))%>% #remove outliers 
  #filter(!Value<0)%>% #remove outliers 
  aov(Value~Temperature*Symbiont, data=.)

summary(model)
```
No effect of temp at medium light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ratio_low")%>%
  #filter(Value>0)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
Ratios decrease at higher light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="ler_inverse")%>%
  filter(!Value< -0.02)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
LER is affected by temperature and is lower at 33 than 27. After 33 degrees, LER increases again.  

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="rd_inverse")%>%
  filter(Value>0)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Symbiont"))
TukeyHSD(model, c("Temperature"))
```
WT and MIX are different, WT is higher than MIX. No difference between MIX and C or WT and C. There is higher dark respiration in WT. Dark respiration is higher at 30 and 33 than 27. Respiration at 36 is lower than at 30 and 33. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_gross_high")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P gross decreases with temperature when at high light. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_gross_medium")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P gross at medium light peaks at 30 then decreases with higher temperatures. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_gross_low")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P gross at low light decreases with temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_net_low")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P net at low light decreases with temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_net_medium")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```
P net at medium light decreases with temperature. 

```{r}
model<-pi_temp_data_calc%>%
  filter(Metric=="p_net_high")%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Symbiont*Temperature, data=.)

summary(model)

TukeyHSD(model, c("Temperature"))
```

P net at high light decreases with temperature. 

## Test effect of light 

Test effect of light on P:R ratio. 
```{r}
model<-pi_temp_data_calc%>%
  filter(Metric %in% c("ratio_low", "ratio_medium", "ratio_high"))%>%
  filter(Value<15)%>%
  filter(Value>-7)%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Metric | Temperature)
pairs(emm)
```
Ratios increase with light and decrease with temperature. PAR only affects P:R ratio when at high temperature (36). 

Test effect of light on P gross.  
```{r}
model<-pi_temp_data_calc%>%
  filter(Metric %in% c("p_gross_low", "p_gross_medium", "p_gross_high"))%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Metric | Temperature)
pairs(emm)
```
Interaction between light and temperature for P gross. 

Test effect of light on P net.   
```{r}
model<-pi_temp_data_calc%>%
  filter(Metric %in% c("p_net_low", "p_net_medium", "p_net_high"))%>%
    filter(Value<0.09)%>%
    filter(Value>-0.08)%>%
  aov(Value~Metric*Temperature*Symbiont, data=.)

summary(model)

TukeyHSD(model, "Metric")
TukeyHSD(model, "Temperature")

emm<-emmeans(model, ~ Metric | Temperature)
pairs(emm)
```
Interaction between temperature and light on P net. 

# TPC on dark respiration 

DOUBLE CHECK THIS, DOESN'T LOOK LIKE DATA IN PLOTS 

Run a TPC on the dark respiration data across 27-36C. 

```{r}
tpc<-pi_temp_data_calc%>%
  filter(Metric=="rd_inverse")%>%
  pivot_wider(names_from = Metric, values_from = Value)%>%
  filter(!rd_inverse< -0.02)%>%
  filter(rd_inverse>0)%>%
  mutate(Temperature=as.numeric(as.character(Temperature)))
```

Run TPC using same method as in `tpc_sdr_analysis.Rmd` script. 

## Run for C symbionts 

```{r}
tpc.C<-tpc%>%
  filter(Symbiont=="C")

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc.C$Temperature, tpc.C$rd_inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc.C$Temperature, tpc.C$rd_inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc.C$Temperature, tpc.C$rd_inverse, model_name = 'sharpeschoolhigh_1981')

fit.C <- nls_multstart(rd_inverse~sharpeschoolhigh_1981(temp = Temperature, r_tref,e,eh,th, tref = 27),
                                                     data = tpc.C,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit.C

# predict new data
new_data.C <- data.frame(temp = seq(min(tpc.C$Temperature), max(tpc.C$Temperature), by=((max(tpc.C$Temperature)-min(tpc.C$Temperature))/(nrow(tpc.C)-1))))

preds.C <- augment(fit.C, newdata = new_data.C)

# plot data and model fit
ggplot(tpc.C, aes(x=Temperature, y=rd_inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds.C, col = 'orange') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures: C')
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM.C <- minpack.lm::nlsLM(rd_inverse~sharpeschoolhigh_1981(temp = Temperature, r_tref, e, eh, th, tref = 27),
                        data = tpc.C,
                        start = coef(fit.C),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc.C)))

# bootstrap using case resampling
boot.C <- Boot(fit_nlsLM.C, method = 'case')

# look at the data
head(boot.C$t)

# create predictions of each bootstrapped model
boot_preds.C <- boot.C$t %>%
  as.data.frame() %>%
  drop_na() %>%
  dplyr::mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc.C$Temperature), max(tpc.C$Temperature), length.out = 100))) %>%
  ungroup() %>%
  dplyr::mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 27))

# calculate bootstrapped confidence intervals
boot_conf_preds.C <- group_by(boot_preds.C, temp) %>%
  dplyr::summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot.C <- ggplot() +
  geom_line(aes(temp, .fitted), preds.C, col = 'orange') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.C, fill = 'orange', alpha = 0.3) +
  geom_point(aes(Temperature, rd_inverse), tpc.C, size = 2, alpha = 0.5,col = 'orange') +
  theme_bw(base_size = 12) +
    theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: C");CI.plot.C

```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit.C) %>%
  # round for easy viewing
  mutate_all(round, 2)
```


## Run for MIX symbionts 

```{r}
tpc.MIX<-tpc%>%
  #filter(R.inverse<0.03)%>%
  #filter(!c(R.inverse>0.02 & Temp_corrected==37))%>%
  #filter(!c(R.inverse>0.02 & Temp_corrected==13))%>%
  #filter(!c(R.inverse< -0.01 & Temp_corrected==23))%>%
  #filter(!c(R.inverse< 0 & Temp_corrected==32))%>%
  #filter(!c(R.inverse< -0.005 & Temp_corrected==37))%>%
  #filter(!c(R.inverse< -0.005 & Temp_corrected==34))%>%
  filter(Symbiont=="MIX")

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc.MIX$Temperature, tpc.MIX$rd_inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc.MIX$Temperature, tpc.MIX$rd_inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc.MIX$Temperature, tpc.MIX$rd_inverse, model_name = 'sharpeschoolhigh_1981')

fit.MIX <- nls_multstart(rd_inverse~sharpeschoolhigh_1981(temp = Temperature, r_tref,e,eh,th, tref = 27),
                                                     data = tpc.MIX,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit.MIX

# predict new data
new_data.MIX <- data.frame(temp = seq(min(tpc.MIX$Temperature), max(tpc.MIX$Temperature), by=((max(tpc.MIX$Temperature)-min(tpc.MIX$Temperature))/(nrow(tpc.MIX)-1))))

preds.MIX <- augment(fit.MIX, newdata = new_data.MIX)

# plot data and model fit
ggplot(tpc.MIX, aes(x=Temperature, y=rd_inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds.MIX, col = 'brown4') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures: C')
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM.MIX <- minpack.lm::nlsLM(rd_inverse~sharpeschoolhigh_1981(temp = Temperature, r_tref, e, eh, th, tref = 27),
                        data = tpc.MIX,
                        start = coef(fit.MIX),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc.MIX)))

# bootstrap using case resampling
boot.MIX <- Boot(fit_nlsLM.MIX, method = 'case')

# look at the data
head(boot.MIX$t)

# create predictions of each bootstrapped model
boot_preds.MIX <- boot.MIX$t %>%
  as.data.frame() %>%
  drop_na() %>%
  dplyr::mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc.MIX$Temperature), max(tpc.MIX$Temperature), length.out = 100))) %>%
  ungroup() %>%
  dplyr::mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 27))

# calculate bootstrapped confidence intervals
boot_conf_preds.MIX <- group_by(boot_preds.MIX, temp) %>%
  dplyr::summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot.MIX <- ggplot() +
  geom_line(aes(temp, .fitted), preds.MIX, col = 'brown4') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.MIX, fill = 'brown4', alpha = 0.3) +
  geom_point(aes(Temperature, rd_inverse), tpc.C, size = 2, alpha = 0.5,col = 'brown4') +
  theme_bw(base_size = 12) +
    theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: MIX");CI.plot.MIX

```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit.MIX) %>%
  # round for easy viewing
  mutate_all(round, 2)
```


## Run for WT symbionts 

```{r}
tpc.WT<-tpc%>%
  filter(Symbiont=="WT")

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc.WT$Temperature, tpc.WT$rd_inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc.WT$Temperature, tpc.WT$rd_inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc.WT$Temperature, tpc.WT$rd_inverse, model_name = 'sharpeschoolhigh_1981')

fit.WT <- nls_multstart(rd_inverse~sharpeschoolhigh_1981(temp = Temperature, r_tref,e,eh,th, tref = 27),
                                                     data = tpc.WT,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit.WT

# predict new data
new_data.WT <- data.frame(temp = seq(min(tpc.WT$Temperature), max(tpc.WT$Temperature), by=((max(tpc.WT$Temperature)-min(tpc.WT$Temperature))/(nrow(tpc.WT)-1))))

preds.WT <- augment(fit.WT, newdata = new_data.WT)

# plot data and model fit
ggplot(tpc.WT, aes(x=Temperature, y=rd_inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds.WT, col = 'gray') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures: C')
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM.WT <- minpack.lm::nlsLM(rd_inverse~sharpeschoolhigh_1981(temp = Temperature, r_tref, e, eh, th, tref = 27),
                        data = tpc.WT,
                        start = coef(fit.WT),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc.WT)))

# bootstrap using case resampling
boot.WT <- Boot(fit_nlsLM.WT, method = 'case')

# look at the data
head(boot.WT$t)

# create predictions of each bootstrapped model
boot_preds.WT <- boot.WT$t %>%
  as.data.frame() %>%
  drop_na() %>%
  dplyr::mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc.WT$Temperature), max(tpc.WT$Temperature), length.out = 100))) %>%
  ungroup() %>%
  dplyr::mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 27))

# calculate bootstrapped confidence intervals
boot_conf_preds.WT <- group_by(boot_preds.WT, temp) %>%
  dplyr::summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot.WT <- ggplot() +
  geom_line(aes(temp, .fitted), preds.WT, col = 'gray') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.WT, fill = 'gray', alpha = 0.3) +
  geom_point(aes(Temperature, rd_inverse), tpc.C, size = 2, alpha = 0.5,col = 'gray') +
  theme_bw(base_size = 12) +
    theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: WT");CI.plot.WT

```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit.WT) %>%
  # round for easy viewing
  mutate_all(round, 2)
```

## Print all parameters together 

Wildtype
```{r}
# calculate additional traits
calc_params(fit.WT) %>%
  # round for easy viewing
  mutate_all(round, 3)
```

Mixed 
```{r}
# calculate additional traits
calc_params(fit.MIX) %>%
  # round for easy viewing
  mutate_all(round, 3)
```

Cladocopium
```{r}
# calculate additional traits
calc_params(fit.C) %>%
  # round for easy viewing
  mutate_all(round, 3)
```

## Show all plots together 

```{r}
CI.plot.C
CI.plot.MIX
CI.plot.WT
```

```{r}
TPCs<-plot_grid(CI.plot.C, CI.plot.MIX, CI.plot.WT, nrow=1, ncol=3, align="vh")

ggsave("figures/temp_pi_curves/TPCs.png", TPCs, width=14, height=5, dpi=300)
```

## Plot TPC parameters

```{r}
#set plot colors
cols <- c("C"="orange", "MIX"="brown4", "WT"="gray")
  
# plot data and model fit
TPC.plot <- ggplot(data=tpc, aes(x=Temperature)) +
   
   geom_point(aes(x=Temperature, y=rd_inverse, color="C"), data=tpc.C, size = 2, alpha = 0.5) +
   geom_point(aes(x=Temperature, y=rd_inverse, color="MIX"), data=tpc.MIX, size = 2, alpha = 0.5) +
   geom_point(aes(x=Temperature, y=rd_inverse, color="WT"), data=tpc.WT, size = 2, alpha = 0.5) +
   
   geom_line(aes(temp, .fitted), preds.C, col = 'orange', size=2) +
   geom_line(aes(temp, .fitted), preds.MIX, col = "brown4", size=2) +
   geom_line(aes(temp, .fitted), preds.WT, col = "gray", size=2) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.C, fill = "orange", alpha = 0.3) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.MIX, fill = 'brown4', alpha = 0.3) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.WT, fill = "gray", alpha = 0.3) +
   #xlim(21,33)+
   #scale_x_continuous(breaks=c(13,23,26,28,32,35,37,40))+
   theme_bw(base_size = 12) +
   theme_classic()+
   scale_colour_manual(name="Symbiont",values=cols)+
   theme(legend.position = "right",
         panel.border = element_blank(), panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
   labs(x = 'Temperature (ºC)',
        y = expression("Respiration Rate"~nmol~O[2] ~larva^{-1}~min^{-1}));TPC.plot 

ggsave("figures/temp_pi_curves/TPC_AllTypes.png", TPC.plot, dpi=300, w=8, h=8, units="in")

```

## Confidence intervals of TPC parameters   

```{r}
broom::tidy(fit_nlsLM.C)
broom::tidy(fit_nlsLM.MIX)
broom::tidy(fit_nlsLM.WT)
```

Cladocopium parameters
```{r}
#Cladocopium
#calculate all the TPC parameters
C.extra_params <- calc_params(fit_nlsLM.C) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
C.ci_extra_params <- Boot(fit_nlsLM.C, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM.C)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  dplyr::rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
C.ci_extra_params <- left_join(C.ci_extra_params, C.extra_params)
C.ci_extra_params$Symbiont <- "Cladocopium"
```

Mixed parameters 
```{r}
#Mix
#calculate all the TPC parameters
MIX.extra_params <- calc_params(fit_nlsLM.MIX) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
MIX.ci_extra_params <- Boot(fit_nlsLM.MIX, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM.MIX)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  dplyr::rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')
  
#join the parameters and CIs  
MIX.ci_extra_params <- left_join(MIX.ci_extra_params, MIX.extra_params)
MIX.ci_extra_params$Symbiont <- "Mixed"
```

Wildtype parameters 
```{r}
#Mix
#calculate all the TPC parameters
WT.extra_params <- calc_params(fit_nlsLM.WT) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
WT.ci_extra_params <- Boot(fit_nlsLM.WT, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM.WT)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  dplyr::rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')
  
#join the parameters and CIs  
WT.ci_extra_params <- left_join(WT.ci_extra_params, WT.extra_params)
WT.ci_extra_params$Symbiont <- "Wildtype"
```

```{r}
#Join 
All_params <- rbind(C.ci_extra_params, MIX.ci_extra_params)
All_params <- rbind(All_params, WT.ci_extra_params)

All_params <- All_params %>% 
 mutate_if(is.numeric, round, 3)

#Plot all of the estimates
estimate.plots <- ggplot(All_params, aes(Symbiont, estimate, color=Symbiont)) +
  geom_point(size = 2) +
  scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_classic() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')+
  theme_classic();estimate.plots

ggsave("figures/temp_pi_curves/TPC_estimates.png", estimate.plots, dpi=300, w=8, h=8, units="in")
``` 

View specific values.
```{r}
All_params%>%
  filter(param=="ctmax")

All_params%>%
  filter(param=="thermal_safety_margin")

All_params%>%
  filter(param=="thermal_tolerance")
```



