---
title: "Respirometry Extraction with LoLinR - Hawaii 2023 - SDR TPCs"
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')

```

## Read in files  

Set the path of all respirometry files within the R project.   

```{r, warning=FALSE, message=FALSE}
path.p<-"data/tpc_sdr/runs" #location of files
```

Bring in the file names.
```{r, warning=FALSE, message=FALSE}
# bring in the respiration file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 
```

## Extract respiration rates  

Generate respiration data frames.

```{r, warning=FALSE, message=FALSE}
#generate a 6 column dataframe with specific column names
#respiration
Resp.R <- data.frame(matrix(NA, ncol=6))
colnames(Resp.R) <- c("Date", "Run","Sample.ID","Chamber.ID","Intercept", "umol.L.min")

Resp.Rb <- data.frame(matrix(NA, ncol=6))
colnames(Resp.Rb) <- c("Date", "Run","Sample.ID","Chamber.ID","Intercept", "umol.L.min")

```

Load in the sample information file. It is important to have chambers in the order "A1, B1, C1, D1", rather than "A1, A2, A3, A4".  

```{r, warning=FALSE, message=FALSE}
Sample.Info<-read.csv("data/tpc_sdr/TPC_SDR_Sample_Info.csv") #information on life stage, number of individuals, runs, etc.
rename <- Sample.Info$Chamber.ID
samp <- Sample.Info$Sample.ID
run <- str_sub(file.names, 17, 21) #grab run from file name
date <- str_sub(file.names, 1, str_length(file.names)-24) #grab date from file name
```

Run loop to extract slopes from respiration data. 

Remember to save data as umol/L NOT % saturation. 

```{r, results=FALSE, warning=FALSE, message=FALSE}
for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Resp.Data <-read.table(file.path(path.p,file.names[i]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Resp.Data$Temp <- Resp.Data[,31] #assigns temp column
  Resp.Data$Time.Min <- seq.int(from=0, to=((nrow(Resp.Data)*0.25)-0.25), by = 0.25) #set time in min
  #Resp.Data <- Resp.Data %>% #filters data by phase (respiration only)
     #filter(Time.Min > rtime[i])
  Resp.Data.N <- Resp.Data[,3:26] #subset desired columns

  #add column names back in 
  Resp.Data.N<-as.data.frame(Resp.Data.N)
  
  for(j in 1:(ncol(Resp.Data.N))){
    model <- rankLocReg(
      xall=Resp.Data$Time.Min, yall=as.numeric(Resp.Data.N[, j]), 
      alpha=0.4, method="pc", verbose=TRUE) #extract slopes, percentile rank method with minimum window size of 0.4. This means that in order to fit a slope, it has to encompass at least 40% of available datapoints. 
    
    pdf(paste0("output/tpc_sdr/RespirationPlots/",date[i], "_",run[i],"_",rename[j],"_regression_trunc.pdf")) #generate output file names
    plot(model)
    dev.off()
    
    Resp.Rb[j,1] <- as.character(date[i]) #stores the date
    Resp.Rb[j,2] <- as.character(run[i]) #stores the run number
    Resp.Rb[j,3] <- as.character(samp[j+(i-1)*ncol(Resp.Data.N)]) #stores the sample ID
    Resp.Rb[j,4] <- as.character(rename[j]) #stores the chamber ID
    Resp.Rb[j,5:6] <- model$allRegs[i,c(4,5)] #inserts slope and intercept in the dataframe
    
  }
  Resp.R <- rbind(Resp.R, Resp.Rb) #bind final data frame
}

```

Calculate average temperature of each run and export to a table.  
```{r}
# list files 
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE)) 

#generate matrix to populate
Temp.R <- data.frame(matrix(NA, ncol=3))
colnames(Temp.R) <- c("Date", "Run","Temp.C")

Temp.Rb <- data.frame(matrix(NA, ncol=3))
colnames(Temp.Rb) <- c("Date", "Run","Temp.C")

#read in temps and generate mean values  
for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Temp.Data <-read.table(file.path(path.p,file.names[i]), skip = 56, header=T, sep=",", na.string="NA", fill = TRUE, as.is=TRUE, fileEncoding="latin1") #reads in the data files
  Temp.Data$Temp <- Temp.Data[,31] #assigns temp column
  
    Temp.Rb[j,1] <- as.character(date[i]) #stores the date
    Temp.Rb[j,2] <- as.character(run[i]) #stores the run number
    Temp.Rb[j,3] <- mean(Temp.Data$Temp) #stores the sample ID
    
    Temp.R <- rbind(Temp.R, Temp.Rb) #bind final data frame
    Temp.R <- na.omit(Temp.R)
  
}

#remove the first empty row  
#Temp.R <- Temp.R[-1,]

write.csv(Temp.R, paste0("output/tpc_sdr/runs_temp.csv")) #save respiration rate data
```

Save respiration data frames.  

```{r, results=FALSE, warning=FALSE, message=FALSE}
Resp.R <- Resp.R[-1,] #remove empty column

write.csv(Resp.R, paste0("output/tpc_sdr/tpc_sdr_respiration_rates_raw.csv")) #save respiration rate data

plot(Resp.R$umol.L.min~as.factor(Resp.R$Run), side = 2, las = 2, xlab="" )

```

```{r, warning=FALSE, message=FALSE}
Resp.Rates <- read.csv(file = "output/tpc_sdr/tpc_sdr_respiration_rates_raw.csv") #read file back in so slopes don't have to be generated every time
Resp.Rates = subset(Resp.Rates, select = -c(X) ) #remove empty column

#format "run" column
Resp.Rates<-Resp.Rates %>%
  mutate(Run = str_sub(Run, 4, -1))
Resp.Rates$Run<-as.integer(Resp.Rates$Run) #format as # rather than run #, set as integer
```

## Standardize and normalize  

Merge P and R files with sample info for testing and manipulation.  

```{r, warning=FALSE, message=FALSE}
Sample.Info$Date<-as.factor(Sample.Info$Date)
Resp.Rates$Date<-as.factor(Resp.Rates$Date)
oxygen<-dplyr::full_join(Sample.Info, Resp.Rates) #add respiration data
colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Resp.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'umol.L.min'] <- 'R.umol.L.min' #rename to specify R

```

Remove samples that have inaccurate slope extraction for respiration rates (determined by PDF output files)  

```{r}
#Run 3 B3
#Run 3 D4
#Run 7 D6 
#Run 8 B4 
#Run 12 A3
#Run 12 B5
#Run 13 D6
#Run 15 A4
#Run 15 D6
#Run 6 D4
#Run 8 C4
#Run 10 D4

#oxygen<-oxygen%>%
  #mutate(code=paste(Run, Chamber.ID))%>%
  #filter(!code=="3 B3")%>%
  #filter(!code=="3 D4")%>%
  #filter(!code=="7 D6")%>%
  #filter(!code=="8 B4")%>%
  #filter(!code=="12 A3")%>%
  #filter(!code=="12 B5")%>%
  #filter(!code=="13 D6")%>%
  #filter(!code=="15 A4")%>%
  #filter(!code=="15 D6")%>%
  #filter(!code=="6 D5")%>%
  #filter(!code=="8 C4")%>%
  #filter(!code=="10 D4")%>%
  #select(!code)
```

Account for volume to obtain umol per minute.   









LEFT OFF HERE , NEED TO QC BELOW SECTION 


```{r, results=TRUE, warning=FALSE, message=FALSE}

#Account for chamber volume to convert from umol L-1 m-1 to umol m-1. This removes per Liter
oxygen$R.umol.min <- oxygen$R.umol.L.min * oxygen$Volume #calculate

plot(oxygen$Run, oxygen$R.umol.L.min, xlab="Run", ylab="rate") #resp rates
plot(oxygen$Run, oxygen$R.umol.min, xlab="Run", ylab="rate") #resp rates
```

Substract blank values. Average blank calculated for each run. Display mean blank value.  
```{r}
blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only
plot(as.factor(blank_data$Run), blank_data$R.umol.min, xlab="Run", ylab="umol O2 min-1") #blanks during dark phase

#display mean blankvalues
mean(blank_data$R.umol.min, na.rm=TRUE) #mean R phase blanks

resp.blnk <- aggregate(R.umol.min ~ Date*Run, data=blank_data, mean) #calculate average blank during light for each run
colnames(resp.blnk)[colnames(resp.blnk) == 'R.umol.min'] <- 'R.Blank.umol.min' #rename to specify blank for R

oxygen <- full_join(oxygen, resp.blnk) #add R blanks to master
```

Subtract blank values to generate a "corrected" value for umol O2 min-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$R.umol.min.corr<-oxygen$R.umol.min-oxygen$R.Blank.umol.min #subtract R blanks
```

```{r}
plot(oxygen$R.umol.min.corr)
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Type == "Sample") #isolate only biological samples and drop unused factor levels 
oxygen.bio <- droplevels(oxygen.bio) #drop unused factor levels

#respiration
oxygen.bio$R.umol.org.min <- oxygen.bio$R.umol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism 
oxygen.bio$R.nmol.org.min <- oxygen.bio$R.umol.org.min*1000 #calculate nanomoles 

```

Plot values.  
```{r}
plot(oxygen.bio$R.nmol.org.min)
```

Save as .csv file.  
```{r, warning=FALSE, message=FALSE}
write.csv(oxygen.bio, paste0("output/tpc_sdr/calculated_normalized_rates.csv")) #save final file
```


