---
title: "Respirometry Plotting and Analysis - Hawaii 2023" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')
if ("nlme" %in% rownames(installed.packages()) == 'FALSE') install.packages('nlme')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
library('nlme')
```

# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
tpc_data<-read.csv("output/tpc_sdr/calculated_normalized_rates.csv") #load data
sample_info<-read.csv("data/tpc_sdr/TPC_SDR_Sample_Info.csv")
```

Keep only experimental plate (removing test plates).  
```{r}
tpc_data<-tpc_data%>%
  filter(!Date=="20230621")
```

Separate project specific data.  
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
tpc_data<-tpc_data[!is.na(tpc_data$Type),]

#format columns
tpc_data$Temperature<-as.factor(tpc_data$Temperature)
tpc_data$Tank<-as.factor(tpc_data$Tank)
tpc_data$Symbiont<-as.factor(tpc_data$Symbiont)
tpc_data$SDR<-as.factor(tpc_data$SDR)
tpc_data$Plate<-as.factor(tpc_data$Plate)
```

Look for outliers in the data.  

```{r}
boxplot(tpc_data$R.nmol.org.min)
tpc_data<-tpc_data%>%filter(R.nmol.org.min > -0.03)
boxplot(tpc_data$R.nmol.org.min)
```

If a respiration value is above 0 (indicating oxygen production), turn it into 0. 
```{r}
tpc_data<-tpc_data%>%
  mutate(R.nmol.org.min=if_else(R.nmol.org.min>0, 0, R.nmol.org.min))
```

Next log transform the data.  
```{r}
tpc_data<-tpc_data%>%
  mutate(log.R=log10(1+R.nmol.org.min))

boxplot(tpc_data$log.R)
```

Calculate the inverse for oxygen consumption expressed as a positive value. 
```{r}
tpc_data<-tpc_data%>%
  mutate(log.R.inverse=log.R*-1)
```

View data now. 
```{r}
boxplot(tpc_data$log.R.inverse)
```

Calculate mean temperature values and reassign measurement treatment values if necessary.  

```{r}
resp.temps<-read.csv("output/tpc_sdr/runs_temp.csv")
resp.temps = subset(resp.temps, select = -c(X) ) #remove empty column
resp.temps<-resp.temps%>%
  filter(!Date=="20230621")

#format "run" column
resp.temps<-resp.temps %>% #Format as "#" rather than "Run#"
  mutate(Plate = str_sub(Plate, 6, -1))
resp.temps$Plate<-as.integer(resp.temps$Plate) #format as integer
resp.temps$Plate<-as.factor(resp.temps$Plate) #format as factor
```

Add temperature data to master data frame.  
```{r}
tpc_data<-left_join(tpc_data,resp.temps)

#round to 0.1°C 
tpc_data<-tpc_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

Add temporary corrected temperature. 
```{r}
tpc_data$Temp_corrected<-sample_info$Temp_corrected[match(tpc_data$Plate, sample_info$Plate)]
```

# Plot respiration over temperature treatments  

Plot data with means   
```{r}
r_plot1<-tpc_data %>%
    filter(!Temperature==28)%>%
    filter(!Temperature==39)%>%
    filter(!Temperature==40)%>%
    group_by(Temp.C, Symbiont, Date)%>%
    dplyr::summarise(mean=mean(log.R.inverse, na.rm=TRUE), sd=sd(log.R.inverse, na.rm=TRUE), N=length(log.R.inverse), se=sd/sqrt(N))%>%
    #mutate(mean=mean*-1)%>%
    
    ggplot(., aes(x = Temp.C, y = mean, group=interaction(Temp.C, Symbiont))) +
    facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", linewidth=0.75)+
    geom_point(aes(group=interaction(Temp.C, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Temp.C, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot1
```

Plot data with a loess line   
```{r}
r_plot2<-tpc_data %>%
  filter(!Temperature==28)%>%
  filter(!Temperature==39)%>%
  filter(!Temperature==40)%>%
    
    ggplot(., aes(x = Temp.C, y = (log.R.inverse), group=Symbiont)) +
    facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=FALSE)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot2
```

Plot data with means   
```{r}
r_plot3<-tpc_data %>%
  filter(!Temperature==28)%>%
  filter(!Temperature==39)%>%
  filter(!Temperature==40)%>%
    
    ggplot(., aes(x = Temp.C, y = (log.R.inverse), group=interaction(Temp.C, Symbiont))) +
    facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temp.C, Symbiont), colour=Symbiont), size=4, position = position_jitterdodge(0.4)) + 
    #geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Temperature, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot3
```

## Running anova on respiration rates 

```{r}
aov<-tpc_data%>%
    #filter(!Temperature==28)%>%
    #filter(!Temperature==39)%>%
    #filter(!Temperature==40)%>%
    aov(log.R.inverse~Temp.C*Symbiont, data=.)

summary(aov)
```
Effect of temp on respiration. 

# Outlier removal 

## Look for outliers 

Change respiration to inverse. 
```{r}
tpc_data<-tpc_data%>%
  mutate(R.inverse=R.nmol.org.min*-1)
```

Filter temps that didn't work. Keeping all right now.  
```{r}
tpc_data_filt<-tpc_data %>%
  filter(!Temperature==28)%>%
  filter(!Temperature==39)%>%
  filter(!Temperature==40)
```

View and remove outliers.   
```{r, warning=FALSE, message=FALSE}
#identify outliers by Temperature and Treatment groups
outlier.plot <- ggbetweenstats(tpc_data_filt, Temp.C, log.R.inverse, outlier.tagging = TRUE)

outlier.plot

ggsave("figures/tpc_sdr/outliersbygroup.png", outlier.plot, dpi=300, w=16, h=8, units="in")

#set quantile values
q <- c(0.1, 0.9)

# calculate quantile values by Temperature and Treatment groups
Quants <- tpc_data_filt %>%
  group_by(Temp.C, Symbiont) %>%
  summarize(quant25 = quantile(log.R.inverse, probs = q[1], na.rm=TRUE),
            quant75 = quantile(log.R.inverse, probs = q[2], na.rm=TRUE),
            IQRbyGroup=IQR(log.R.inverse, na.rm=TRUE))

# add a group name by pasting Temperature and Treatment
Quants$group <-paste0(Quants$Temp.C,"_", Quants$Symbiont)

#Calculate Quantile upper and lower ranges 
Quants$upper <-  Quants$quant75+1.5*Quants$IQRbyGroup # Upper Range  
Quants$lower <- Quants$quant25-1.5*Quants$IQRbyGroup # Lower Range

#join outlier cutoffs with rate data
tpc_data_filt <- left_join(tpc_data_filt, Quants)

#remove outliers from rates
tpc_data_outlier <- tpc_data_filt %>%
  filter(log.R.inverse < upper) %>%
  filter(log.R.inverse > lower) #%>%
  #filter(rate < 0.125)

# Visualize data following outlier removal
outlier.plot2 <- ggbetweenstats(tpc_data_outlier, Temp.C, log.R.inverse, outlier.tagging = TRUE)
outlier.plot2

ggsave("figures/tpc_sdr/removed_outliersbygroup.png", outlier.plot2, dpi=300, w=16, h=8, units="in")
```

Use this if you do not want outlier removal from above. 
```{r}
tpc_data_filt<-tpc_data_outlier %>%
  filter(!Temperature==28)%>%
  filter(!Temperature==39)%>%
  filter(!Temperature==40)

outlier.plot2 <- ggbetweenstats(tpc_data_filt, Temp.C, log.R.inverse, outlier.tagging = TRUE)
outlier.plot2

ggsave("figures/tpc_sdr/removed_outliersbygroup.png", outlier.plot2, dpi=300, w=16, h=8, units="in")
```

## Plot data after outlier removal 

Plot data with means   
```{r}
r_plot4<-tpc_data_filt %>%
    group_by(Temp.C, Symbiont, Date)%>%
    dplyr::summarise(mean=mean(log.R.inverse, na.rm=TRUE), sd=sd(log.R.inverse, na.rm=TRUE), N=length(log.R.inverse), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = Temp.C, y = mean, group=interaction(Temp.C, Symbiont))) +
    facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(group=interaction(Temp.C, Symbiont), colour=Symbiont), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=interaction(Temp.C, Symbiont)), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot4

ggsave("figures/tpc_sdr/tpc_means.png", r_plot4, dpi=300, w=7, h=5, units="in")
```

Plot data with a loess line   
```{r}
r_plot5<-tpc_data_filt %>%
    
    ggplot(., aes(x = as.numeric(Temp.C), y = (log.R.inverse), group=Symbiont)) +
    facet_wrap(~Date)+
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_point(aes(color=Symbiont, group=Symbiont), size=3, shape=21)+
    geom_smooth(aes(color=Symbiont, group=Symbiont), method="loess", alpha=0.4, se=FALSE, size=2)+
    xlab("Temperature") + 
    scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
    ylab(expression(bold(paste("log R (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); r_plot5

ggsave("figures/tpc_sdr/tpc_dots.png", r_plot5, dpi=300, w=7, h=5, units="in")
```

```{r}
summary(aov(log.R.inverse~Temp.C*Symbiont, data=tpc_data_filt))
```
Effect of temperature. 

# Fitting TPCs 

TPC fitting 
Padifeld et al **rTPC and nls.multstart: A new pipeline to fit thermal performance curves in r**  
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13585  

Sharpe Schoolfield 1981 model
Schoolfield, R. M., Sharpe, P. J. H., & Magnuson, C. E. (1981). Non-linear regression of biological temperature-dependent rate models based on absolute reaction-rate theory. Journal of theoretical biology, 88(4), 719-731. https://doi.org/10.1016/0022-5193(81)90246-0

## Run for all symbionts 

```{r}
tpc_data_filt.All<-tpc_data_filt #%>%
  #filter(R.inverse<0.03)%>%
  #filter(!c(R.inverse>0.02 & Temp_corrected==37))%>%
  #filter(!c(R.inverse>0.02 & Temp_corrected==13))%>%
  #filter(!c(R.inverse< -0.01 & Temp_corrected==23))%>%
  #filter(!c(R.inverse< 0 & Temp_corrected==32))%>%
  #filter(!c(R.inverse< -0.005 & Temp_corrected==37))%>%
  #filter(!c(R.inverse< -0.005 & Temp_corrected==34))

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc_data_filt.All$Temp.C, tpc_data_filt.All$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc_data_filt.All$Temp.C, tpc_data_filt.All$log.R.inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc_data_filt.All$Temp.C, tpc_data_filt.All$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

fit <- nls_multstart(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 26),
                                                     data = tpc_data_filt.All,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit

# predict new data
new_data <- data.frame(temp = seq(min(tpc_data_filt.All$Temp.C), max(tpc_data_filt.All$Temp.C), by=((max(tpc_data_filt.All$Temp.C)-min(tpc_data_filt.All$Temp.C))/(nrow(tpc_data_filt.All)-1))))

preds <- augment(fit, newdata = new_data)

# plot data and model fit
ggplot(tpc_data_filt.All, aes(x=Temp.C, y=log.R.inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds, col = 'black') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures')
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM <- minpack.lm::nlsLM(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref, e, eh, th, tref = 26),
                        data = tpc_data_filt.All,
                        start = coef(fit),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc_data_filt.All)))

# bootstrap using case resampling
boot <- Boot(fit_nlsLM, method = 'case')

# look at the data
head(boot$t)


# create predictions of each bootstrapped model
boot_preds <- boot$t %>%
  as.data.frame() %>%
  drop_na() %>%
  dplyr::mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc_data_filt.All$Temp.C), max(tpc_data_filt.All$Temp.C), length.out = 100))) %>%
  ungroup() %>%
 dplyr::mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 26))

# calculate bootstrapped confidence intervals
boot_conf_preds <- group_by(boot_preds, temp) %>%
  dplyr::summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), preds, col = 'black') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds, fill = 'black', alpha = 0.3) +
  geom_point(aes(Temp.C, log.R.inverse), tpc_data_filt.All, size = 2, alpha = 0.5,col = 'black') +
  theme_bw(base_size = 12) +
  theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'log Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: All");CI.plot 

```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit) %>%
  # round for easy viewing
  mutate_all(round, 2)
```


## Run for C symbionts 

```{r}
tpc_data_filt.C<-tpc_data_filt%>%
  filter(Symbiont=="C")

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc_data_filt.C$Temp.C, tpc_data_filt.C$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc_data_filt.C$Temp.C, tpc_data_filt.C$log.R.inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc_data_filt.C$Temp.C, tpc_data_filt.C$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

fit.C <- nls_multstart(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 26),
                                                     data = tpc_data_filt.C,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit.C

# predict new data
new_data.C <- data.frame(temp = seq(min(tpc_data_filt.C$Temp.C), max(tpc_data_filt.C$Temp.C), by=((max(tpc_data_filt.C$Temp.C)-min(tpc_data_filt.C$Temp.C))/(nrow(tpc_data_filt.C)-1))))

preds.C <- augment(fit.C, newdata = new_data.C)

# plot data and model fit
ggplot(tpc_data_filt.C, aes(x=Temp.C, y=log.R.inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds.C, col = 'orange') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures: C')
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM.C <- minpack.lm::nlsLM(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref, e, eh, th, tref = 26),
                        data = tpc_data_filt.C,
                        start = coef(fit.C),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc_data_filt.C)))

# bootstrap using case resampling
boot.C <- Boot(fit_nlsLM.C, method = 'case')

# look at the data
head(boot.C$t)

# create predictions of each bootstrapped model
boot_preds.C <- boot.C$t %>%
  as.data.frame() %>%
  drop_na() %>%
  dplyr::mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc_data_filt.C$Temp.C), max(tpc_data_filt.C$Temp.C), length.out = 100))) %>%
  ungroup() %>%
  dplyr::mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 26))

# calculate bootstrapped confidence intervals
boot_conf_preds.C <- group_by(boot_preds.C, temp) %>%
  dplyr::summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot.C <- ggplot() +
  geom_line(aes(temp, .fitted), preds.C, col = 'orange') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.C, fill = 'orange', alpha = 0.3) +
  geom_point(aes(Temp.C, log.R.inverse), tpc_data_filt.C, size = 2, alpha = 0.5,col = 'orange') +
  theme_bw(base_size = 12) +
    theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'log Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: C");CI.plot.C

```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit.C) %>%
  # round for easy viewing
  mutate_all(round, 2)
```

## Run for MIX symbionts 

```{r}
tpc_data_filt.MIX<-tpc_data_filt%>%
  filter(Symbiont=="MIX")

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc_data_filt.MIX$Temp.C, tpc_data_filt.MIX$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc_data_filt.MIX$Temp.C, tpc_data_filt.MIX$log.R.inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc_data_filt.MIX$Temp.C, tpc_data_filt.MIX$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

fit.MIX <- nls_multstart(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 26),
                                                     data = tpc_data_filt.MIX,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit.MIX

# predict new data
new_data.MIX <- data.frame(temp = seq(min(tpc_data_filt.MIX$Temp.C), max(tpc_data_filt.MIX$Temp.C), by=((max(tpc_data_filt.MIX$Temp.C)-min(tpc_data_filt.MIX$Temp.C))/(nrow(tpc_data_filt.MIX)-1))))

preds.MIX <- augment(fit.MIX, newdata = new_data.MIX)

# plot data and model fit
ggplot(tpc_data_filt.MIX, aes(x=Temp.C, y=log.R.inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds.MIX, col = 'brown4') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'log Respiration across temperatures: MIX')
```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit.MIX) %>%
  # round for easy viewing
  mutate_all(round, 2)
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM.MIX <- minpack.lm::nlsLM(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref, e, eh, th, tref = 26),
                        data = tpc_data_filt.MIX,
                        start = coef(fit.MIX),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc_data_filt.MIX)))

# bootstrap using case resampling
boot.MIX <- Boot(fit_nlsLM.MIX, method = 'case')

# look at the data
head(boot.MIX$t)

# create predictions of each bootstrapped model
boot_preds.MIX <- boot.MIX$t %>%
  as.data.frame() %>%
  drop_na() %>%
  dplyr::mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc_data_filt.MIX$Temp.C), max(tpc_data_filt.MIX$Temp.C), length.out = 100))) %>%
  ungroup() %>%
  dplyr::mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 26))

# calculate bootstrapped confidence intervals
boot_conf_preds.MIX <- group_by(boot_preds.MIX, temp) %>%
  dplyr::summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot.MIX <- ggplot() +
  geom_line(aes(temp, .fitted), preds.MIX, col = 'brown4') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.MIX, fill = 'brown4', alpha = 0.3) +
  geom_point(aes(Temp.C, log.R.inverse), tpc_data_filt.MIX, size = 2, alpha = 0.5,col = 'brown4') +
  theme_bw(base_size = 12) +
    theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'log Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: MIX");CI.plot.MIX

```

## Run for WT symbionts 

```{r}
tpc_data_filt.WT<-tpc_data_filt%>%
  filter(Symbiont=="WT")

# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(tpc_data_filt.WT$Temp.C, tpc_data_filt.WT$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(tpc_data_filt.WT$Temp.C, tpc_data_filt.WT$log.R.inverse, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(tpc_data_filt.WT$Temp.C, tpc_data_filt.WT$log.R.inverse, model_name = 'sharpeschoolhigh_1981')

#manually specify start values since eh is NA - try eh=5 
#start_vals["eh"] <- 5

fit.WT <- nls_multstart(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref,e,eh,th, tref = 26),
                                                     data = tpc_data_filt.WT,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit.WT

# predict new data
new_data.WT <- data.frame(temp = seq(min(tpc_data_filt.WT$Temp.C), max(tpc_data_filt.WT$Temp.C), by=((max(tpc_data_filt.WT$Temp.C)-min(tpc_data_filt.WT$Temp.C))/(nrow(tpc_data_filt.WT)-1))))

preds.WT <- augment(fit.WT, newdata = new_data.WT)

# plot data and model fit
ggplot(tpc_data_filt.WT, aes(x=Temp.C, y=log.R.inverse)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds.WT, col = 'gray') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'log Respiration across temperatures: WT')
```

Print parameters 
```{r}
# calculate additional traits
calc_params(fit.WT) %>%
  # round for easy viewing
  mutate_all(round, 2)
```

Bootstrap for confidence intervals. 
```{r}
#refit 
fit_nlsLM.WT <- minpack.lm::nlsLM(log.R.inverse~sharpeschoolhigh_1981(temp = Temp.C, r_tref, e, eh, th, tref = 26),
                        data = tpc_data_filt.WT,
                        start = coef(fit.WT),
                        lower = low_lims,
                        upper = upper_lims,
                        weights = rep(1, times = nrow(tpc_data_filt.WT)))

# bootstrap using case resampling
boot.WT <- Boot(fit_nlsLM.WT, method = 'case')

# look at the data
head(boot.WT$t)

# create predictions of each bootstrapped model
boot_preds.WT <- boot.WT$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(tpc_data_filt.WT$Temp.C), max(tpc_data_filt.WT$Temp.C), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 26))

# calculate bootstrapped confidence intervals
boot_conf_preds.WT <- group_by(boot_preds.WT, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
CI.plot.WT <- ggplot() +
  geom_line(aes(temp, .fitted), preds.MIX, col = 'gray') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.MIX, fill = 'gray', alpha = 0.3) +
  geom_point(aes(Temp.C, log.R.inverse), tpc_data_filt.MIX, size = 2, alpha = 0.5,col = 'gray') +
  theme_bw(base_size = 12) +
    theme_classic()+
  labs(x = 'Temperature (ºC)',
       y = 'log Respiration Rate (nmol O2/larva/min)', 
       title = "TPC: WT");CI.plot.WT

```

## Print all parameters together 

All together
```{r}
# calculate additional traits
calc_params(fit) %>%
  # round for easy viewing
  mutate_all(round, 3)
```

Wildtype
```{r}
# calculate additional traits
calc_params(fit.WT) %>%
  # round for easy viewing
  mutate_all(round, 3)
```

Mixed 
```{r}
# calculate additional traits
calc_params(fit.MIX) %>%
  # round for easy viewing
  mutate_all(round, 3)
```

Cladocopium
```{r}
# calculate additional traits
calc_params(fit.C) %>%
  # round for easy viewing
  mutate_all(round, 3)
```


## Show all plots together 

```{r}
CI.plot
CI.plot.C
CI.plot.MIX
CI.plot.WT
```

```{r}
TPCs<-plot_grid(CI.plot, CI.plot.C, CI.plot.MIX, CI.plot.WT, nrow=1, ncol=4, align="vh")
#TPCs<-plot_grid(CI.plot, CI.plot.C, CI.plot.MIX, nrow=1, ncol=4, align="vh")

ggsave("figures/tpc_sdr/TPCs.png", TPCs, width=14, height=5, dpi=300)
```

## Plot TPC parameters

```{r}
#set plot colors
cols <- c("C"="orange", "MIX"="brown4", "WT"="gray")
  
# plot data and model fit
TPC.plot <- ggplot(data=tpc_data_filt.All, aes(x=Temp.C)) +
   
   geom_point(aes(x=Temp.C, y=log.R.inverse, color="C"), data=tpc_data_filt.C, size = 2, alpha = 0.5) +
   geom_point(aes(x=Temp.C, y=log.R.inverse, color="MIX"), data=tpc_data_filt.MIX, size = 2, alpha = 0.5) +
   geom_point(aes(x=Temp_corrected, y=log.R.inverse, color="WT"), data=tpc_data_filt.WT, size = 2, alpha = 0.5) +
   
   geom_line(aes(temp, .fitted), preds.C, col = 'orange', size=2) +
   geom_line(aes(temp, .fitted), preds.MIX, col = "brown4", size=2) +
   geom_line(aes(temp, .fitted), preds.WT, col = "gray", size=2) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.C, fill = "orange", alpha = 0.3) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.MIX, fill = 'brown4', alpha = 0.3) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds.WT, fill = "gray", alpha = 0.3) +
   #xlim(21,33)+
   scale_x_continuous(breaks=c(13,23,26,28,32,35,37,40))+
   theme_bw(base_size = 12) +
   theme_classic()+
   scale_colour_manual(name="Symbiont",values=cols)+
   theme(legend.position = "right",
         panel.border = element_blank(), panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
   labs(x = 'Temperature (ºC)',
        y = expression("log Respiration Rate"~nmol~O[2] ~larva^{-1}~min^{-1}));TPC.plot 

ggsave("figures/tpc_sdr/TPC_AllTypes.png", TPC.plot, dpi=300, w=8, h=8, units="in")

```

## Confidence intervals of TPC parameters   

```{r}
broom::tidy(fit_nlsLM)
broom::tidy(fit_nlsLM.C)
broom::tidy(fit_nlsLM.MIX)
broom::tidy(fit_nlsLM.WT)
```

All symbiont parameters
```{r}
#all
#calculate all the TPC parameters
All.extra_params <- calc_params(fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
All.ci_extra_params <- Boot(fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  dplyr::rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  dplyr::mutate(method = 'case bootstrap')

#join the parameters and CIs  
All.ci_extra_params <- left_join(All.ci_extra_params, All.extra_params)
All.ci_extra_params$Symbiont <- "All"
```

Cladocopium parameters
```{r}
#Cladocopium
#calculate all the TPC parameters
C.extra_params <- calc_params(fit_nlsLM.C) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
C.ci_extra_params <- Boot(fit_nlsLM.C, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM.C)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  dplyr::rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
C.ci_extra_params <- left_join(C.ci_extra_params, C.extra_params)
C.ci_extra_params$Symbiont <- "Cladocopium"
```

Mixed parameters 
```{r}
#Mix
#calculate all the TPC parameters
MIX.extra_params <- calc_params(fit_nlsLM.MIX) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
MIX.ci_extra_params <- Boot(fit_nlsLM.MIX, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM.MIX)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  dplyr::rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')
  
#join the parameters and CIs  
MIX.ci_extra_params <- left_join(MIX.ci_extra_params, MIX.extra_params)
MIX.ci_extra_params$Symbiont <- "Mixed"
```

Wildtype parameters 
```{r}
#Mix
#calculate all the TPC parameters
WT.extra_params <- calc_params(fit_nlsLM.WT) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
WT.ci_extra_params <- Boot(fit_nlsLM.WT, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(fit_nlsLM.WT)), R = 200, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')
  
#join the parameters and CIs  
WT.ci_extra_params <- left_join(WT.ci_extra_params, WT.extra_params)
WT.ci_extra_params$Symbiont <- "Wildtype"
```

```{r}
#Join 
All_params <- rbind(C.ci_extra_params, MIX.ci_extra_params)
All_params <- rbind(All_params, WT.ci_extra_params)

All_params <- All_params %>% 
 mutate_if(is.numeric, round, 3)

#Plot all of the estimates
estimate.plots <- ggplot(All_params, aes(Symbiont, estimate, color=Symbiont)) +
  geom_point(size = 2) +
  scale_color_manual(name="Symbiont", values=c("orange","brown4", "gray"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_classic() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')+
  theme_classic();estimate.plots

ggsave("figures/tpc_sdr/TPC_estimates.png", estimate.plots, dpi=300, w=8, h=8, units="in")
``` 

View values.
```{r}
All_params%>%
  filter(param=="topt")

All_params%>%
  filter(param=="thermal_safety_margin")

All_params%>%
  filter(param=="ctmax")

All_params%>%
  filter(param=="eh")
```

View parameters estimated from all larvae together. 
```{r}
All_symbiont_params <- All.ci_extra_params %>% 
 mutate_if(is.numeric, round, 3)

#Plot all of the estimates
estimate.plots_all <- ggplot(All_symbiont_params, aes(y=estimate, x=Symbiont)) +
  geom_point(size = 2, color="black") +
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_classic() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')+
  theme_classic();estimate.plots_all

ggsave("figures/tpc_sdr/TPC_estimates_all_symbionts.png", estimate.plots_all, dpi=300, w=8, h=8, units="in")
```



